### 2018-12-10 main.tex
Checkpoint conversion from GDocs

\documentclass[12pt]{article}

\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{hyperref}

\newif\ifdraft\drafttrue

\newcommand{\bcp}[1]{{\bf #1}}

\begin{document}

\title{\Huge ACM and Climate Change}
\author{Benjamin C. Pierce (University of Pennsylvania) \\
Michael Hicks (University of Maryland) \\
Crista Lopes (University California, Irvine) \\
Jens Palsberg (University of California, Los Angeles)}

\maketitle

ACM’s vibrant conference culture has a dark side: it contributes
significantly to climate change.

\begin{figure}[tb]
\centering
\includegraphics[width=4in]{wmo_stripes.png}
\caption{Annual global temperatures, 1850 to 2018 (range from blue to red:
  1.35 degrees C) \cite{WarmingStripes}}
\label{fig:warming}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[width=2in]{IPCC-cover.png}
\caption{2018 IPCC report \cite{IPCC18}.  Key finding: Emissions must
  decline by at least 40 percent by 2030 and reach net zero by 2050, if we
  are to hold warming to 1.5 degrees.  ``No documented historic precedent''
  for the transformation of the world economy needed to achieve this.  }
\label{fig:warming}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[width=6in]{SIGPLAN-confs.png}
\caption{Carbon footprint per participant for travel to recent SIGPLAN
  conferences\bcp{Get better file from Crista}} 
\label{fig:confs}
\end{figure}

There is a broad scientific consensus that the earth is warming due to human
emissions of carbon dioxide and other greenhouse gases (Figure
\ref{fig:warming}). This is not a problem we can leave to future
generations: we are up against it now. Indeed, the UN’s Intergovernmental
Panel on Climate Change (IPCC) recently reported that a 40\% decrease in
emissions by 2030 is needed to avoid irreversible damage
(Figure~\ref{IPCC-cover}). Achieving this will require urgent and sustained
commitment at all levels of society, including non-governmental communities
such as universities, companies, and scientific societies like ACM.

Why ACM? Because one of ACM’s main activities is organizing international
conferences, and these produce significant emissions of greenhouse gases,
particularly from air travel. For example, a single round-trip flight from
Philadelphia to Paris emits the equivalent of about 2 tons of carbon dioxide
(CO2e) per passenger. Compare this against the fact that the average US
resident emits about 16.5 tons per year, while the average European emits
about 7 tons [10].  Moreover, these emissions have no near-term
technological fix because jet fuel is difficult to replace with renewable
energy sources [7]. So reducing the CO2e footprint of conferences is going
to require a reduction in the footprint of the attendees’ air travel.

We are the members of an {\em ad hoc} Committee on Climate Change within
SIGPLAN (ACM’s Special Interest Group on Programming Languages) that was
formed in 2016 to consider possible strategies for mitigating the emissions
of our conferences [3]. Among the ones we consider most attractive are:
making it easier to use carbon offsets to mitigate aviation emissions for
conferences, choosing conference locations with an eye to minimizing
emissions, co-locating or merging conferences, virtualizing program
committees, and supporting remote participation.

Buying carbon offsets to reduce the impact of air travel is an ``easy win''
[4].  The idea of carbon offsets is to pay somebody to plant trees, capture
methane (from landfills or cattle herds, for example), buy cookstoves to
replace inefficient open-fire cooking in poor communities, or carry out
other actions that remove greenhouse gases from the air or prevent their
emission. At the moment, the average air travel to an international
conference can be offset by paying around \$30 [4].  ACM could streamline
the purchase of such offsets and make offsetting the default by folding them
into its conference registration fees.

Unfortunately, carbon offsetting is only a short-term mitigation, since the
potential supply of offsets is far less than the potential demand [11]. We
must also begin significantly reducing emissions, which means reducing the
distance traveled by the sum of all conference participants.

One approach to doing so is to choose conference locations that are close to
the center of mass of their potential participants.  For example, SIGGRAPH
(the Special Interest Group on Graphics) locates in Los Angeles every other
year---an easy call as this is near an international airport and a major hub
of the worldwide entertainment industry. For other research areas, the
effects of location choice are less obvious. For example,
Figure~\ref{fig:SIGPLAN-confs} plots the 
locations of the four major SIGPLAN conferences over the past five years,
colored according to the estimated average carbon footprint per
participant. The figure shows that there were significant differences across
conferences and across locations---indeed, it is a factor of two between the
lowest and highest. Why is this? One reason might be that the locus of a
conference’s population is closer to one location than another, and changing
locations reduces the total cost. Another might be the presence of a local
population that will only attend if the conference is very close to
them. Our preliminary analysis suggests it may be both. The larger question
is how to use this kind of information to choose the best locations over
time, for a diffuse population.

Another natural idea is to co-locate or consolidate conferences.  This is
already standard practice in some communities---e.g., SIGGRAPH, SC, and CVPR
are all ``the'' conference for their areas.  More communities---for example,
the programming languages and security communities, each of which holds four
main conferences every year---could consider following suit.  A different
twist on this idea would be to hold a single conference in two locations
simultaneously (e.g., Boston and Paris), with a video link between the
sites. Switching to virtual program committee meetings for conferences is
another obvious step.  Indeed, this switch is already happening: we found in
a recent poll of ACM SIG chairs that, across ACM, a majority of the SIGs’
main conferences now use online PC meetings rather than in-person ones.

We can attend some conferences online, saving both emissions and travel
time.  Indeed, SIGPLAN already livestreams all its four main conferences,
while SIGACCESS and SIGCHI have experimented with telepresence robots at
their main conferences [8,9]. Journal-first initiatives, which make
conference attendance optional, could also help.  Looking further ahead,
some communities are already experimenting with fully virtual conferences
[6].  ACM members can contribute to the development of technologies for
connecting people effectively at a distance.

Each of these strategies will have some beneficial effect on emissions.
Each may also have downsides---for example, locating conferences near their
existing ``centers of mass'' may discourage holding them in countries like
India and China where the community is still developing, while virtual PC
meetings and remote conference attendance may reduce opportunities for
informal connections and mentoring.  Getting beyond easy, short-term
solutions will require balancing competing concerns, demanding discussion,
compromise, careful analysis of data, and creative experimentation.

The problem of climate change is too large and too urgent to leave to world
leaders: organizations at every level must play their part in dealing with
their own contributions, raising awareness, and establishing new ways of
doing business in the lower-carbon future that is coming. For ACM, this
especially means developing strategies for reducing the climate impact of
conference-related travel. You can help in a number of ways: Get involved in
raising awareness about this issue in the organizations you are a part of.
For example, ask the organizers of the ACM conferences you attend what they
are doing about reducing or offsetting emissions from participants’ travel.
Analyze your organization’s carbon-emitting activities to inform a
discussion of how best to reduce its footprint.  Connect with others that
are engaging with these issues---in your workplace, in your other
communities, and within ACM (e.g., on the acm-climate mailing list [5]).

\bibliographystyle{unsrt}
\bibliography{local}

\section*{More References (to be bibtexified)}
\begin{itemize}
\item 
[3] Engaging with Climate Change: Possible Steps for SIGPLAN. Preliminary Report of the SIGPLAN Climate Committee.  Michael Hicks,​ ​Crista Lopes, and​ ​Benjamin C. Pierce. Version 1.2, June 2018. http://www.cis.upenn.edu/~bcpierce/papers/sigplan-climate-report.pdf
\item [4] Carbon Offsets: An Overview for Scientific Societies.  Richard Kim and ​Benjamin C. Pierce. Version 1.2 June 24, 2018.  http://www.cis.upenn.edu/~bcpierce/papers/carbon-offsets.pdf
\item[5] acm-climate mailing list, https://groups.google.com/forum/\#!forum/acm-climate
[6] OpenSimulator Community Conference, https://conference.opensimulator.org/2018/
\item [7] Why Aren’t There Electric Airplanes Yet? Venkat Viswanathan, Shashank Sripad, and William Leif Fredericks. November 2018. https://www.universal-sci.com/headlines/2018/11/30/-why-arent-there-electric-airplanes-yet
\item [8] CHI 2018, Remote Attendance, https://chi2018.acm.org/attending/remote-attendance/
\item [9] Neustaedter, C., Tang, J., Singhal, S., Pan, R., Heshmat, Y., and Forghani, A. (2017) Telepresence Attendance at the ACM CSCW 2016 Conference, Connections Lab Technical Report 2017-0419-01, Simon Fraser University.
\item [10] CO2 emissions (metric tons per capita). The World Bank Data Bank, using data from the Carbon Dioxide Information Analysis Center, Environmental Sciences Division, Oak Ridge National Laboratory, Tennessee, United States. Accessed December 2018. https://data.worldbank.org/indicator/EN.ATM.CO2E.PC
\item [11]  Supply and sustainability of carbon offsets and alternative fuels for international aviation, by
Rob Bailis, Derik Broekhoff, and Carrie M. Lee.  Stockholm Environment Institute, Working Paper 2016-03. https://www.sei.org/publications/carbon-offsets-aviation/			
\end{itemize}

\ifdraft
\newpage
\section*{Permissions for figures}

[1] Creative Commons license
[2] ``Reproduction of a limited number of figures or short excerpts of IPCC material is authorized free of charge and without formal written permission provided that the original source is properly acknowledged, with mention of the name of the report, the publisher and the numbering of the page(s) or the figure(s).''
[3] Authors

\section*{Things to do}

Final things to do:
Check where we’ve used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
Reformat (in latex?) to sharpen images

Discussions:
About Crista’s figure:
BCP: Would be great to supplement with a more refined analysis of registrations for each of these conferences; then we could add a "center of mass" dot for each of the conferences to the map. This would still not tell the whole story, since it's possible (likely) that the registration totals would have been different if the conferences had been located differently. But it would help tell the story.
Also about Crista’s figure:
BCP: We should quantify what the red and blue dots mean -- i.e., give readers an idea what overall difference a change of location would make to carbon footprint.
Mike: The dot colors mean per-capita footprint. Could we make their size represent the total footprint? Far-flung places might be high per-capita, but lower overall.
Crista: The order by total footprint is very similar to the normalized one. Here are the 5 top offenders by that criteria: ICFP-Japan, POPL-LA, POPL-SD, SPLASH-Vancouver, PLDI-Barcelona. There are some changes in the order of the conferences in the middle. I hesitated which plot to draw (each has positives and negatives) but decided this one would be less problematic. I don't think we have space for two, and overlaying the 2 data series will be confusing...
BTW, there's a LOT more that can be said about this data... But I think that will distract from the main message of this article. We need to chose whether this 2-page article is about the data analysis of our conferences or about a call for action. Maybe we can write a data analysis paper, upload it on arXiv, and link to it from this article?
For example: POPL-Paris was the absolute winner by all criteria: it had the highest number of participants, the second-lowest normalized carbon footprint, and was mid-point in total carbon footprint, even with all those participants.
This is just one tidbit; there are many more interesting facts in this data.
From Michael Coblenz:
What about research on energy-efficient computing? The impact of this research might be far more significant than that of mitigating our direct carbon emissions, and this is more in line with the core skills of the community.
Also, what about research on telepresence and virtual meetings? There’s obviously a lot of work in this area, but we may need more in order to (a) let us effectively trade off options, e.g. virtual PC meeting vs in-person; (b) improve the state of the art to make remote meetings more effective.
Consider also discussing the diversity implications on travel. Reducing travel expectations might make the field more attractive to women, who may face higher social expectations to stay home (and are more likely to be single parents). Surely there is a good article to cite here. I found https://journals.sagepub.com/doi/abs/10.1177/0950017006066999 but maybe there is something better. I’m looking…
Consider citing the new IEA report as another piece of evidence that action is needed
From Sophia: Does the article need an abstract to "set the reader's expectations"? I expected that there would be a list of actions in the end, or recommended resolutions. Instead, the article is more about what some communities do, and where individuals may be involved.  Perhaps say: "In this article we give a short overview of how different SCM communities tackle this problem, what thought are relevant to it, and conclude with suggestions on how to get involved". It breaks the flow, but it clarifies what the article is about.
Is citation 10 CO2 or CO2e?
Is there a ``more authoritative'' source for price of carbon offsets than [4]?
\fi

\end{document}

### 2018-12-10 main.tex
Checkpoint

@@
+\newif\ifdraft\drafttrue   % comments / todos for now
+\newif\iflater\laterfalse  % comments / todos for final version
+
 \documentclass[12pt]{article}
@@
 \usepackage{hyperref}
+\usepackage{multicol}
 
-\newif\ifdraft\drafttrue
-
-\newcommand{\bcp}[1]{{\bf #1}}
+\newcommand{\bcp}[1]{\ifdraft{\bf #1}\fi}
 
@@
 
+% \begin{multicols}{3}
+
 ACM’s vibrant conference culture has a dark side: it contributes
@@
 
-\begin{figure}[tb]
+\begin{figure}[t]
 \centering
 \includegraphics[width=4in]{wmo_stripes.png}
-\caption{Annual global temperatures, 1850 to 2018 (range from blue to red:
-  1.35 degrees C) \cite{WarmingStripes}}
+\caption{Variation in annual global temperature, 1850 to 2018 (range from
+  blue to red: $1.35^\circ$C) \cite{WarmingStripes}}
 \label{fig:warming}
@@
 
-\begin{figure}[tb]
+\begin{figure}[t]
 \centering
@@
   decline by at least 40 percent by 2030 and reach net zero by 2050, if we
-  are to hold warming to 1.5 degrees.  ``No documented historic precedent''
-  for the transformation of the world economy needed to achieve this.  }
-\label{fig:warming}
+  are to hold warming to 1.5 degrees.  There is ``no documented historic
+  precedent'' for the transformation of the world economy needed to achieve
+  this.  }
+\label{fig:IPCC-cover}
 \end{figure}
 
-\begin{figure}[tb]
+\begin{figure}[t]
 \centering
@@
 \caption{Carbon footprint per participant for travel to recent SIGPLAN
-  conferences\bcp{Get better file from Crista}} 
+  conferences\iflater\bcp{Get better file from the other paper, or get
+    better version of this one from Crista.}\fi} 
 \label{fig:confs}
@@
 emissions by 2030 is needed to avoid irreversible damage
-(Figure~\ref{IPCC-cover}). Achieving this will require urgent and sustained
+(Figure~\ref{fig:IPCC-cover}). Achieving this will require urgent and sustained
 commitment at all levels of society, including non-governmental communities
@@
 especially means developing strategies for reducing the climate impact of
-conference-related travel. You can help in a number of ways: Get involved in
-raising awareness about this issue in the organizations you are a part of.
-For example, ask the organizers of the ACM conferences you attend what they
-are doing about reducing or offsetting emissions from participants’ travel.
-Analyze your organization’s carbon-emitting activities to inform a
-discussion of how best to reduce its footprint.  Connect with others that
-are engaging with these issues---in your workplace, in your other
-communities, and within ACM (e.g., on the acm-climate mailing list [5]).
+conference-related travel. You can help in a number of ways: 
+\begin{itemize}
+\item {\bf Get involved} in raising awareness about this issue in the
+organizations you are a part of.  For example, ask the organizers of the ACM
+conferences you attend what they are doing about reducing or offsetting
+emissions from participants’ travel.
+\item {\bf Analyze} your organization’s carbon-emitting activities to inform a
+discussion of how best to reduce its footprint.
+\item {\bf Connect} with others that are engaging with these issues---in
+your workplace, in your other communities, and within ACM (e.g., on the
+acm-climate mailing list [5]).
+\end{itemize}
 
@@
 
-Final things to do:
+\paragraph*{Final things to do:}
+\begin{itemize}
+\item 
 Check where we’ve used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
-Reformat (in latex?) to sharpen images
+\end{itemize}
 
-Discussions:
+\paragraph*{Discussions:}
+\begin{itemize}
+\item 
 About Crista’s figure:
@@
 But it would help tell the story.
+\item 
 Also about Crista’s figure:
+\begin{itemize}
+\item 
 BCP: We should quantify what the red and blue dots mean -- i.e., give readers an idea what overall difference a change of location would make to carbon footprint.
-Mike: The dot colors mean per-capita footprint. Could we make their size represent the total footprint? Far-flung places might be high per-capita, but lower overall.
-Crista: The order by total footprint is very similar to the normalized one.
+\item Mike: The dot colors mean per-capita footprint. Could we make their size represent the total footprint? Far-flung places might be high per-capita, but lower overall.
+\item Crista: The order by total footprint is very similar to the normalized one.
 Here are the 5 top offenders by that criteria: ICFP-Japan, POPL-LA, POPL-SD, SPLASH-Vancouver, PLDI-Barcelona.
@@
 I don't think we have space for two, and overlaying the 2 data series will be confusing...
-BTW, there's a LOT more that can be said about this data...
+\item BTW, there's a LOT more that can be said about this data...
 But I think that will distract from the main message of this article.
@@
 Maybe we can write a data analysis paper, upload it on arXiv, and link to it from this article?
+\item
 For example: POPL-Paris was the absolute winner by all criteria: it had the highest number of participants, the second-lowest normalized carbon footprint, and was mid-point in total carbon footprint, even with all those participants.
 This is just one tidbit; there are many more interesting facts in this data.
+\end{itemize}
+\item 
 From Michael Coblenz:
@@
 I’m looking…
+\item 
 Consider citing the new IEA report as another piece of evidence that action is needed
+\item 
 From Sophia: Does the article need an abstract to "set the reader's expectations"?
@@
 It breaks the flow, but it clarifies what the article is about.
+\item 
 Is citation 10 CO2 or CO2e?
+\item 
 Is there a ``more authoritative'' source for price of carbon offsets than [4]?
+\end{itemize}
 \fi
 
+% \end{multicols}
 \end{document}

### 2018-12-11 main.tex
Finished first draft

@@
 \newcommand{\bcp}[1]{\ifdraft{\bf #1}\fi}
+\newcommand{\COtwoE}{CO$_2$e}
 
@@
 Philadelphia to Paris emits the equivalent of about 2 tons of carbon dioxide
-(CO2e) per passenger. Compare this against the fact that the average US
+(\COtwoE) per passenger. Compare this against the fact that the average US
 resident emits about 16.5 tons per year, while the average European emits
-about 7 tons [10].  Moreover, these emissions have no near-term
+about 7 tons~\cite{emissions}.  Moreover, these emissions have no near-term
 technological fix because jet fuel is difficult to replace with renewable
-energy sources [7]. So reducing the CO2e footprint of conferences is going
+energy sources~\cite{elec-air}. So reducing the \COtwoE{} footprint of conferences is going
 to require a reduction in the footprint of the attendees’ air travel.
@@
 formed in 2016 to consider possible strategies for mitigating the emissions
-of our conferences [3]. Among the ones we consider most attractive are:
-making it easier to use carbon offsets to mitigate aviation emissions for
-conferences, choosing conference locations with an eye to minimizing
-emissions, co-locating or merging conferences, virtualizing program
-committees, and supporting remote participation.
+of our conferences \cite{ClimateCommitteReport}. Among the ones we find most
+attractive are: making it easier to use carbon offsets to mitigate aviation
+emissions for conferences, choosing conference locations with an eye to
+minimizing emissions, co-locating or merging conferences, virtualizing
+program committees, and supporting remote participation.
 
 Buying carbon offsets to reduce the impact of air travel is an ``easy win''
-[4].  The idea of carbon offsets is to pay somebody to plant trees, capture
+\cite{CarbonOFfsetReport}.  The idea of carbon offsets is to pay somebody to plant trees, capture
 methane (from landfills or cattle herds, for example), buy cookstoves to
@@
 emission. At the moment, the average air travel to an international
-conference can be offset by paying around \$30 [4].  ACM could streamline
+conference can be offset by paying around \$30 \cite{CarbonOFfsetReport}.
+ACM could streamline 
 the purchase of such offsets and make offsetting the default by folding them
@@
 Unfortunately, carbon offsetting is only a short-term mitigation, since the
-potential supply of offsets is far less than the potential demand [11]. We
+potential supply of offsets is far less than the potential demand~\cite{SEI-Report}. We
 must also begin significantly reducing emissions, which means reducing the
@@
 effects of location choice are less obvious. For example,
-Figure~\ref{fig:SIGPLAN-confs} plots the 
+Figure~\ref{fig:confs} plots the 
 locations of the four major SIGPLAN conferences over the past five years,
@@
 while SIGACCESS and SIGCHI have experimented with telepresence robots at
-their main conferences [8,9]. Journal-first initiatives, which make
-conference attendance optional, could also help.  Looking further ahead,
-some communities are already experimenting with fully virtual conferences
-[6].  ACM members can contribute to the development of technologies for
-connecting people effectively at a distance.
+their main conferences~\cite{CHI-remote,Neustaedter}. Journal-first
+initiatives, which make conference attendance optional, could also help.
+Looking further ahead, some communities are already experimenting with fully
+virtual conferences~\cite{OS}.  ACM members can contribute to the
+development of technologies for connecting people effectively at a distance.
 
@@
 your workplace, in your other communities, and within ACM (e.g., on the
-acm-climate mailing list [5]).
+acm-climate mailing list~\cite{acm-climate}).
 \end{itemize}
@@
 
-\section*{More References (to be bibtexified)}
-\begin{itemize}
-\item 
-[3] Engaging with Climate Change: Possible Steps for SIGPLAN.
-Preliminary Report of the SIGPLAN Climate Committee.
-Michael Hicks,​ ​Crista Lopes, and​ ​Benjamin C.
-Pierce.
-Version 1.2, June 2018.
-http://www.cis.upenn.edu/~bcpierce/papers/sigplan-climate-report.pdf
-\item [4] Carbon Offsets: An Overview for Scientific Societies.  Richard Kim and ​Benjamin C. Pierce. Version 1.2 June 24, 2018.  http://www.cis.upenn.edu/~bcpierce/papers/carbon-offsets.pdf
-\item[5] acm-climate mailing list, https://groups.google.com/forum/\#!forum/acm-climate
-[6] OpenSimulator Community Conference, https://conference.opensimulator.org/2018/
-\item [7] Why Aren’t There Electric Airplanes Yet?
-Venkat Viswanathan, Shashank Sripad, and William Leif Fredericks.
-November 2018.
-https://www.universal-sci.com/headlines/2018/11/30/-why-arent-there-electric-airplanes-yet
-\item [8] CHI 2018, Remote Attendance, https://chi2018.acm.org/attending/remote-attendance/
-\item [9] Neustaedter, C., Tang, J., Singhal, S., Pan, R., Heshmat, Y., and Forghani, A.
-(2017) Telepresence Attendance at the ACM CSCW 2016 Conference, Connections Lab Technical Report 2017-0419-01, Simon Fraser University.
-\item [10] CO2 emissions (metric tons per capita).
-The World Bank Data Bank, using data from the Carbon Dioxide Information Analysis Center, Environmental Sciences Division, Oak Ridge National Laboratory, Tennessee, United States.
-Accessed December 2018.
-https://data.worldbank.org/indicator/EN.ATM.CO2E.PC
-\item [11]  Supply and sustainability of carbon offsets and alternative fuels for international aviation, by
-Rob Bailis, Derik Broekhoff, and Carrie M. Lee.  Stockholm Environment Institute, Working Paper 2016-03. https://www.sei.org/publications/carbon-offsets-aviation/			
-\end{itemize}
-
-\ifdraft
+\iflater
 \newpage
@@
 [1] Creative Commons license
+
 [2] ``Reproduction of a limited number of figures or short excerpts of IPCC material is authorized free of charge and without formal written permission provided that the original source is properly acknowledged, with mention of the name of the report, the publisher and the numbering of the page(s) or the figure(s).''
+
 [3] Authors
 
-\section*{Things to do}
-
-\paragraph*{Final things to do:}
-\begin{itemize}
-\item 
-Check where we’ve used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
-\end{itemize}
-
-\paragraph*{Discussions:}
+\section*{More Discussions:}
 \begin{itemize}
@@
 \item 
-Is citation 10 CO2 or CO2e?
+Is the citation for emissions by region CO2 or CO2e?
 \item 
-Is there a ``more authoritative'' source for price of carbon offsets than [4]?
+Is there a ``more authoritative'' source for price of carbon offsets than
+our carbon report?
 \end{itemize}
+\item \item 
+Check where we’ve used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
 \fi

### 2018-12-11 main.tex
Notes

@@
 \begin{itemize}
+\item Old comments from Andrew Chien:
+\begin{quote}
+So, should we really be so conference intensive? (we seem to be far more so than other disciplines)
+
+If we are going to have SIG conferences, as a staple of CS, do they actually need to be face to face?  (could
+make the f2f ones less frequent)  What would the scientific impact be?
+
+We seem to have wired into the economics of our conferences that "bigger attendance is better".  this supports
+the fixed costs, makes the hotels happy, etc.  Can we make conferences where smaller in-person attendance, but
+larger total engagement (whatever that means... perhaps virtual or after the fact interaction) is maximized?
+
+What does a zero-carbon professional society look like?  (if you peek at my zccloud.cs.uchicago.edu page, you'll
+see that I don't mean offset, but rather actual zero-incremental carbon emissions)
+
+If you've already thought about all of these, then that's great.  If not, please do take them as a set of potential
+challenges to consider (all voluntary... as I intend them only as suggestions).
+\end{quote}
 \item 

### 2018-12-11 main.tex
Added couple sentences of vision

@@
 
-\newcommand{\bcp}[1]{\ifdraft{\bf #1}\fi}
+\newcommand{\bcp}[1]{\ifdraft{\bf [bcp: #1]}\fi}
 \newcommand{\COtwoE}{CO$_2$e}
@@
 
-Each of these strategies will have some beneficial effect on emissions.
-Each may also have downsides---for example, locating conferences near their
-existing ``centers of mass'' may discourage holding them in countries like
-India and China where the community is still developing, while virtual PC
-meetings and remote conference attendance may reduce opportunities for
-informal connections and mentoring.  Getting beyond easy, short-term
-solutions will require balancing competing concerns, demanding discussion,
-compromise, careful analysis of data, and creative experimentation.
+Each of these strategies will have some beneficial effect on emissions.  Of
+course, each may also have downsides---for example, locating conferences
+near their existing ``centers of mass'' may discourage holding them in
+countries like India and China where the community is still developing,
+while virtual PC meetings and remote conference attendance may reduce
+opportunities for informal connections and mentoring.  Getting beyond easy,
+short-term solutions will require balancing competing concerns, demanding
+discussion, compromise, careful analysis of data, and creative
+experimentation.  
+%
+\bcp{Added remainder of paragraph.  I don't think the two sentences work
+together very well, but I can't decide which to keep!}
+It will also require questioning some fundamental assumptions about how
+research gets done---e.g., that more and bigger conferences are better, and
+that personal success is (positively) correlated with time spent in
+airports.
+%
+And to thrive in the long term, we need to begin imagining what a
+zero-carbon professional society might look like.
 

### 2018-12-11 main.tex
Nits.

@@
 \newcommand{\bcp}[1]{\ifdraft{\bf [bcp: #1]}\fi}
+\newcommand{\mwh}[1]{\ifdraft{\bf [mwh: #1]}\fi}
 \newcommand{\COtwoE}{CO$_2$e}
@@
 particularly from air travel. For example, a single round-trip flight from
-Philadelphia to Paris emits the equivalent of about 2 tons of carbon dioxide
-(\COtwoE) per passenger. Compare this against the fact that the average US
+Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon dioxide
+(\COtwoE) per passenger.\mwh{Cite the carboncalculator.org site I used
+  to calculate this?} Compare this against the fact that the average US
 resident emits about 16.5 tons per year, while the average European emits
@@
 sites. Switching to virtual program committee meetings for conferences is
-another obvious step.  Indeed, this switch is already happening: we found in
+another obvious step. \mwh{data about surprising connection to
+  conference's overall footprint?}  Indeed, this switch is already happening: we found in
 a recent poll of ACM SIG chairs that, across ACM, a majority of the SIGs’
@@
 Looking further ahead, some communities are already experimenting with fully
-virtual conferences~\cite{OS}.  ACM members can contribute to the
+virtual \mwh{reality?} conferences~\cite{OS}.  ACM members can contribute to the
 development of technologies for connecting people effectively at a distance.
 
-Each of these strategies will have some beneficial effect on emissions.  Of
-course, each may also have downsides---for example, locating conferences
+Each of these strategies will have some beneficial effect on emissions.  
+Each may also have downsides. For example, locating conferences
 near their existing ``centers of mass'' may discourage holding them in
@@
 And to thrive in the long term, we need to begin imagining what a
-zero-carbon professional society might look like.
+zero-carbon professional society will look like.
 

### 2018-12-11 main.tex
Small polishing and responses to Mike

@@
 particularly from air travel. For example, a single round-trip flight from
-Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon dioxide
-(\COtwoE) per passenger.\mwh{Cite the carboncalculator.org site I used
-  to calculate this?} Compare this against the fact that the average US
-resident emits about 16.5 tons per year, while the average European emits
-about 7 tons~\cite{emissions}.  Moreover, these emissions have no near-term
+Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon
+dioxide (\COtwoE) per passenger~\cite{carboncalculator}.\iflater\bcp{Trim
+  citation?}\fi{} Compare this against the fact that the average US resident
+emits about 16.5 tons per year, while the average European emits about 7
+tons~\cite{emissions}.  Moreover, these emissions have no near-term
 technological fix because jet fuel is difficult to replace with renewable
-energy sources~\cite{elec-air}. So reducing the \COtwoE{} footprint of conferences is going
-to require a reduction in the footprint of the attendees’ air travel.
+energy sources~\cite{elec-air}. So reducing the \COtwoE{} footprint of
+conferences is going to require a reduction in the footprint of the
+attendees’ air travel.
 
@@
 sites. Switching to virtual program committee meetings for conferences is
-another obvious step. \mwh{data about surprising connection to
-  conference's overall footprint?}  Indeed, this switch is already happening: we found in
-a recent poll of ACM SIG chairs that, across ACM, a majority of the SIGs’
-main conferences now use online PC meetings rather than in-person ones.
+another obvious step. \mwh{data about surprising connection to conference's
+  overall footprint?}\bcp{Or leave it for next round of revisions when we
+  may have done more/better analysis?}  Indeed, this switch is already
+happening: we found in a recent poll of ACM SIG chairs that, across ACM, a
+majority of the SIGs’ main conferences now use online PC meetings rather
+than in-person ones.
 
@@
 Looking further ahead, some communities are already experimenting with fully
-virtual \mwh{reality?} conferences~\cite{OS}.  ACM members can contribute to the
+virtual 
+% \mwh{reality?}
+% BCP: Seems redundant with "fully"
+conferences~\cite{OS}.  ACM members can contribute to the
 development of technologies for connecting people effectively at a distance.
@@
 %
-\bcp{Added remainder of paragraph.  I don't think the two sentences work
-together very well, but I can't decide which to keep!}
+\bcp{Added remainder of paragraph.  I'm not sure the two sentences work
+  together very well, but I can't decide which to keep!  (Mike likes them
+  both.)}%
 It will also require questioning some fundamental assumptions about how
@@
 %
-And to thrive in the long term, we need to begin imagining what a
-zero-carbon professional society will look like.
+To thrive over the long haul, we need to begin imagining what a zero-carbon
+professional society will look like.
 

### 2018-12-12 main.tex
Polish

@@
 
-ACM’s vibrant conference culture has a dark side: it contributes
+ACM's vibrant conference culture has a dark side: it contributes
 significantly to climate change.
@@
 
-There is a broad scientific consensus that the earth is warming due to human
+A broad scientific consensus warns that the earth is warming due to human
 emissions of carbon dioxide and other greenhouse gases (Figure
 \ref{fig:warming}). This is not a problem we can leave to future
-generations: we are up against it now. Indeed, the UN’s Intergovernmental
-Panel on Climate Change (IPCC) recently reported that a 40\% decrease in
-emissions by 2030 is needed to avoid irreversible damage
-(Figure~\ref{fig:IPCC-cover}). Achieving this will require urgent and sustained
-commitment at all levels of society, including non-governmental communities
-such as universities, companies, and scientific societies like ACM.
+generations: we are up against it now. Indeed, the UN's Intergovernmental
+Panel on Climate Change (IPCC) recently recommended a 40\% decrease in
+emissions by 2030 to avoid irreversible damage
+(Figure~\ref{fig:IPCC-cover}). Achieving reductions on this scale will
+require urgent and sustained commitment at all levels of society, including
+non-governmental communities such as universities, companies, and scientific
+societies like ACM.
 
-Why ACM? Because one of ACM’s main activities is organizing international
-conferences, and these produce significant emissions of greenhouse gases,
+Why ACM? Because one of ACM's main activities is organizing international
+conferences, which produce significant emissions of greenhouse gases,
 particularly from air travel. For example, a single round-trip flight from
 Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon
-dioxide (\COtwoE) per passenger~\cite{carboncalculator}.\iflater\bcp{Trim
-  citation?}\fi{} Compare this against the fact that the average US resident
-emits about 16.5 tons per year, while the average European emits about 7
-tons~\cite{emissions}.  Moreover, these emissions have no near-term
-technological fix because jet fuel is difficult to replace with renewable
-energy sources~\cite{elec-air}. So reducing the \COtwoE{} footprint of
-conferences is going to require a reduction in the footprint of the
-attendees’ air travel.
+dioxide (\COtwoE) per passenger~\cite{carboncalculator}\iflater\bcp{Trim
+  citation?}\fi---a substantial fraction of the {\em total} yearly emissions
+for an average resident of the US (16.5 tons) or Europe (7
+tons)~\cite{emissions}.  Moreover, these emissions have no near-term
+technological fix because jet fuel is difficult to replace with
+renewable energy sources~\cite{elec-air}. 
+\iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
+of conferences will require a reduction in the footprint of the attendees'
+air travel.}\fi
 
 We are the members of an {\em ad hoc} Committee on Climate Change within
-SIGPLAN (ACM’s Special Interest Group on Programming Languages) that was
+ACM's Special Interest Group on Programming Languages (SIGPLAN) that was
 formed in 2016 to consider possible strategies for mitigating the emissions
-of our conferences \cite{ClimateCommitteReport}. Among the ones we find most
-attractive are: making it easier to use carbon offsets to mitigate aviation
+of our conferences \cite{ClimateCommitteReport}. Among the ones we find 
+attractive are making it easier to use carbon offsets to mitigate aviation
 emissions for conferences, choosing conference locations with an eye to
@@
 
-Buying carbon offsets to reduce the impact of air travel is an ``easy win''
-\cite{CarbonOFfsetReport}.  The idea of carbon offsets is to pay somebody to plant trees, capture
-methane (from landfills or cattle herds, for example), buy cookstoves to
-replace inefficient open-fire cooking in poor communities, or carry out
-other actions that remove greenhouse gases from the air or prevent their
-emission. At the moment, the average air travel to an international
+Buying carbon offsets to reduce the impact of air travel
+\cite{CarbonOFfsetReport} is an ``easy win'' that conferences should
+implement now.  The idea of carbon offsets is to pay somebody to plant
+trees, capture methane (from landfills or cattle herds, for example), buy
+cookstoves to replace inefficient open-fire cooking in poor communities, or
+carry out other actions that remove greenhouse gases from the air or prevent
+their emission. At the moment, the average air travel to an international
 conference can be offset by paying around \$30 \cite{CarbonOFfsetReport}.
-ACM could streamline 
-the purchase of such offsets and make offsetting the default by folding them
-into its conference registration fees.
+ACM could streamline the purchase of such offsets and make offsetting the
+default by folding them into its conference registration fees.
 
-Unfortunately, carbon offsetting is only a short-term mitigation, since the
-potential supply of offsets is far less than the potential demand~\cite{SEI-Report}. We
-must also begin significantly reducing emissions, which means reducing the
-distance traveled by the sum of all conference participants.
+Sadly, carbon offsetting is only a short-term mitigation, since the
+potential supply of offsets is far less than the potential
+demand~\cite{SEI-Report}. We must also begin significantly reducing
+emissions, which means reducing the distance traveled by the sum of all
+conference participants.
 
-One approach to doing so is to choose conference locations that are close to
-the center of mass of their potential participants.  For example, SIGGRAPH
-(the Special Interest Group on Graphics) locates in Los Angeles every other
-year---an easy call as this is near an international airport and a major hub
+One way to do this is to choose conference locations that are close to
+the center of mass of their likely participants.  For example, the
+SIGGRAPH conference on computer graphics
+locates in Los Angeles every other
+year---an easy call, as this is near an international airport and a major hub
 of the worldwide entertainment industry. For other research areas, the
@@
 colored according to the estimated average carbon footprint per
-participant. The figure shows that there were significant differences across
-conferences and across locations---indeed, it is a factor of two between the
+participant. The figure shows significant differences across
+conferences and across locations---indeed, a factor of two difference
+between the 
 lowest and highest. Why is this? One reason might be that the locus of a
-conference’s population is closer to one location than another, and changing
+conference's population is closer to one location than another, and changing
 locations reduces the total cost. Another might be the presence of a local
 population that will only attend if the conference is very close to
-them. Our preliminary analysis suggests it may be both. The larger question
-is how to use this kind of information to choose the best locations over
+them. Our preliminary analysis suggests it may be both (and that deeper
+analysis is needed to understand the situation better). The larger question
+is how to use this kind of information to choose the best locations, over
 time, for a diffuse population.
@@
 are all ``the'' conference for their areas.  More communities---for example,
-the programming languages and security communities, each of which holds four
+the programming languages and security communities, each with four
 main conferences every year---could consider following suit.  A different
@@
 sites. Switching to virtual program committee meetings for conferences is
-another obvious step. \mwh{data about surprising connection to conference's
+another obvious step.\iflater \mwh{data about surprising connection to conference's
   overall footprint?}\bcp{Or leave it for next round of revisions when we
-  may have done more/better analysis?}  Indeed, this switch is already
+  may have done more/better analysis?}\fi{}  Indeed, this switch is already
 happening: we found in a recent poll of ACM SIG chairs that, across ACM, a
-majority of the SIGs’ main conferences now use online PC meetings rather
+majority of the SIGs' main conferences now use online PC meetings rather
 than in-person ones.
 
-We can attend some conferences online, saving both emissions and travel
-time.  Indeed, SIGPLAN already livestreams all its four main conferences,
-while SIGACCESS and SIGCHI have experimented with telepresence robots at
-their main conferences~\cite{CHI-remote,Neustaedter}. Journal-first
-initiatives, which make conference attendance optional, could also help.
-Looking further ahead, some communities are already experimenting with fully
-virtual 
+Another approach is to allow some conference attendees to participate
+remotely, saving both emissions and travel time.  Indeed, SIGPLAN already
+livestreams all its four main conferences, while SIGACCESS and SIGCHI have
+experimented with telepresence robots at their main
+conferences~\cite{CHI-remote,Neustaedter}. \iflater\bcp{Next sentence breaks
+the flow a bit.  Cut it?}\fi{}Journal-first initiatives, which
+make conference attendance optional, can also help.  Looking further
+ahead, some communities are already experimenting with fully virtual
 % \mwh{reality?}
@@
 conferences you attend what they are doing about reducing or offsetting
-emissions from participants’ travel.
-\item {\bf Analyze} your organization’s carbon-emitting activities to inform a
+emissions from participants' travel.
+\item {\bf Analyze} your organization's carbon-emitting activities to inform a
 discussion of how best to reduce its footprint.
@@
 \item 
-About Crista’s figure:
+About Crista's figure:
 BCP: Would be great to supplement with a more refined analysis of registrations for each of these conferences; then we could add a "center of mass" dot for each of the conferences to the map.
@@
 \item 
-Also about Crista’s figure:
+Also about Crista's figure:
 \begin{itemize}
@@
 Also, what about research on telepresence and virtual meetings?
-There’s obviously a lot of work in this area, but we may need more in order to (a) let us effectively trade off options, e.g.
+There's obviously a lot of work in this area, but we may need more in order to (a) let us effectively trade off options, e.g.
 virtual PC meeting vs in-person; (b) improve the state of the art to make remote meetings more effective.
@@
 I found https://journals.sagepub.com/doi/abs/10.1177/0950017006066999 but maybe there is something better.
-I’m looking…
+I'm looking…
 \item 
@@
 \item \item 
-Check where we’ve used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
+Check where we've used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
 \fi

### 2018-12-12 main.tex
Submission version

@@
 
-Sadly, carbon offsetting is only a short-term mitigation, since the
+However, carbon offsetting is only a short-term mitigation, since the
 potential supply of offsets is far less than the potential
@@
 locates in Los Angeles every other
-year---an easy call, as this is near an international airport and a major hub
+year---an easy call, as this is near both an international airport and a major hub
 of the worldwide entertainment industry. For other research areas, the
@@
 
-Another approach is to allow some conference attendees to participate
+Another approach is allowing some conference attendees to participate
 remotely, saving both emissions and travel time.  Indeed, SIGPLAN already
@@
 Each may also have downsides. For example, locating conferences
-near their existing ``centers of mass'' may discourage holding them in
+near their existing centers of mass may discourage holding them in
 countries like India and China where the community is still developing,
@@
 opportunities for informal connections and mentoring.  Getting beyond easy,
-short-term solutions will require balancing competing concerns, demanding
+short-term solutions and fat-trimming exercises requires balancing competing
+concerns; this will demand significant  
 discussion, compromise, careful analysis of data, and creative
@@
 %
-\bcp{Added remainder of paragraph.  I'm not sure the two sentences work
+\iflater\bcp{Added remainder of paragraph.  I'm not sure the two sentences work
   together very well, but I can't decide which to keep!  (Mike likes them
-  both.)}%
-It will also require questioning some fundamental assumptions about how
+  both.)}\fi%
+It will also involve questioning some fundamental assumptions about how
 research gets done---e.g., that more and bigger conferences are better, and
@@
 The problem of climate change is too large and too urgent to leave to world
-leaders: organizations at every level must play their part in dealing with
-their own contributions, raising awareness, and establishing new ways of
+leaders: organizations at every level must play their part by addressing
+their own contributions, raising awareness among their members, and
+establishing new ways of 
 doing business in the lower-carbon future that is coming. For ACM, this
 especially means developing strategies for reducing the climate impact of
-conference-related travel. You can help in a number of ways: 
+conference-related travel.
+%
+You can help accelerate this process in a number of ways: 
 \begin{itemize}
@@
 your workplace, in your other communities, and within ACM (e.g., on the
-acm-climate mailing list~\cite{acm-climate}).
+{\tt acm-climate} mailing list~\cite{acm-climate}).
 \end{itemize}

### 2018-12-26 main.tex
todo

@@
 for an average resident of the US (16.5 tons) or Europe (7
-tons)~\cite{emissions}.  Moreover, these emissions have no near-term
+tons)~\cite{emissions}.  \iflater\bcp{For comparison, ...  (put in some
+  figures about the footprints of a few other things that people might think
+  of reducing---driving, eating meat, upgrading their fridge...)}\fi{}
+Moreover, these emissions have no near-term 
 technological fix because jet fuel is difficult to replace with

### 2019-05-14 main.tex
Comments

@@
 \begin{itemize}
+\item New comments from Judith Bishop:
+
+1. Figure 3 is unclear. I did not understand it. Maybe it is missing a key, or the circles are too small. A rework would be very beneficial.
+
+2. All articles benefit from headings. It reveals the structure of the paper at a glance and assists in avoiding repetition. Think about some headings. Have a look at some recent Viewpoints papers.
+
+3.
+The title is rather opaque.
+The article is mostly about conferences, so maybe that should be in the title.
+Viewpoints likes catchy titles, so something more personal such as "Your conference trip and climate change" might get more readers.
+
+4.
+The article does not mention the journal publication alternative to conferences.
+We all know that computer scientists fight fiercely for the status of their conference papers on their CVs.
+Maybe climate change will be the tipping point that will cause them to become more like their colleagues in other disciplines and publish in journals, with just one conference a year, to meet colleagues and catch up.
+From my side, I do think this aspect should be mentioned.
+
 \item Old comments from Andrew Chien:

### 2019-05-14 rewrite.tex
Start at a rewrite.

--- 2019-05-14 main.tex
@@
 \usepackage{fullpage}
+\usepackage{times}
 \usepackage{graphicx}
@@
 ACM's vibrant conference culture has a dark side: it contributes
-significantly to climate change.
+significantly to climate change. What should we do about it?
+
+\bigskip{}
+
+Experts say that human civilization must reduce CO2 emissions by 40\%
+by 2030. To do that, we all need to tighten our CO2-emitting belts. For
+academic computer scientists, from what activity should the CO2 reduction come?
+Look no further than our professional conferences. Air
+travel is CO2 heavy, on a per-capita basis. A few trans-oceanic trips
+is XXX CO2. Wow!
+
+As responsible scientists wanting to do our part to sustain our
+environment, should we eliminate (most) conferences, and thereby the
+travel to organize and attend them? Doing so could impose a high cost
+on scientific progress. Science is a fundamentally social process. The
+conference system leverages that, accelerating research through easy
+interaction, direct dissemination of results, and network building to
+further new work. It works well enough that conferences are
+proliferating and attendance is on the rise. So the challenge is: How
+do we weigh the benefits to science of conferences against their
+attendant CO2 (and other) costs? This seems like an extremely
+difficult question to answer.
+
+But not so fast: such difficult questions are answered all the time
+through the magic of markets. By mapping choices to currency amounts,
+markets facilitate individual valuation of costs and benefits across
+vastly different activities, like the cost of flying and staying in a
+hotel vs. the benefit of attending a scientific meeting. If the former
+was too high, we would not do the latter. To facilitate answers to
+difficult questions about CO2 costs, many are advocating CO2-centered
+markets, by placing a dollar cost on CO2 emissions. California has
+already done this. In a world in which CO2 had a cost, air travel
+would be more expensive because flyers would have to bear its CO2
+cost, too. The result would place downward pressure on conferences
+that were CO2-inefficient for their attendees.
+
+There’s no global market for CO2 just yet, so does that mean we should
+do nothing until there is one? Not necessarily. We can start to impose
+the discipline on ourselves right now.
+
+First, we can recognize there is an emerging price for CO2, which is
+the price of CO2 offsets. [Explain what they are.]
+
+With a CO2 cost in hand, we can establish an acceptable baseline CO2
+cost for an event. This baseline could be established by analyzing
+[...] Planners can track this cost according to the CO2 footprint of
+their attendees. [...]
+
+\hrule
+\clearpage
+
+\texdtbf{OLD STUFF HERE}
 

### 2019-05-25 main.tex
Progress on a rewrite

@@
 
-A broad scientific consensus warns that the earth is warming due to human
+A broad scientific consensus warns that the earth is warming, due to human
 emissions of carbon dioxide and other greenhouse gases (Figure
@@
 air travel.}\fi
+
+Obviously, conferences also create value.  Science is a fundamentally social
+process, and the conference system accelerates research through
+opportunities for high-bandwidth interaction, direct dissemination of
+results, network building, and serendipitous cross-fertilization.
+
+This creates a dilemma: How do we weigh the environmental harm caused by
+conference travel against the social benefits of the scientific progress
+that comes from conferences?  This seems an extremely difficult question to
+answer: the harm and the compensating benefits cannot be directly
+compared---they are apples and oranges.
+%
+Fortunately, we have a well-established mechanism for comparing apples and
+oranges---indeed, for comparing almost anything: the magic of markets.  By
+mapping choices to currency amounts, markets facilitate individual valuation
+of costs and benefits across vastly different activities.\bcp{Give a simple,
+  concrete example.} For this reason, many policy experts advocate some form
+of carbon pricing, which places a monetary cost on greenhouse gas emissions,
+giving a market signal that changes aggregate behavior.  Exactly what the
+cost should be is a challenging question, but it has a clear ``success
+criterion'': The cost should be set so that the world arrives at zero
+emissions within the timeframe that climate scientists tell us is safe.
+
+% (Experts seem to agree that the current price of around \$15 / ton is too
+% low.)
+
+Many countries and states (e.g., California) have already adopted carbon
+pricing policies; more will follow.  The upshot is that, sooner or later,
+carbon is going to be expensive.
+\bcp{Yes, this may not happen.  But if it does not, we are probably
+  screwed.  So might as well behave as if we knew it was going to happen.}
+%
+In the meantime, knowing that this is coming, what should we rationally (and
+ethically) do now?
+
+First, we should immediately start planning for how ACM, and computer-science
+research more broadly, are going to continue to thrive in a world where air
+travel is significantly more expensive.  Indeed, this may well involve
+rethinking the whole basis of how computer science is done. 
+\bcp{Add some meat to that.}
+
+Second, we can begin gathering and analyzing data, to better understand both
+the current carbon impact of our conferences and the likely effects (both
+environmental and scientific) of potential changes in the way they are
+organized.
+\bcp{Put in the most interesting discussion and perhaps show an interesting
+  diagram.}
+
+\bcp{Polish and expand with reworked old text and new points.}
+Finally, we can start imposing a cost on ourselves by using carbon
+offsets.  (People counter-argue that this "lets people off the hook" and
+encourages inaction, but we believe it is consistent with encouraging action
+-- again, if offsets are priced right.)
+
+The climate crisis is too large and too urgent to just delegate to world
+leaders: organizations at every social scale must play their part by addressing
+their own contributions, raising awareness among their members, and
+establishing new ways of 
+doing business in the lower-carbon future that is coming. For ACM, this
+especially means developing strategies for reducing the climate impact of
+conference-related travel.
+%
+You can help accelerate this process in a number of ways: 
+\begin{itemize}
+\item {\bf Get involved} in raising awareness about this issue in the
+organizations you are a part of.  \bcp{Does this smack of incrementalism?
+  Maybe delete it.
+  ``For example, ask the organizers of the conferences you attend what they
+  are doing about reducing or offsetting emissions from participants'
+  travel.''}
+\item {\bf Connect} with others that are engaging with these issues---in
+your workplace, in your other communities, and within ACM (e.g., on the
+{\tt acm-climate} mailing list~\cite{acm-climate}).
+\item {\bf Analyze} your organization's carbon-emitting activities to inform
+a discussion of how reduce its footprint and prepare for the era of
+expensive carbon.
+\end{itemize}
+
+\bcp{Things still missing:
+  \begin{itemize}
+  \item citations for the claims we make
+  \item an excuse for citing the Climate Committee report and perhaps
+  mentioning other SIGPLAN activities 
+  \end{itemize}
+}
+
+\bibliographystyle{unsrt}
+\bibliography{local}
+
+\iflater
+\newpage
+\section*{Permissions for figures}
+
+[1] Creative Commons license
+
+[2] ``Reproduction of a limited number of figures or short excerpts of IPCC material is authorized free of charge and without formal written permission provided that the original source is properly acknowledged, with mention of the name of the report, the publisher and the numbering of the page(s) or the figure(s).''
+
+[3] Authors
+
+\section*{Old Text}
 
@@
 
-The problem of climate change is too large and too urgent to leave to world
-leaders: organizations at every level must play their part by addressing
-their own contributions, raising awareness among their members, and
-establishing new ways of 
-doing business in the lower-carbon future that is coming. For ACM, this
-especially means developing strategies for reducing the climate impact of
-conference-related travel.
-%
-You can help accelerate this process in a number of ways: 
-\begin{itemize}
-\item {\bf Get involved} in raising awareness about this issue in the
-organizations you are a part of.  For example, ask the organizers of the ACM
-conferences you attend what they are doing about reducing or offsetting
-emissions from participants' travel.
-\item {\bf Analyze} your organization's carbon-emitting activities to inform a
-discussion of how best to reduce its footprint.
-\item {\bf Connect} with others that are engaging with these issues---in
-your workplace, in your other communities, and within ACM (e.g., on the
-{\tt acm-climate} mailing list~\cite{acm-climate}).
-\end{itemize}
-
-\bibliographystyle{unsrt}
-\bibliography{local}
-
-\iflater
-\newpage
-\section*{Permissions for figures}
-
-[1] Creative Commons license
-
-[2] ``Reproduction of a limited number of figures or short excerpts of IPCC material is authorized free of charge and without formal written permission provided that the original source is properly acknowledged, with mention of the name of the report, the publisher and the numbering of the page(s) or the figure(s).''
-
-[3] Authors
-
 \section*{More Discussions:}
 \begin{itemize}
+\item Our notes:
+
+Remove the IPCC image and mark the other “uninformative” image as up to the design department (i.e., we’re happy to keep it or drop it or change it to something else)
+
+The map of registration data should be improved at least by reformatting and possibly by better data analysis, if we can get any of that done in time.
+I need to schedule a meeting with Aalok and Yannick about that.
+
+Some of the more “incremental” ideas can be removed — we included some that we ourselves were not very excited about.
+
+We can make clearer that carbon offsets are controversial, and why
+
+We can use the space gained by some of the above to add some discussion at the end about:
+
+Why it’s important that ACM be serious about making serious cuts (i.e., because it gives us moral authority and sets a good example for other parts of society), even though in some sense anything that we (or any individual or “small” group does is a drop in the ocean).
+
+The fact that ACM should also be thinking further ahead, which may mean fundamentally questioning its entire business model, along with a call to start this conversation.
+
 \item New comments from Judith Bishop:

### 2019-05-25 rewrite.tex
Progress on a rewrite

@@
 ACM's vibrant conference culture has a dark side: it contributes
-significantly to climate change. What should we do about it?
+significantly to climate change. What should we do about this?
 
@@
 
-Experts say that human civilization must reduce CO2 emissions by 40\%
-by 2030. To do that, we all need to tighten our CO2-emitting belts. For
-academic computer scientists, from what activity should the CO2 reduction come?
-Look no further than our professional conferences. Air
-travel is CO2 heavy, on a per-capita basis. A few trans-oceanic trips
-is XXX CO2. Wow!
+Experts say that human civilization must reduce CO2 emissions by 40\% by
+2030. To do that, we all need to tighten our CO2-emitting belts.  Where
+should this CO2 reduction come from for computer science researchers?  Look
+no further than our professional conferences. Air travel is CO2 heavy, on a
+per-capita basis. A few trans-oceanic trips is XXX CO2. Wow!
 
-As responsible scientists wanting to do our part to sustain our
-environment, should we eliminate (most) conferences, and thereby the
-travel to organize and attend them? Doing so could impose a high cost
-on scientific progress. Science is a fundamentally social process. The
+As responsible scientists wanting to do our part to sustain the
+habitability of the planet, should we eliminate most or all conferences, and
+thus the travel to organize and attend them? Doing so could impose a high
+cost on scientific progress. Science is a fundamentally social process. The
 conference system leverages that, accelerating research through easy
 interaction, direct dissemination of results, and network building to
-further new work. It works well enough that conferences are
-proliferating and attendance is on the rise. So the challenge is: How
-do we weigh the benefits to science of conferences against their
-attendant CO2 (and other) costs? This seems like an extremely
-difficult question to answer.
+further new work. It works well enough that conferences are proliferating
+and attendance is on the rise. So the challenge is: How do we weigh the
+benefits to science of conferences against their attendant CO2 (and other)
+costs? This seems like an extremely difficult question to answer.
 
@@
 
-\texdtbf{OLD STUFF HERE}
+\textbf{OLD STUFF HERE}
 

### 2019-05-25 main.tex
Progress

@@
 {\tt acm-climate} mailing list~\cite{acm-climate}).
-\item {\bf Analyze} your organization's carbon-emitting activities to inform
-a discussion of how reduce its footprint and prepare for the era of
+\item {\bf Analyze} your organization's carbon-emitting activities to
+understand how to reduce its footprint now and prepare for the coming era of
 expensive carbon.
@@
   mentioning other SIGPLAN activities
+  \item Decide what figures we want, if any
+  \item go through all the suggestions below and in the referee reports,
+  once more
+  \item Write a cover letter
+  \begin{itemize}
+  \item We tried hard to remove all incrementalism and make it much more
+  hard hitting
+  \item However, we did keep the suggestion of carbon offsets, hopefully
+  in a more convincing / palatable / better argued form
+  \item what else?
+  \end{itemize}
   \end{itemize}

### 2019-05-25 main.tex
An outline for the carbon offset part

@@
 
-\bcp{Polish and expand with reworked old text and new points.}
 Finally, we can start imposing a cost on ourselves by using carbon
-offsets.  (People counter-argue that this "lets people off the hook" and
+offsets.  
+\begin{itemize}
+\item There are two good reasons for doing this, as well as some significant
+concerns.
+\begin{itemize}
+\item Reasons in favor:
+\begin{itemize}
+\item It is the right thing to do: We should be changing our behavior
+starting now, and giving ourselves a ``price signal''
+\item Offsets reduce carbon emissions now.  (Yes, there are bogus ones, but
+not all are bogus.)
+\end{itemize}
+\item Concerns:
+\end{itemize}
+\item It is hard to choose the price (right now there is an oversupply), and setting it too low leads to complacency
+\item Organizations are not set up to do it.  Our experience with getting
+ACM to introduce an offsetting policy shows that it requires significant
+effort.
+\end{itemize}
+People counter-argue that this "lets people off the hook" and
 encourages inaction, but we believe it is consistent with encouraging action
--- again, if offsets are priced right.)
+-- again, if offsets are priced right.
 

### 2019-05-25 main.tex
Fix build

@@
 \item It is the right thing to do: We should be changing our behavior
-starting now, and giving ourselves a ``price signal''
+starting now, and giving ourselves a ``price signal'' is a good way to
+achieve this.
 \item Offsets reduce carbon emissions now.  (Yes, there are bogus ones, but
@@
 \item Concerns:
-\end{itemize}
-\item It is hard to choose the price (right now there is an oversupply), and setting it too low leads to complacency
-\item Organizations are not set up to do it.  Our experience with getting
+\begin{itemize}
+\item 
+Organizations are not set up to do it.  Our experience with getting
 ACM to introduce an offsetting policy shows that it requires significant
 effort.
+\item It is hard to choose the price (right now there is an oversupply), and
+setting it too low leads to complacency --- it doesn't really bite, so it
+doesn't change people's behavior
 \end{itemize}
-People counter-argue that this "lets people off the hook" and
-encourages inaction, but we believe it is consistent with encouraging action
--- again, if offsets are priced right.
+\item People argue that offsets "let people off the hook" and encourage
+inaction, but we believe it is consistent with encouraging action -- again,
+if offsets are priced right.
+\end{itemize}
+\end{itemize}
 
-The climate crisis is too large and too urgent to just delegate to world
-leaders: organizations at every social scale must play their part by addressing
-their own contributions, raising awareness among their members, and
-establishing new ways of 
-doing business in the lower-carbon future that is coming. For ACM, this
-especially means developing strategies for reducing the climate impact of
-conference-related travel.
+The climate crisis is too large and too urgent to leave it to world leaders
+to address at their own pace: organizations at every social scale should
+confront their own contributions, raise awareness among their members, and
+begin establishing new ways of doing business in the lower-carbon future
+that is coming. For ACM, this especially means confronting the
+carbon-intensive nature of our conferences.
 %
@@
 \item {\bf Get involved} in raising awareness about this issue in the
-organizations you are a part of.  \bcp{Does this smack of incrementalism?
-  Maybe delete it.
-  ``For example, ask the organizers of the conferences you attend what they
-  are doing about reducing or offsetting emissions from participants'
-  travel.''}
+organizations you are a part of.  \bcp{Does the rest of this smack of
+  incrementalism?  Maybe delete it.  ``For example, ask the organizers of
+  the conferences you attend what they are doing about reducing or
+  offsetting emissions from participants' travel.''}
 \item {\bf Connect} with others that are engaging with these issues---in

### 2019-05-25 main.tex
Better outline

@@
 concerns.
-\begin{itemize}
-\item Reasons in favor:
-\begin{itemize}
-\item It is the right thing to do: We should be changing our behavior
-starting now, and giving ourselves a ``price signal'' is a good way to
-achieve this.
-\item Offsets reduce carbon emissions now.  (Yes, there are bogus ones, but
-not all are bogus.)
-\end{itemize}
-\item Concerns:
-\begin{itemize}
-\item 
+\item Concern:
 Organizations are not set up to do it.  Our experience with getting
@@
 effort.
-\item It is hard to choose the price (right now there is an oversupply), and
+\item Concern: It is hard to choose the price (right now there is an oversupply), and
 setting it too low leads to complacency --- it doesn't really bite, so it
@@
 \end{itemize}
-\item People argue that offsets "let people off the hook" and encourage
+\item This issue leads many people to argue that offsets "let people off the hook" and encourage
 inaction, but we believe it is consistent with encouraging action -- again,
 if offsets are priced right.
-\end{itemize}
+\item Moreover, offsets are an effective way to reduce carbon emissions {\em
+  now}.  (Yes, there are bogus ones, but not all are bogus.)
+\item Most importantly, it is the right thing to do: We should be changing
+our behavior starting now, and giving ourselves a (correct) ``price signal''
+is an effective way to achieve this.
 \end{itemize}

### 2019-05-26 main.tex
Notes

@@
 of carbon pricing, which places a monetary cost on greenhouse gas emissions,
-giving a market signal that changes aggregate behavior.  Exactly what the
+giving a market signal that changes aggregate behavior.\bcp{See
+  https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
+for lots of good phrasing and citations.}
+Exactly what the
 cost should be is a challenging question, but it has a clear ``success

### 2019-05-28 main.tex
Fix build
Tidy

@@
 \ref{fig:warming}). This is not a problem we can leave to future
-generations: we are up against it now. Indeed, the UN's Intergovernmental
+generations: we are up against it now. The UN's Intergovernmental
 Panel on Climate Change (IPCC) recently recommended a 40\% decrease in
@@
 require urgent and sustained commitment at all levels of society, including
-non-governmental communities such as universities, companies, and scientific
+non-governmental organizations such as universities, companies, and scientific
 societies like ACM.
 
-Why ACM? Because one of ACM's main activities is organizing international
+Indeed, ACM arguably has an especially important role to play, since one of
+its main activities is organizing international
 conferences, which produce significant emissions of greenhouse gases,
-particularly from air travel. For example, a single round-trip flight from
+particularly from air travel.   A single round-trip flight from
 Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon
@@
 
-Obviously, conferences also create value.  Science is a fundamentally social
-process, and the conference system accelerates research through
-opportunities for high-bandwidth interaction, direct dissemination of
-results, network building, and serendipitous cross-fertilization.
+On the other hand, conferences obviously also create value.  Science is a
+fundamentally social process, and the conference system accelerates research
+through opportunities for high-bandwidth interaction, direct dissemination
+of results, network building, and serendipitous cross-fertilization.
 
-This creates a dilemma: How do we weigh the environmental harm caused by
+This leads to a puzzle: How do we weigh the environmental harm caused by
 conference travel against the social benefits of the scientific progress
@@
 %
-Fortunately, we have a well-established mechanism for comparing apples and
+Fortunately, there is a well-established mechanism for comparing apples and
 oranges---indeed, for comparing almost anything: the magic of markets.  By
@@
 doesn't change people's behavior
-\end{itemize}
 \item This issue leads many people to argue that offsets "let people off the hook" and encourage
@@
   once more
+  \item Think about adding headings, per Bishop's recommendation
+  \item Go through all the comments and ideas below (including the CACM
+  reviews) one last time
+  \item Change to a catchier title (``so something more personal such as
+  "Your conference trip and climate change" might get more readers'') 
   \item Write a cover letter
@@
 \end{itemize}
-\item \item 
+\item 
 Check where we've used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.

### 2019-05-28 main.tex
Tidy more

@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=4in]{wmo_stripes.png}
-\caption{Variation in annual global temperature, 1850 to 2018 (range from
-  blue to red: $1.35^\circ$C) \cite{WarmingStripes}}
-\label{fig:warming}
-\end{figure}
+% \begin{figure}[t]
+% \centering
+% \includegraphics[width=4in]{wmo_stripes.png}
+% \caption{Variation in annual global temperature, 1850 to 2018 (range from
+%   blue to red: $1.35^\circ$C) \cite{WarmingStripes}}
+% \label{fig:warming}
+% \end{figure}
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=2in]{IPCC-cover.png}
-\caption{2018 IPCC report \cite{IPCC18}.  Key finding: Emissions must
-  decline by at least 40 percent by 2030 and reach net zero by 2050, if we
-  are to hold warming to 1.5 degrees.  There is ``no documented historic
-  precedent'' for the transformation of the world economy needed to achieve
-  this.  }
-\label{fig:IPCC-cover}
-\end{figure}
+% \begin{figure}[t]
+% \centering
+% \includegraphics[width=2in]{IPCC-cover.png}
+% \caption{2018 IPCC report \cite{IPCC18}.  Key finding: Emissions must
+%   decline by at least 40 percent by 2030 and reach net zero by 2050, if we
+%   are to hold warming to 1.5 degrees.  There is ``no documented historic
+%   precedent'' for the transformation of the world economy needed to achieve
+%   this.  }
+% \label{fig:IPCC-cover}
+% \end{figure}
 
@@
 A broad scientific consensus warns that the earth is warming, due to human
-emissions of carbon dioxide and other greenhouse gases (Figure
-\ref{fig:warming}). This is not a problem we can leave to future
+emissions of carbon dioxide and other greenhouse gases.
+% (Figure \ref{fig:warming}).
+This is not a problem we can leave to future
 generations: we are up against it now. The UN's Intergovernmental
 Panel on Climate Change (IPCC) recently recommended a 40\% decrease in
-emissions by 2030 to avoid irreversible damage
-(Figure~\ref{fig:IPCC-cover}). Achieving reductions on this scale will
+emissions by 2030 to avoid irreversible damage~\cite{IPCC-report}.
+% (Figure~\ref{fig:IPCC-cover}).
+Achieving reductions on this scale will
 require urgent and sustained commitment at all levels of society, including

### 2019-05-28 main.tex
Telecon

@@
 
-\title{\Huge ACM and Climate Change}
+\title{\Huge Conferences and the Price of Carbon}
 \author{Benjamin C. Pierce (University of Pennsylvania) \\
@@
   for lots of good phrasing and citations.}
-Exactly what the
-cost should be is a challenging question, but it has a clear ``success
-criterion'': The cost should be set so that the world arrives at zero
+Exactly how this should be done is a challenging question, but it has a
+clear success 
+criterion: The cost should be set so that the world arrives at zero
 emissions within the timeframe that climate scientists tell us is safe.
+%
+For ACM, this means that the carbon cost of conferences needs to be
+reflected in their monetary cost.
 
-% (Experts seem to agree that the current price of around \$15 / ton is too
-% low.)
+Ultimately, this cost should be imposed from outside, by governments.
+Indeed, many countries and states (e.g., California) have already adopted
+carbon pricing policies; more will follow.
+%
+Moreover, although the current market price of carbon is still quite low
+(around \$15 / ton), experts agree that this will need to rise significantly
+over time to achieve the required effect.
+%
+Expecting that this is coming, what should we rationally do now to ensure
+that ACM, and computer-science research broadly, continue to thrive in a
+world where air travel is more expensive?  Believing that it needs to come
+quickly, what should we ethically do now to accelerate the process?
+Recognizing that the climate change is the defining crisis of our time, how
+can ACM and its members participate in addressing it?
 
-Many countries and states (e.g., California) have already adopted carbon
-pricing policies; more will follow.  The upshot is that, sooner or later,
-carbon is going to be expensive.
-\bcp{Yes, this may not happen.  But if it does not, we are probably
-  screwed.  So might as well behave as if we knew it was going to happen.}
-%
-In the meantime, knowing that this is coming, what should we rationally (and
-ethically) do now?
+First, we can begin gathering and analyzing data, to better understand the
+carbon footprint of our conferences and the likely effects (both
+environmental and scientific) of potential changes in the way they are
+organized.  \bcp{Put in the most interesting discussion and perhaps show an
+  interesting diagram.}
 
-First, we should immediately start planning for how ACM, and computer-science
-research more broadly, are going to continue to thrive in a world where air
-travel is significantly more expensive.  Indeed, this may well involve
-rethinking the whole basis of how computer science is done. 
-\bcp{Add some meat to that.}
+% \bcp{Add some meat to that.
+%   \begin{itemize}
+%   \item List a few things we can do
+%   \begin{itemize}
+%   \item journalize conferences
+%   \item go virtual
+%   \item potential parts of a solution: regional conferences, virtual PC
+%   meetings, virtual participation, mega-conferences
+%   \end{itemize}
+%   \item cite our report
+%   \end{itemize}
+% }
 
-Second, we can begin gathering and analyzing data, to better understand both
-the current carbon impact of our conferences and the likely effects (both
-environmental and scientific) of potential changes in the way they are
-organized.
-\bcp{Put in the most interesting discussion and perhaps show an interesting
-  diagram.}
-
-Finally, we can start imposing a cost on ourselves by using carbon
+Second, we can start imposing a cost on ourselves by using carbon
 offsets.  
@@
 
+Ultimately, responding to the inevitability of expensive carbon may well
+involve rethinking the whole basis of how computer science is done.
+\bcp{Add a bit of meat about how the current economic model may be
+  unsustainable and reworking it will involve substantial organizational
+  changes.}
+
 The climate crisis is too large and too urgent to leave it to world leaders
@@
 carbon-intensive nature of our conferences.
-%
-You can help accelerate this process in a number of ways: 
-\begin{itemize}
-\item {\bf Get involved} in raising awareness about this issue in the
-organizations you are a part of.  \bcp{Does the rest of this smack of
-  incrementalism?  Maybe delete it.  ``For example, ask the organizers of
-  the conferences you attend what they are doing about reducing or
-  offsetting emissions from participants' travel.''}
-\item {\bf Connect} with others that are engaging with these issues---in
-your workplace, in your other communities, and within ACM (e.g., on the
-{\tt acm-climate} mailing list~\cite{acm-climate}).
-\item {\bf Analyze} your organization's carbon-emitting activities to
-understand how to reduce its footprint now and prepare for the coming era of
-expensive carbon.
-\end{itemize}
+% %
+% You can help accelerate this process in a number of ways: 
+% \begin{itemize}
+% \item {\bf Get involved} in raising awareness about this issue in the
+% organizations you are a part of.  \bcp{Does the rest of this smack of
+%   incrementalism?  Maybe delete it.  ``For example, ask the organizers of
+%   the conferences you attend what they are doing about reducing or
+%   offsetting emissions from participants' travel.''}
+% \item {\bf Connect} with others that are engaging with these issues---in
+% your workplace, in your other communities, and within ACM (e.g., on the
+% {\tt acm-climate} mailing list~\cite{acm-climate}).
+% \item {\bf Analyze} your organization's carbon-emitting activities to
+% understand how to reduce its footprint now and prepare for the coming era of
+% expensive carbon.
+% \end{itemize}
 
@@
   reviews) one last time
-  \item Change to a catchier title (``so something more personal such as
-  "Your conference trip and climate change" might get more readers'') 
+  \ite Find a place for this: {\tt acm-climate} mailing list~\cite{acm-climate}
   \item Write a cover letter

### 2019-06-03 main.tex
Fix build
CP

@@
 
-\title{\Huge Conferences and the Price of Carbon}
+\title{\Huge Conferences and Cost of Carbon}
 \author{Benjamin C. Pierce (University of Pennsylvania) \\
@@
 criterion: The cost should be set so that the world arrives at zero
-emissions within the timeframe that climate scientists tell us is safe.
+Emissions within the timeframe that climate scientists tell us is safe.
 %
-For ACM, this means that the carbon cost of conferences needs to be
+For ACM, this means that the carbon cost of conferences should be
 reflected in their monetary cost.
 
-Ultimately, this cost should be imposed from outside, by governments.
+% Expecting that this is coming, what should we rationally do now to ensure
+% that ACM, and computer-science research broadly, continue to thrive in a
+% world where air travel is more expensive?  Believing that it needs to come
+% quickly, what should we ethically do now to accelerate the process?
+% Recognizing that the climate change is the defining crisis of our time, how
+% can ACM and its members participate in addressing it?
+
+In the longer term, this cost should be imposed from outside, by governments.
 Indeed, many countries and states (e.g., California) have already adopted
@@
 %
-Expecting that this is coming, what should we rationally do now to ensure
-that ACM, and computer-science research broadly, continue to thrive in a
-world where air travel is more expensive?  Believing that it needs to come
-quickly, what should we ethically do now to accelerate the process?
-Recognizing that the climate change is the defining crisis of our time, how
-can ACM and its members participate in addressing it?
+
+%
+In the short term, we can choose to do it voluntarily using carbon
+offsets.  
+\begin{itemize}
+\item (What they are.)
+\item giving ourselves a ``price signal''
+is an effective way to begin raising awareness and changing organizational
+behavior.
+\item Lots of organizations (universities, companies) are doing this; some
+ACM conferences are starting to experiment with voluntary carbon offset fees
+at registration time. 
+\item There are also some significant
+concerns.
+\item Concern:
+Organizations are not set up to do it.  Our experience with getting
+ACM to introduce an offsetting policy shows that it requires significant
+effort.
+\item Concern: It is hard to choose the price (right now there is an oversupply), and
+setting it too low leads to complacency --- it doesn't really bite, so it
+doesn't change people's behavior
+\item This issue leads many people to argue that offsets "let people off the hook" and encourage
+inaction, but we believe it is consistent with encouraging action -- again,
+if offsets are priced right.
+\end{itemize}
 
@@
 
-Second, we can start imposing a cost on ourselves by using carbon
-offsets.  
-\begin{itemize}
-\item There are two good reasons for doing this, as well as some significant
-concerns.
-\item Concern:
-Organizations are not set up to do it.  Our experience with getting
-ACM to introduce an offsetting policy shows that it requires significant
-effort.
-\item Concern: It is hard to choose the price (right now there is an oversupply), and
-setting it too low leads to complacency --- it doesn't really bite, so it
-doesn't change people's behavior
-\item This issue leads many people to argue that offsets "let people off the hook" and encourage
-inaction, but we believe it is consistent with encouraging action -- again,
-if offsets are priced right.
-\item Moreover, offsets are an effective way to reduce carbon emissions {\em
-  now}.  (Yes, there are bogus ones, but not all are bogus.)
-\item Most importantly, it is the right thing to do: We should be changing
-our behavior starting now, and giving ourselves a (correct) ``price signal''
-is an effective way to achieve this.
-\end{itemize}
-
 Ultimately, responding to the inevitability of expensive carbon may well
@@
   changes.} 
+\begin{itemize}
+\item monitoring and mitigating carbon costs should be an ACM-level activity
+\end{itemize}
 
@@
   reviews) one last time
-  \ite Find a place for this: {\tt acm-climate} mailing list~\cite{acm-climate}
+  \item Find a place for this: {\tt acm-climate} mailing list~\cite{acm-climate}
   \item Write a cover letter

### 2019-06-03 main.tex
CP

@@
 Emissions within the timeframe that climate scientists tell us is safe.
-%
-For ACM, this means that the carbon cost of conferences should be
-reflected in their monetary cost.
 
-% Expecting that this is coming, what should we rationally do now to ensure
-% that ACM, and computer-science research broadly, continue to thrive in a
-% world where air travel is more expensive?  Believing that it needs to come
-% quickly, what should we ethically do now to accelerate the process?
-% Recognizing that the climate change is the defining crisis of our time, how
-% can ACM and its members participate in addressing it?
-
-In the longer term, this cost should be imposed from outside, by governments.
+In the long term, this cost must be imposed from outside, by governments.
 Indeed, many countries and states (e.g., California) have already adopted
@@
 (around \$15 / ton), experts agree that this will need to rise significantly
-over time to achieve the required effect.
-%
+over time to achieve the required effect.  Bottom line: carbon is going to
+be expensive.
 
-%
-In the short term, we can choose to do it voluntarily using carbon
-offsets.  
+Expecting that this is coming, what should we rationally do now to ensure
+that ACM, and computer-science research broadly, continue to thrive in a
+world where air travel is more expensive?  Believing that it needs to come
+quickly, what should we ethically do now to accelerate the process?
+
+A short-term response to both questions: ACM can choose to voluntarily impose
+a price on its emissions using carbon offsets. 
 \begin{itemize}
@@
 
-First, we can begin gathering and analyzing data, to better understand the
+Whether we choose to impose carbon offsets on ourselves or wait for
+governments to do it, we need to begin gathering and analyzing data, to
+better understand the 
 carbon footprint of our conferences and the likely effects (both
@@
   interesting diagram.}
+\bcp{Add some meat to that.
+  \begin{itemize}
+  \item List a few things we can do (and whose likely effects we'd need data
+  to estimate)
+  \begin{itemize}
+  \item journalize conferences
+  \item go virtual
+  \item potential parts of a solution: regional conferences, virtual PC
+  meetings, virtual participation, mega-conferences
+  \end{itemize}
+  \item cite our report
+  \end{itemize}
+}
 
-% \bcp{Add some meat to that.
-%   \begin{itemize}
-%   \item List a few things we can do
-%   \begin{itemize}
-%   \item journalize conferences
-%   \item go virtual
-%   \item potential parts of a solution: regional conferences, virtual PC
-%   meetings, virtual participation, mega-conferences
-%   \end{itemize}
-%   \item cite our report
-%   \end{itemize}
-% }
-
-Ultimately, responding to the inevitability of expensive carbon may well
-involve rethinking the whole basis of how computer science is done.
-\bcp{Add a bit of meat about how the current economic model may be
+Ultimately, responding to the inevitability of expensive carbon may involve
+rethinking the whole basis of how academic computer science is done.
+\bcp{
+\begin{itemize}
+\iten Add a bit of meat about how the current economic model may be
   unsustainable and reworking it will involve substantial organizational
-  changes.} 
-\begin{itemize}
+  changes.
 \item monitoring and mitigating carbon costs should be an ACM-level activity
 \end{itemize}
+}
 

### 2019-06-03 main.tex
Final CP after call

@@
 
-A short-term response to both questions: ACM can choose to voluntarily impose
-a price on its emissions using carbon offsets. 
+One possible short-term response to both questions: ACM can choose to
+voluntarily impose a price on its emissions using carbon offsets.
 \begin{itemize}
@@
 Whether we choose to impose carbon offsets on ourselves or wait for
-governments to do it, we need to begin gathering and analyzing data, to
-better understand the 
-carbon footprint of our conferences and the likely effects (both
-environmental and scientific) of potential changes in the way they are
-organized.  \bcp{Put in the most interesting discussion and perhaps show an
-  interesting diagram.}
-\bcp{Add some meat to that.
+governments to do it, there are also many things we can do to begin reducing
+our footprint.  
+\bcp{
   \begin{itemize}
-  \item List a few things we can do (and whose likely effects we'd need data
+  \item Briefly list a few things we can do (and whose likely effects we'd need data
   to estimate)
@@
 }
+However, we urgently need to begin gathering and analyzing data, to
+better understand the positive and negative effects of any of these
+proposals.   
+\bcp{Put the discussion of Figure 1 (and an accompanying figure
+showing distance traveled to several instances of the same conference) here;
+this illustrates the difficulty and the need for careful data gathering and
+analysis.} 
+All parts of ACM should be doing this kind of analysis.  (The acm-climate
+mailing list is one place to exchange ideas about this.)
 
@@
   \begin{itemize}
-  \item citations for the claims we make
-  \item an excuse for citing the Climate Committee report and perhaps
-  mentioning other SIGPLAN activities
-  \item Decide what figures we want, if any
-  \item go through all the suggestions below and in the referee reports,
-  once more
-  \item Think about adding headings, per Bishop's recommendation
+  \item citations for the claims we make (and the carbon-offsets-bad paper
+  mentioned by the reviewer)
   \item Go through all the comments and ideas below (including the CACM
@@
   \item Find a place for this: {\tt acm-climate} mailing list~\cite{acm-climate}
+  \item Think about adding headings, per Bishop's recommendation
+
   \item Write a cover letter

### 2019-06-03 main.tex
Added distribution of participants' origin

@@
 \includegraphics[width=6in]{SIGPLAN-confs.png}
-\caption{Carbon footprint per participant for travel to recent SIGPLAN
+\includegraphics[width=6in]{ParticipantsOriginAll.png}
+\caption{Top: carbon footprint per participant for travel to recent SIGPLAN
   conferences\iflater\bcp{Get better file from the other paper, or get
-    better version of this one from Crista.}\fi} 
+    better version of this one from Crista.}\fi Bottom: origin of participants for each conference.} 
 \label{fig:confs}
@@
 \begin{itemize}
-\iten Add a bit of meat about how the current economic model may be
+\item Add a bit of meat about how the current economic model may be
   unsustainable and reworking it will involve substantial organizational

### 2019-06-04 main.tex
Checkpoint work on writing

@@
 
-\title{\Huge Conferences and Cost of Carbon}
+\title{\Huge Conferences in the Era of Expensive Carbon}
 \author{Benjamin C. Pierce (University of Pennsylvania) \\
@@
 
-A broad scientific consensus warns that the earth is warming, due to human
-emissions of carbon dioxide and other greenhouse gases.
+A broad scientific consensus warns us that human
+emissions of carbon dioxide and other greenhouse gases are warming the earth.
 % (Figure \ref{fig:warming}).
-This is not a problem we can leave to future
+This is not a problem to be left to future
 generations: we are up against it now. The UN's Intergovernmental
-Panel on Climate Change (IPCC) recently recommended a 40\% decrease in
-emissions by 2030 to avoid irreversible damage~\cite{IPCC-report}.
+Panel on Climate Change (IPCC) says a 40\% decrease in
+emissions by 2030 is needed to avoid irreversible damage~\cite{IPCC-report}.
 % (Figure~\ref{fig:IPCC-cover}).
@@
 require urgent and sustained commitment at all levels of society, including
-non-governmental organizations such as universities, companies, and scientific
-societies like ACM.
+large organizations like universities, companies, and scientific societies
+like ACM.
 
-Indeed, ACM arguably has an especially important role to play, since one of
-its main activities is organizing international
-conferences, which produce significant emissions of greenhouse gases,
+Indeed, scientific societies arguably have an especially important role to
+play, since one of their main activities is organizing international
+conferences, which produce significant greenhouse gas emissions,
 particularly from air travel.   A single round-trip flight from
@@
 Moreover, these emissions have no near-term 
-technological fix because jet fuel is difficult to replace with
+technological fix, since jet fuel is difficult to replace with
 renewable energy sources~\cite{elec-air}. 
-\iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
-of conferences will require a reduction in the footprint of the attendees'
-air travel.}\fi
+% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
+% of conferences will require a reduction in the footprint of the attendees'
+% air travel.}\fi
 
-On the other hand, conferences obviously also create value.  Science is a
+Conferences obviously also create significant value.  Science is a
 fundamentally social process, and the conference system accelerates research
-through opportunities for high-bandwidth interaction, direct dissemination
-of results, network building, and serendipitous cross-fertilization.
+through high-bandwidth interaction, direct dissemination of results, network
+building, and serendipitous cross-fertilization.
 
@@
 conference travel against the social benefits of the scientific progress
-that comes from conferences?  This seems an extremely difficult question to
+that comes from conferences?  It seems an extremely difficult question to
 answer: the harm and the compensating benefits cannot be directly
@@
 Fortunately, there is a well-established mechanism for comparing apples and
-oranges---indeed, for comparing almost anything: the magic of markets.  By
-mapping choices to currency amounts, markets facilitate individual valuation
-of costs and benefits across vastly different activities.\bcp{Give a simple,
-  concrete example.} For this reason, many policy experts advocate some form
-of carbon pricing, which places a monetary cost on greenhouse gas emissions,
-giving a market signal that changes aggregate behavior.\bcp{See
+oranges (indeed, for comparing almost anything): the magic of markets.  By
+mapping choices and preferences in disparate domains to a common scale
+(namely money), economic markets can be viewed as massively distributed
+computations that balance costs an benefits across vastly different
+activities.\bcp{Give a simple, concrete example.}  For this reason, many
+policy experts advocate addressing climate change through some form of
+carbon pricing, which places a monetary cost on greenhouse gas emissions,
+thus giving a market signal that changes aggregate behavior.\bcp{See
   https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
-  for lots of good phrasing and citations.}
-Exactly how this should be done is a challenging question, but it has a
-clear success 
-criterion: The cost should be set so that the world arrives at zero
-Emissions within the timeframe that climate scientists tell us is safe.
+  for lots of good phrasing and citations.}  Exactly how the price of carbon
+should be set is a challenging policy question, but it has a clear goal: The
+cost should be set so that the world arrives at zero emissions within the
+timeframe that climate scientists tell us is safe.
 
-In the long term, this cost must be imposed from outside, by governments.
-Indeed, many countries and states (e.g., California) have already adopted
-carbon pricing policies; more will follow.
+In the long term, a carbon price should be uniform across society---i.e., it
+should be imposed by governments.  Indeed, many countries and states (e.g.,
+California) have already adopted carbon pricing policies; more will follow.
 %
 Moreover, although the current market price of carbon is still quite low
-(around \$15 / ton), experts agree that this will need to rise significantly
-over time to achieve the required effect.  Bottom line: carbon is going to
-be expensive.
+(around \$15 per ton), experts agree that this will need to rise
+significantly over time to achieve the required effect.  Bottom line: We are
+entering an era in which carbon is expensive.
 
-Expecting that this is coming, what should we rationally do now to ensure
-that ACM, and computer-science research broadly, continue to thrive in a
-world where air travel is more expensive?  Believing that it needs to come
-quickly, what should we ethically do now to accelerate the process?
+Expecting that this era is coming, what should we rationally do now to
+ensure that ACM, and computer-science research broadly, continue to thrive
+in a world where air travel is more expensive?  Believing that it needs to
+come as quickly as possible to avert disaster, what can we do to accelerate
+the process?
 
-One possible short-term response to both questions: ACM can choose to
-voluntarily impose a price on its emissions using carbon offsets.
+A possible short-term response to both questions involves {\em carbon
+  offsets}---a tool by which organizations (and individuals) can choose to
+voluntarily impose a price on their own emissions.
+%
+% What they are:
+In a nutshell, the idea would be that a conference could raise funds from
+registration fees or corporate sponsors and give them, via a carbon offset
+{\em vendor}, to someone who will use them to fund an activity (e.g.,
+installing methane capture devices on landfill sites or buying efficient
+cookstoves to replace open-fire cooking in developing areas) that will avoid
+emitting an amount of greenhouse gases equivalent to the emissions from
+participants' travel to the conference.
+%
 \begin{itemize}
-\item (What they are.)
 \item giving ourselves a ``price signal''

### 2019-06-04 main.tex
Checkpoint work on writing

@@
 In the long term, a carbon price should be uniform across society---i.e., it
-should be imposed by governments.  Indeed, many countries and states (e.g.,
-California) have already adopted carbon pricing policies; more will follow.
+should be imposed by governments.  Indeed, many countries\bcp{e.g.?} and
+states (e.g., California) have already adopted carbon pricing policies; more
+will follow.
 %
-Moreover, although the current market price of carbon is still quite low
-(around \$15 per ton), experts agree that this will need to rise
-significantly over time to achieve the required effect.  Bottom line: We are
-entering an era in which carbon is expensive.
+Bottom line: We are entering an era in which carbon will be expensive.
 
@@
 A possible short-term response to both questions involves {\em carbon
-  offsets}---a tool by which organizations (and individuals) can choose to
-voluntarily impose a price on their own emissions.
+  offsets}---a tool by which organizations (and individuals) can voluntarily
+impose a price on their own emissions.
 %
 % What they are:
-In a nutshell, the idea would be that a conference could raise funds from
+In a nutshell, the idea is that a conference can raise funds from
 registration fees or corporate sponsors and give them, via a carbon offset
-{\em vendor}, to someone who will use them to fund an activity (e.g.,
-installing methane capture devices on landfill sites or buying efficient
-cookstoves to replace open-fire cooking in developing areas) that will avoid
-emitting an amount of greenhouse gases equivalent to the emissions from
-participants' travel to the conference.
+{\em vendor}, to a {\em provider} who will use them to support an activity
+(e.g., installing methane capture devices on landfill sites or buying
+efficient cookstoves to replace open-fire cooking in developing areas) that
+avoids emitting an amount of greenhouse gases equivalent to the emissions
+from participants' travel to the conference.
 %
 \begin{itemize}
-\item giving ourselves a ``price signal''
-is an effective way to begin raising awareness and changing organizational
-behavior.
-\item Lots of organizations (universities, companies) are doing this; some
-ACM conferences are starting to experiment with voluntary carbon offset fees
-at registration time. 
+\item Lots of organizations (universities, companies) are already doing
+this; some ACM conferences are starting to experiment with voluntary carbon
+offset fees at registration time.
+\end{itemize}
+
+\begin{itemize}
 \item There are also some significant
 concerns.
-\item Concern:
-Organizations are not set up to do it.  Our experience with getting
-ACM to introduce an offsetting policy shows that it requires significant
-effort.
-\item Concern: It is hard to choose the price (right now there is an oversupply), and
-setting it too low leads to complacency --- it doesn't really bite, so it
-doesn't change people's behavior
-\item This issue leads many people to argue that offsets "let people off the hook" and encourage
-inaction, but we believe it is consistent with encouraging action -- again,
-if offsets are priced right.
+\item Concern: Organizations are not set up to do it, and the piecemeal
+political and bureaucratic work of designing and implementing a separate
+policy for each organization is exhausting.
+\item Concern: It is hard to choose the price (right now there is an
+oversupply), and setting it too low leads to complacency --- it doesn't
+really bite, so it doesn't change people's behavior.  The current market
+price of carbon is still quite low (around \$15 per ton), whereas the
+``social cost of carbon'' is variously estimated at anywhere from \$40 to
+\$400 per ton~\bcp{citation}.
+Experts agree that the current price will need to rise significantly
+over time to achieve the required effect.
+\item This issue leads many people to argue that offsets "let people off the
+hook" and encourage inaction, but we believe it is consistent with
+encouraging action -- especially if offsets are priced right, but even if
+(in the short term) they are too cheap -- because it injects a ``price
+signal'' that makes emissions something for planners to optimize.
+\item We feel that, on balance, voluntary carbon offsets are an effective
+tools for raising awareness and changing organizational behavior in the
+short term.
+\item Moreover, supporting well-chosen carbon mitigation activities has a
+clear benefit and can be done now.
 \end{itemize}
 
-Whether we choose to impose carbon offsets on ourselves or wait for
-governments to do it, there are also many things we can do to begin reducing
-our footprint.  
+Beyond imposing a price on our carbon footprint (or waiting for one to be
+imposed from outside), a sane response to climate change should clearly also
+include taking steps to reduce the size of this footprint.  
 \bcp{
   \begin{itemize}
-  \item Briefly list a few things we can do (and whose likely effects we'd need data
-  to estimate)
+  \item Briefly list a few things we can do
   \begin{itemize}
@@
 }
-However, we urgently need to begin gathering and analyzing data, to
-better understand the positive and negative effects of any of these
-proposals.   
+%
+However, before we undertake large changes, we need to understand the
+situation better.  For most of these ideas, it is hard to confidently
+predict either the potential benefits (in terms of reduced emissions) or the
+potential costs (to scientific exchange, career development of young
+scientists, etc.).  Indeed, a preliminary analysis of registration data from
+a number of SIGPLAN conferences paints a rather complicated picture.
+%
 \bcp{Put the discussion of Figure 1 (and an accompanying figure
@@
   unsustainable and reworking it will involve substantial organizational
-  changes.
+  changes.  (There may be some good words that we can paraphrase from our
+  second reviewer!)
 \item monitoring and mitigating carbon costs should be an ACM-level activity
@@
 
-The climate crisis is too large and too urgent to leave it to world leaders
-to address at their own pace: organizations at every social scale should
-confront their own contributions, raise awareness among their members, and
-begin establishing new ways of doing business in the lower-carbon future
-that is coming. For ACM, this especially means confronting the
-carbon-intensive nature of our conferences.
+The climate crisis is too large and too urgent to leave to world leaders to
+address at their own pace: organizations at every scale should confront
+their own contributions, raise awareness among their members, and begin
+establishing new ways of doing business in the lower-carbon future that is
+coming.
+% For ACM, this especially means confronting the carbon-intensive
+% nature of our conferences.
 % %
@@
   \begin{itemize}
-  \item citations for the claims we make (and the carbon-offsets-bad paper
+  \item citations for the claims we make (and the carbon-offsets-are-bad paper
   mentioned by the reviewer)

### 2019-06-04 main.tex
More work

@@
 %
+Many organizations (including numerous universities, companies such as
+Google, Dell, Microsoft, General Motors, Delta Airlines, Expedia, and
+others) are already buying carbon offsets; some ACM conferences have already
+started to experiment with voluntary carbon offset fees at registration
+time.
+
+We acknowledge that there are some significant concerns with carbon offsets.
 \begin{itemize}
-\item Lots of organizations (universities, companies) are already doing
-this; some ACM conferences are starting to experiment with voluntary carbon
-offset fees at registration time.
-\end{itemize}
-
-\begin{itemize}
-\item There are also some significant
-concerns.
 \item Concern: Organizations are not set up to do it, and the piecemeal

### 2019-06-04 main.tex
Some more work on the middle parts

--- 2019-06-04 main.tex
@@
 \newcommand{\COtwoE}{CO$_2$e}
+
+\newcommand{\SECTION}{\paragraph*}
 
@@
 % \end{figure}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=6in]{SIGPLAN-confs.png}
-\includegraphics[width=6in]{ParticipantsOriginAll.png}
-\caption{Top: carbon footprint per participant for travel to recent SIGPLAN
-  conferences\iflater\bcp{Get better file from the other paper, or get
-    better version of this one from Crista.}\fi Bottom: origin of participants for each conference.} 
-\label{fig:confs}
-\end{figure}
 
@@
 compared---they are apples and oranges.
-%
+
+\SECTION{Putting a Price on Carbon}
+
 Fortunately, there is a well-established mechanism for comparing apples and
@@
 
-A possible short-term response to both questions involves {\em carbon
-  offsets}---a tool by which organizations (and individuals) can voluntarily
-impose a price on their own emissions.
+\SECTION{Carbon Offsets}
+
+In the short term, one possible response to both questions is to use {\em
+  carbon offsets}, which allow organizations (and individuals) to
+voluntarily impose a price on their own emissions~\cite{CarbonOFfsetReport}.
 %
-% What they are:
 In a nutshell, the idea is that a conference can raise funds from
@@
 efficient cookstoves to replace open-fire cooking in developing areas) that
-avoids emitting an amount of greenhouse gases equivalent to the emissions
-from participants' travel to the conference.
+is {\em certified} (by yet another organization) to permanently avoid
+emitting an amount of greenhouse gases equivalent to the emissions from
+participants' travel to the conference.
 %
-\begin{itemize}
-\item Lots of organizations (universities, companies) are already doing
-this; some ACM conferences are starting to experiment with voluntary carbon
-offset fees at registration time.
-\end{itemize}
+Many organizations (including companies such as Google, Dell, Microsoft,
+General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
+and a number of academic societies) are already buying carbon offsets, and
+some ACM conferences have already started to experiment with voluntary
+carbon offset fees at registration time.
 
-\begin{itemize}
-\item There are also some significant
-concerns.
-\item Concern: Organizations are not set up to do it, and the piecemeal
-political and bureaucratic work of designing and implementing a separate
-policy for each organization is exhausting.
-\item Concern: It is hard to choose the price (right now there is an
-oversupply), and setting it too low leads to complacency --- it doesn't
-really bite, so it doesn't change people's behavior.  The current market
-price of carbon is still quite low (around \$15 per ton), whereas the
-``social cost of carbon'' is variously estimated at anywhere from \$40 to
-\$400 per ton~\bcp{citation}.
-Experts agree that the current price will need to rise significantly
-over time to achieve the required effect.
-\item This issue leads many people to argue that offsets "let people off the
-hook" and encourage inaction, but we believe it is consistent with
-encouraging action -- especially if offsets are priced right, but even if
-(in the short term) they are too cheap -- because it injects a ``price
-signal'' that makes emissions something for planners to optimize.
-\item We feel that, on balance, voluntary carbon offsets are an effective
-tools for raising awareness and changing organizational behavior in the
-short term.
-\item Moreover, supporting well-chosen carbon mitigation activities has a
-clear benefit and can be done now.
-\end{itemize}
+Carbon offsets do come with some significant concerns.
+%
+A superficial but potentially sigificant issue is that most
+organizations do not currently have accounting procedures in place to deal
+with them, and the piecemeal political and bureaucratic work of designing
+and implementing a separate policy for each organization can consume
+significant energy.
+%
+A more fundamental problem is that it is not straightforward to decide how
+much it should cost to offset a ton of carbon: current prices vary widely
+around an average of around \$15 per ton, depending on the vendor, the
+nature of the offset project, the region of the world where it will be
+implemented, etc..  Worse, in order for offsets (or any other form of a
+carbon pricing) to fully motivate the needed changes in behavior, the price
+may need to go much higher---the true ``social cost'' of a ton of carbon is
+variously estimated at anywhere from \$40 to \$400 per
+ton~\bcp{citation}---and setting it too low may encourage complacency.
+%
+And finally, 
+%
+We feel that, on balance, voluntary carbon offsetting can be an
+effective tool for raising awareness and changing organizational behavior
+in the short term.
+%
+Moreover, supporting well-chosen carbon mitigation activities has a clear
+benefit in reducing global net emissions now.
 
-Beyond imposing a price on our carbon footprint (or waiting for one to be
-imposed from outside), a sane response to climate change should clearly also
-include taking steps to reduce the size of this footprint.  
-\bcp{
-  \begin{itemize}
-  \item Briefly list a few things we can do
-  \begin{itemize}
-  \item journalize conferences
-  \item go virtual
-  \item potential parts of a solution: regional conferences, virtual PC
-  meetings, virtual participation, mega-conferences
-  \end{itemize}
-  \item cite our report
-  \end{itemize}
-}
+\SECTION{Measure, Analyze, Optimize}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=6in]{SIGPLAN-confs.png}
+\includegraphics[width=6in]{ParticipantsOriginAll.png}
+\caption{Top: carbon footprint per participant for travel to recent SIGPLAN
+  conferences\iflater\bcp{Get better file from the other paper, or get
+    better version of this one from Crista.}\fi Bottom: origin of participants for each conference.} 
+\label{fig:confs}
+\end{figure}
+
+Beyond imposing a price on our carbon footprint (or waiting for one
+to be imposed from outside), a sane response to climate change should
+clearly also include taking steps to reduce the size of this footprint.  For
+conferences, there are many options that one might
+contemplate~\cite{ClimateCommitteReport}, from easy 
+but incremental changes like making program committee meetings virtual
+instead of physical and enabling virtual participation via live-streaming,
+all the way to complete re-imagining of conferences and their role in
+science (holding conferences completely in virtual spaces, replacing
+conferences with journals, etc.), with some intriguing hybrid forms in the
+middle (regional conferences, mega-conferences, conferences split into
+two locations on different continents, etc.).  
 %
@@
 mailing list is one place to exchange ideas about this.)
+
+\SECTION{Deeper Changes}
 

### 2019-06-05 main.tex
A couple of comments.

@@
 conferences, which produce significant greenhouse gas emissions,
-particularly from air travel.   A single round-trip flight from
+particularly from air travel. \mwh{The previous is the wrong
+  transition, since again it seems to suggest that travel is a big
+  deal in terms of the overall problem. The sentence also fails to
+  convey the other reason scientific societies should be involved: As
+  role models and advocates for general change.}  A single round-trip flight from
 Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon
@@
 come as quickly as possible to avert disaster, what can we do to accelerate
-the process?
+the process? \mwh{The question now seems to have changed. The paper started by
+  saying ACM has an ``especially important role to play,'' but now
+  seems to suggest it should be reacting to an inevitable/coming
+  future. These questions probably need to go much earlier, and the earlier
+  text reworked around them.}
 
@@
 to be imposed from outside), a sane response to climate change should
-clearly also include taking steps to reduce the size of this footprint.  For
+clearly also include taking steps to reduce the size of this
+footprint.
+\mwh{Don't like previous sentence. It seems to have gotten away from
+  the point of the paper, which is using price to change balance of
+  apples to oranges. Analyzing data is useful for figuring out ACM's
+  internal price, not tweaking around the edges, which is how it reads
+  below.} For
 conferences, there are many options that one might

### 2019-06-05 main.tex
More writing

@@
 %
-In a nutshell, the idea is that a conference can raise funds from
-registration fees or corporate sponsors and give them, via a carbon offset
+In a nutshell, the idea is for a conference to raise funds from
+registration fees or corporate sponsors and donate them, via a carbon offset
 {\em vendor}, to a {\em provider} who will use them to support an activity
 (e.g., installing methane capture devices on landfill sites or buying
-efficient cookstoves to replace open-fire cooking in developing areas) that
+fuel-efficient stoves to replace open-fire cooking in poor communities) that
 is {\em certified} (by yet another organization) to permanently avoid
@@
 %
-A superficial but potentially sigificant issue is that most
-organizations do not currently have accounting procedures in place to deal
-with them, and the piecemeal political and bureaucratic work of designing
-and implementing a separate policy for each organization can consume
-significant energy.
+One superficial impediment is that most organizations do not currently have
+accounting procedures in place to deal with them, and the piecemeal
+political and bureaucratic work of designing and implementing a separate
+policy for each organization can consume significant energy.
+%
+Another point to be careful of is that the long-term climate benefit of some
+offsetting strategies (for example, paying people to plant trees) is
+controversial.
 %
 A more fundamental problem is that it is not straightforward to decide how
-much it should cost to offset a ton of carbon: current prices vary widely
+much one should pay to offset a ton of carbon: current prices vary widely
 around an average of around \$15 per ton, depending on the vendor, the
 nature of the offset project, the region of the world where it will be
-implemented, etc..  Worse, in order for offsets (or any other form of a
+implemented, etc.  Moreover, in order for offsets (or any other form of a
 carbon pricing) to fully motivate the needed changes in behavior, the price
@@
 %
-And finally, 
-%
-We feel that, on balance, voluntary carbon offsetting can be an
-effective tool for raising awareness and changing organizational behavior
-in the short term.
-%
-Moreover, supporting well-chosen carbon mitigation activities has a clear
-benefit in reducing global net emissions now.
+However, while acknowledging these concerns, we believe that voluntary
+carbon offsetting can be an effective near-term tool for raising awareness
+and beginning to shift organizational behavior because it associates
+emissions with a number, giving decision makers within the organization
+something concrete to optimize.
 

### 2019-06-05 main.tex
Bit more writing

@@
 
-\SECTION{Putting a Price on Carbon}
+\SECTION{Putting a Price on Emissions}
 
 Fortunately, there is a well-established mechanism for comparing apples and
-oranges (indeed, for comparing almost anything): the magic of markets.  By
-mapping choices and preferences in disparate domains to a common scale
-(namely money), economic markets can be viewed as massively distributed
-computations that balance costs an benefits across vastly different
-activities.\bcp{Give a simple, concrete example.}  For this reason, many
-policy experts advocate addressing climate change through some form of
-carbon pricing, which places a monetary cost on greenhouse gas emissions,
-thus giving a market signal that changes aggregate behavior.\bcp{See
+oranges---indeed, for comparing almost anything!  Economists call it ``the
+magic of markets.''  By mapping choices and preferences in disparate domains
+onto a common scale (money), economic markets can be viewed as massively
+distributed computations that optimize costs and benefits across vastly
+different activities.\bcp{Give a simple, concrete example.}  For this
+reason, many policy experts advocate addressing climate change through some
+form of carbon pricing, which places a monetary cost on greenhouse gas
+emissions, sending a clear market signal that changes aggregate
+behavior.\bcp{See
   https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
   for lots of good phrasing and citations.}  Exactly how the price of carbon
-should be set is a challenging policy question, but it has a clear goal: The
-cost should be set so that the world arrives at zero emissions within the
+should be set is a challenging policy question, but it has a clear goal: It
+should be chosen so that the world arrives at zero emissions within the
 timeframe that climate scientists tell us is safe.
 
-In the long term, a carbon price should be uniform across society---i.e., it
-should be imposed by governments.  Indeed, many countries\bcp{e.g.?} and
-states (e.g., California) have already adopted carbon pricing policies; more
-will follow.
+To be effective in the long term, a carbon price needs to be uniform across
+society---i.e., it should be imposed by governments.  Indeed, many
+countries\bcp{e.g.?} and states (e.g., California) have already adopted
+carbon pricing policies; more will follow.
 %
@@
 
-Expecting that this era is coming, what should we rationally do now to
-ensure that ACM, and computer-science research broadly, continue to thrive
-in a world where air travel is more expensive?  Believing that it needs to
-come as quickly as possible to avert disaster, what can we do to accelerate
-the process? \mwh{The question now seems to have changed. The paper started by
-  saying ACM has an ``especially important role to play,'' but now
-  seems to suggest it should be reacting to an inevitable/coming
-  future. These questions probably need to go much earlier, and the earlier
-  text reworked around them.}
+Anticipating this era, what should we rationally do now to ensure that ACM,
+and computer-science research more broadly, continue to thrive in a world
+where air travel is more expensive?  Believing that it needs to come quickly
+to avert disaster, what can we do to accelerate the process? \mwh{The
+  question now seems to have changed. The paper started by saying ACM has an
+  ``especially important role to play,'' but now seems to suggest it should
+  be reacting to an inevitable/coming future. These questions probably need
+  to go much earlier, and the earlier text reworked around them.}
 

### 2019-06-10 mike.tex
Reworked.

--- 2019-06-05 main.tex
@@
 
-\title{\Huge Conferences in the Era of Expensive Carbon}
+\title{\Huge Conferences in an Era of Expensive Carbon}
+% \title{\Huge ACM Conferences and Cost of Carbon}
 \author{Benjamin C. Pierce (University of Pennsylvania) \\
@@
 
-% \begin{multicols}{3}
-
-ACM's vibrant conference culture has a dark side: it contributes
-significantly to climate change.
-
-% \begin{figure}[t]
-% \centering
-% \includegraphics[width=4in]{wmo_stripes.png}
-% \caption{Variation in annual global temperature, 1850 to 2018 (range from
-%   blue to red: $1.35^\circ$C) \cite{WarmingStripes}}
-% \label{fig:warming}
-% \end{figure}
-
-% \begin{figure}[t]
-% \centering
-% \includegraphics[width=2in]{IPCC-cover.png}
-% \caption{2018 IPCC report \cite{IPCC18}.  Key finding: Emissions must
-%   decline by at least 40 percent by 2030 and reach net zero by 2050, if we
-%   are to hold warming to 1.5 degrees.  There is ``no documented historic
-%   precedent'' for the transformation of the world economy needed to achieve
-%   this.  }
-% \label{fig:IPCC-cover}
-% \end{figure}
+% ACM's vibrant conference culture has a dark side: it contributes
+% significantly to climate change.
 
@@
 conferences, which produce significant greenhouse gas emissions,
-particularly from air travel. \mwh{The previous is the wrong
-  transition, since again it seems to suggest that travel is a big
-  deal in terms of the overall problem. The sentence also fails to
-  convey the other reason scientific societies should be involved: As
-  role models and advocates for general change.}  A single round-trip flight from
+particularly from air travel. A single round-trip flight from
 Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon
@@
 
-Conferences obviously also create significant value.  Science is a
-fundamentally social process, and the conference system accelerates research
-through high-bandwidth interaction, direct dissemination of results, network
-building, and serendipitous cross-fertilization.
+We formed the ACM SIGPLAN Climate Committee with this situation in
+mind. We made presentations at SIGPLAN events, had discussions with
+ACM and SIG leaders, and ultimately put together a report of ideas for
+reducing the \COtwoE{} costs of conference
+activities~\cite{ClimateCommitteReport}. The report considers easy 
+but incremental changes like making program committee meetings virtual
+instead of physical and enabling virtual participation via live-streaming,
+all the way to complete re-imagining of conferences and their role in
+science (holding conferences completely in virtual spaces, replacing
+conferences with journals, etc.), with some intriguing hybrid forms in the
+middle (regional conferences, mega-conferences, conferences split into
+two locations on different continents, etc.).
 
-This leads to a puzzle: How do we weigh the environmental harm caused by
-conference travel against the social benefits of the scientific progress
-that comes from conferences?  It seems an extremely difficult question to
-answer: the harm and the compensating benefits cannot be directly
-compared---they are apples and oranges.
+\SECTION{The Challenge of CO2 Reduction}
+
+Those we spoke to were almost universally supportive of reducing
+\COtwoE{} emissions due to conferences, but very few were eager
+to change the way we organize them. This
+hesitation makes sense. Science
+is a fundamentally social process, and the conference system
+accelerates research through high-bandwidth interaction, direct
+dissemination of results, network building, and serendipitous
+cross-fertilization. How to weigh the benefit to the environment of a
+particular \COtwoE-reducing change, such as on-line PC meetings,
+against the harm the change could cause to the scientific process,
+such as the lost opportunity to build networks and germinate
+synergistic ideas?
+
+Scientific value and environmental harm are apples and oranges, so
+they are hard to compare. They are also very hard to measure. While we
+can reasonably estimate the \COtwoE{} saved by a particular change, it
+is harder to guess the impact of the example it sets. This impact is
+important. Emissions from air travel to conferences are the lion's
+share of ACM's \COtwoE{} footprint, but air travel only contributes 
+2\% of current global emissions~\cite{emissions}. This small fraction is not an excuse
+for inaction. Just because any individual's or
+organization's share of global emissions is miniscule does not mean it
+doesn't matter. We must all contribute to the global 40\% reduction
+needed by 2030 in a way that makes the most difference, and we must set an example for
+others to do the same. It is especially important for scientists---who
+as a community have been raising the alarm about climate change---to
+not appear to be declaring themselves exempt from cutting back. But
+they must do so in a smart way, or else risk disproportionately
+harming the progress that science can bring.
 
@@
 reason, many policy experts advocate addressing climate change through some
-form of carbon pricing, which places a monetary cost on greenhouse gas
+form of carbon tax, which places a monetary cost on greenhouse gas
 emissions, sending a clear market signal that changes aggregate
@@
   for lots of good phrasing and citations.}  Exactly how the price of carbon
-should be set is a challenging policy question, but it has a clear goal: It
+should be set, and what should be done with the raised funds, is a
+challenging policy question, but it has a clear goal: It 
 should be chosen so that the world arrives at zero emissions within the
@@
 
-To be effective in the long term, a carbon price needs to be uniform across
-society---i.e., it should be imposed by governments.  Indeed, many
-countries\bcp{e.g.?} and states (e.g., California) have already adopted
-carbon pricing policies; more will follow.
+While a global \COtwoE{} market is the ultimate goal, we do not need
+to get there in one fell swoop.
+Some countries\bcp{e.g.?}, US states (e.g., California), universities
+(e.g., UMD and UCLA), and companies (\mwh{Jane Street?}) have
+begun to adopt internal carbon pricing policies and markets, to push
+their own behavior closer to a sustainable equilibrium. We feel that ACM
+should do the same thing. Here's how it might start.
+
+\SECTION{A Carbon Market within ACM}
+
+The main idea is to for ACM to enforce a carbon tax on conference
+budgets. At present, ACM charges an ``allocation'' percentage of
+12\% on total revenue. It can also impose a tax on
+the \COtwoE{} emissions induced by the conference. This tax has two
+benefits. First, it creates an incentive for SIGs and conference organizers to
+lower total costs by reducing emissions, so they can pay less
+tax. They could do this in many ways, such as the ones outlined in our 
+report. Second, ACM can use the extra revenue for ``good works.'' Such
+good works could directly offset the carbon that paid for them,
+or they could provide some other benefit, such as defraying open access
+publishing costs.
+
+Getting this idea rolling requires three things: (1) processes to
+measure \COtwoE{} used by a conference, for assessing the tax; (2) setting
+a tax to the conference budget for the \COtwoE{} it consumed; and (3)
+allocating and disbursing collected funds to good works.
+
+For the first, we can start by estimating \COtwoE{} emissions due to
+conference travel by analyzing the conference's registration data. We
+worked the ACM SGB to develop a \COtwoE{} calculator that takes such
+registration data and estimates the \COtwoE{} cost borne by each
+individual's travel to the conference. It could become part of
+standard practice to run this calculator when closing out the
+conference, to assess the tax. \mwh{other \COtwoE{} costs?}
 %
-Bottom line: We are entering an era in which carbon will be expensive.
-
-Anticipating this era, what should we rationally do now to ensure that ACM,
-and computer-science research more broadly, continue to thrive in a world
-where air travel is more expensive?  Believing that it needs to come quickly
-to avert disaster, what can we do to accelerate the process? \mwh{The
-  question now seems to have changed. The paper started by saying ACM has an
-  ``especially important role to play,'' but now seems to suggest it should
-  be reacting to an inevitable/coming future. These questions probably need
-  to go much earlier, and the earlier text reworked around them.}
-
-\SECTION{Carbon Offsets}
-
-In the short term, one possible response to both questions is to use {\em
-  carbon offsets}, which allow organizations (and individuals) to
-voluntarily impose a price on their own emissions~\cite{CarbonOFfsetReport}.
-%
-In a nutshell, the idea is for a conference to raise funds from
-registration fees or corporate sponsors and donate them, via a carbon offset
-{\em vendor}, to a {\em provider} who will use them to support an activity
-(e.g., installing methane capture devices on landfill sites or buying
-fuel-efficient stoves to replace open-fire cooking in poor communities) that
-is {\em certified} (by yet another organization) to permanently avoid
-emitting an amount of greenhouse gases equivalent to the emissions from
-participants' travel to the conference.
-%
-Many organizations (including companies such as Google, Dell, Microsoft,
-General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
-and a number of academic societies) are already buying carbon offsets, and
-some ACM conferences have already started to experiment with voluntary
-carbon offset fees at registration time.
-
-Carbon offsets do come with some significant concerns.
-%
-One superficial impediment is that most organizations do not currently have
-accounting procedures in place to deal with them, and the piecemeal
-political and bureaucratic work of designing and implementing a separate
-policy for each organization can consume significant energy.
-%
-Another point to be careful of is that the long-term climate benefit of some
-offsetting strategies (for example, paying people to plant trees) is
-controversial.
-%
-A more fundamental problem is that it is not straightforward to decide how
-much one should pay to offset a ton of carbon: current prices vary widely
-around an average of around \$15 per ton, depending on the vendor, the
-nature of the offset project, the region of the world where it will be
-implemented, etc.  Moreover, in order for offsets (or any other form of a
-carbon pricing) to fully motivate the needed changes in behavior, the price
-may need to go much higher---the true ``social cost'' of a ton of carbon is
-variously estimated at anywhere from \$40 to \$400 per
-ton~\bcp{citation}---and setting it too low may encourage complacency.
-%
-However, while acknowledging these concerns, we believe that voluntary
-carbon offsetting can be an effective near-term tool for raising awareness
-and beginning to shift organizational behavior because it associates
-emissions with a number, giving decision makers within the organization
-something concrete to optimize.
-
-\SECTION{Measure, Analyze, Optimize}
+Moreover, intrepid SIG leaders and conference organizers can analyze
+the gathered data to look for opportunities for reduction. For
+example, we used the calculator to compute the per-capita and total
+\COtwoE{} costs of several SIGPLAN conferences over the last few years.
+\mwh{Discuss.}
+\bcp{Put the discussion of Figure 1 (and an accompanying figure
+showing distance traveled to several instances of the same conference) here;
+this illustrates the difficulty and the need for careful data gathering and
+analysis.}
 
@@
 
-Beyond imposing a price on our carbon footprint (or waiting for one
-to be imposed from outside), a sane response to climate change should
-clearly also include taking steps to reduce the size of this
-footprint.
-\mwh{Don't like previous sentence. It seems to have gotten away from
-  the point of the paper, which is using price to change balance of
-  apples to oranges. Analyzing data is useful for figuring out ACM's
-  internal price, not tweaking around the edges, which is how it reads
-  below.} For
-conferences, there are many options that one might
-contemplate~\cite{ClimateCommitteReport}, from easy 
-but incremental changes like making program committee meetings virtual
-instead of physical and enabling virtual participation via live-streaming,
-all the way to complete re-imagining of conferences and their role in
-science (holding conferences completely in virtual spaces, replacing
-conferences with journals, etc.), with some intriguing hybrid forms in the
-middle (regional conferences, mega-conferences, conferences split into
-two locations on different continents, etc.).  
+How should we set the price of a conference's \COtwoE? There are several
+possibilities. One idea is simply to adopt the price from an existing
+market, such as California's; between 2012 and 2018, the price of a
+ton of \COtwoE{} fluctuated from about \$12 to \$23~\cite{CACO2}. Another idea is to
+match the price to the equivalent in \emph{carbon
+  offsets}~\cite{CarbonOFfsetReport}. Offset providers use raised funds
+to support an activity (e.g., installing methane capture devices on
+landfill sites or buying fuel-efficient stoves to replace open-fire
+cooking in poor communities) that is {\em certified} (by yet another
+organization) to permanently avoid emitting an amount of greenhouse
+gases specified. Current prices vary widely around an average of around
+\$15 per ton, depending on the vendor, the nature of the offset
+project, the region of the world where it will be implemented, etc.
+
+However we estimate the price, we could imagine setting it a little low at the
+outset, perhaps based on analysis of existing conference \COtwoE{} usage, and then raising it
+over time as organizers start to adapt to it. On the other hand, the true
+``social cost'' of a ton of carbon is variously estimated to be much
+higher than the above---anywhere from \$40 to \$400 per
+ton~\bcp{citation}---so current pricing may already be on the low
+side.
+
+Lastly, having collected the carbon tax from the conference budget,
+what should ACM do with it? One obvious thing to do is to buy \COtwoE{}
+offsets.\footnote{Some ACM conferences have already started to
+  experiment with voluntary carbon offset fees at registration time,
+  but with little luck.}
 %
-However, before we undertake large changes, we need to understand the
-situation better.  For most of these ideas, it is hard to confidently
-predict either the potential benefits (in terms of reduced emissions) or the
-potential costs (to scientific exchange, career development of young
-scientists, etc.).  Indeed, a preliminary analysis of registration data from
-a number of SIGPLAN conferences paints a rather complicated picture.
+Carbon offsets are not a long-term solution to global warming, and the
+details matter, but at this point we believe there are options that do
+far more good than harm.
 %
-\bcp{Put the discussion of Figure 1 (and an accompanying figure
-showing distance traveled to several instances of the same conference) here;
-this illustrates the difficulty and the need for careful data gathering and
-analysis.} 
-All parts of ACM should be doing this kind of analysis.  (The acm-climate
-mailing list is one place to exchange ideas about this.)
+Many organizations (including companies such as Google, Dell, Microsoft,
+General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
+and a number of academic societies) are already buying carbon offsets.
+%
+One superficial impediment is that most organizations do not currently have
+accounting procedures in place to deal with them, and the piecemeal
+political and bureaucratic work of designing and implementing a separate
+policy for each organization can consume significant energy. \mwh{Not
+  sure the previous is the right thing to say? Basically: Need to work
+  out the legality. Cite offset report?}
 
-\SECTION{Deeper Changes}
+Another thing is to do \mwh{finish off. Here, talk about other good
+  works, like open access fees, developing technology for CO2e
+  reduction of hosting conferences, perhaps pay for remote presence,
+  ...}
 
-Ultimately, responding to the inevitability of expensive carbon may involve
-rethinking the whole basis of how academic computer science is done.
-\bcp{
-\begin{itemize}
-\item Add a bit of meat about how the current economic model may be
-  unsustainable and reworking it will involve substantial organizational
-  changes.  (There may be some good words that we can paraphrase from our
-  second reviewer!)
-\item monitoring and mitigating carbon costs should be an ACM-level activity
-\end{itemize}
-}
+\SECTION{Conclusion}
 

### 2019-06-10 mike.tex
Comments on Mike proposal, and some polishing on it

@@
 % (Figure \ref{fig:warming}).
-This is not a problem to be left to future
-generations: we are up against it now. The UN's Intergovernmental
+This is not a problem we can leave to future
+generations: The UN's Intergovernmental
 Panel on Climate Change (IPCC) says a 40\% decrease in
@@
 require urgent and sustained commitment at all levels of society, including
-large organizations like universities, companies, and scientific societies
+not only city, state, and national governments, but also 
+universities, companies, and scientific societies
 like ACM.
 
-Indeed, scientific societies arguably have an especially important role to
+Indeed, scientific societies have an especially important role to
 play, since one of their main activities is organizing international
@@
 Moreover, these emissions have no near-term 
-technological fix, since jet fuel is difficult to replace with
+technological fix, jet fuel being difficult to replace with
 renewable energy sources~\cite{elec-air}. 
@@
 
-We formed the ACM SIGPLAN Climate Committee with this situation in
-mind. We made presentations at SIGPLAN events, had discussions with
-ACM and SIG leaders, and ultimately put together a report of ideas for
+Motivated by this situation, in 2016 we formed the ACM SIGPLAN Climate
+Committee. Since then, we made presentations at SIGPLAN events, spoken with
+ACM and SIG leaders, and assembled a report assessing a number of ideas for
 reducing the \COtwoE{} costs of conference
-activities~\cite{ClimateCommitteReport}. The report considers easy 
-but incremental changes like making program committee meetings virtual
-instead of physical and enabling virtual participation via live-streaming,
-all the way to complete re-imagining of conferences and their role in
-science (holding conferences completely in virtual spaces, replacing
-conferences with journals, etc.), with some intriguing hybrid forms in the
-middle (regional conferences, mega-conferences, conferences split into
-two locations on different continents, etc.).
+activities~\cite{ClimateCommitteReport}.  Our report considers a range of
+options, fromn easy-but-incremental changes like making program committee
+meetings virtual instead of physical and enabling virtual participation in
+conferences via live-streaming, all the way to complete
+re-conceptualizations of conferences and their role in science (e.g.,
+replacing conferences with journals, holding conferences completely in
+virtual spaces, etc.), as well as some intriguing points in the middle of
+the spectrum (regional conferences, mega-conferences, conferences held
+simultaneously at two locations on different continents, etc.).
 
@@
 
-Those we spoke to were almost universally supportive of reducing
-\COtwoE{} emissions due to conferences, but very few were eager
-to change the way we organize them. This
-hesitation makes sense. Science
-is a fundamentally social process, and the conference system
-accelerates research through high-bandwidth interaction, direct
+Those we spoke to have been almost universally supportive of reducing
+\COtwoE{} emissions due to conferences, but very few are eager to
+fundamentally change the way conferences are organized. This hesitation
+makes sense. Science is a fundamentally social process, and the conference
+system accelerates research through high-bandwidth interaction, direct
 dissemination of results, network building, and serendipitous
-cross-fertilization. How to weigh the benefit to the environment of a
-particular \COtwoE-reducing change, such as on-line PC meetings,
-against the harm the change could cause to the scientific process,
-such as the lost opportunity to build networks and germinate
-synergistic ideas?
+cross-fertilization.  So how can we weight the ennvironmental benefit of a
+particular \COtwoE-reducing change, such as on-line PC meetings, against the
+harm the change could cause to the scientific process, such as the lost
+opportunity to build networks and germinate synergistic ideas?
 
-Scientific value and environmental harm are apples and oranges, so
-they are hard to compare. They are also very hard to measure. While we
-can reasonably estimate the \COtwoE{} saved by a particular change, it
-is harder to guess the impact of the example it sets. This impact is
-important. Emissions from air travel to conferences are the lion's
-share of ACM's \COtwoE{} footprint, but air travel only contributes 
-2\% of current global emissions~\cite{emissions}. This small fraction is not an excuse
-for inaction. Just because any individual's or
+Scientific value and environmental harm are apples and oranges---hard to
+compare.  Indeed, they are even hard to measure. In particular, while we can
+reasonably estimate the \COtwoE{} directly saved by a particular change, it
+is much harder to assess the impact of the example it sets.
+
+This impact is important---perhaps more important than \COtwoE{} reductions
+themselves. Emissions from air travel to conferences are the lion's share of
+ACM's \COtwoE{} footprint, but air travel only contributes 2\% of current
+global emissions~\cite{emissions}. But this small fraction should not be
+taken as an excuse for inaction: the fact that any individual's or
 organization's share of global emissions is miniscule does not mean it
-doesn't matter. We must all contribute to the global 40\% reduction
-needed by 2030 in a way that makes the most difference, and we must set an example for
-others to do the same. It is especially important for scientists---who
-as a community have been raising the alarm about climate change---to
-not appear to be declaring themselves exempt from cutting back. But
-they must do so in a smart way, or else risk disproportionately
-harming the progress that science can bring.
+doesn't matter. {\em No} individual's (or organization's) contributions
+amount to a nontrivial percentage of global emissions; {\em everyone} will
+need to contribute, to achieve the global 40\% reduction needed by 2030. To
+get others on board, we must set an example. Indeed, it is especially
+important for scientists---who, as a community, have been instrumental in
+raising the alarm about climate change---to not appear to be exempting
+themselves from cutting back.  But they must do so in a smart way, or else
+risk disproportionately harming the progress that science can bring.
 
@@
 magic of markets.''  By mapping choices and preferences in disparate domains
-onto a common scale (money), economic markets can be viewed as massively
-distributed computations that optimize costs and benefits across vastly
-different activities.\bcp{Give a simple, concrete example.}  For this
+onto a common scale (i.e., money), economic markets function as
+massively distributed computations that optimize costs and benefits across
+vastly different activities.\bcp{Give a simple, concrete example.}  For this
 reason, many policy experts advocate addressing climate change through some
-form of carbon tax, which places a monetary cost on greenhouse gas
+form of carbon pricing, which places a monetary cost on greenhouse gas
 emissions, sending a clear market signal that changes aggregate
@@
 should be set, and what should be done with the raised funds, is a
-challenging policy question, but it has a clear goal: It 
-should be chosen so that the world arrives at zero emissions within the
-timeframe that climate scientists tell us is safe.
+challenging policy question, but it has a clear goal: It should be chosen so
+that the world arrives at zero emissions within the timeframe that climate
+scientists tell us is safe.
 
@@
 Some countries\bcp{e.g.?}, US states (e.g., California), universities
-(e.g., UMD and UCLA), and companies (\mwh{Jane Street?}) have
+(e.g., UMD and UCLA), and companies (\mwh{Jane Street?}) have already
 begun to adopt internal carbon pricing policies and markets, to push
-their own behavior closer to a sustainable equilibrium. We feel that ACM
+their own behavior closer to a sustainable equilibrium.  We argue that ACM
 should do the same thing. Here's how it might start.
@@
 or they could provide some other benefit, such as defraying open access
-publishing costs.
+publishing costs.  \bcp{I foresee a tricky legal question of whether it's OK for
+  ACM to charge such a tax simply for ``good works.''}
 
@@
 conference travel by analyzing the conference's registration data. We
-worked the ACM SGB to develop a \COtwoE{} calculator that takes such
+worked with the ACM SGB to develop a \COtwoE{} calculator that takes such
 registration data and estimates the \COtwoE{} cost borne by each
@@
 standard practice to run this calculator when closing out the
-conference, to assess the tax. \mwh{other \COtwoE{} costs?}
+conference, to assess the tax. \mwh{other \COtwoE{} costs?}  \bcp{But the
+  conference planners would need at least an estimate of what the tax was
+  going to amount to, to avoid a budget shortfall.}
 %
@@
 project, the region of the world where it will be implemented, etc.
-
-However we estimate the price, we could imagine setting it a little low at the
-outset, perhaps based on analysis of existing conference \COtwoE{} usage, and then raising it
-over time as organizers start to adapt to it. On the other hand, the true
+On the other hand, the true
 ``social cost'' of a ton of carbon is variously estimated to be much
@@
 side.
+
+However we estimate the price, we could imagine setting it a little low at the
+outset, perhaps based on analysis of existing conference \COtwoE{} usage, and then raising it
+over time to give organizers time to adapt.
 
@@
 The climate crisis is too large and too urgent to leave to world leaders to
-address at their own pace: organizations at every scale should confront
+address at their own pace: organizations at every scale must confront
 their own contributions, raise awareness among their members, and begin
@@
 coming.
+\bcp{A sentence restating our proposal.}
+
 % For ACM, this especially means confronting the carbon-intensive

### 2019-06-10 main.tex
Nits
Some work on the original draft

@@
 % \begin{multicols}{3}
-
-ACM's vibrant conference culture has a dark side: it contributes
-significantly to climate change.
 
@@
 
-A broad scientific consensus warns us that human
+A broad scientific consensus warns that human
 emissions of carbon dioxide and other greenhouse gases are warming the earth.
 % (Figure \ref{fig:warming}).
-This is not a problem to be left to future
-generations: we are up against it now. The UN's Intergovernmental
+This is not a problem we can leave to future
+generations: The UN's Intergovernmental
 Panel on Climate Change (IPCC) says a 40\% decrease in
@@
 require urgent and sustained commitment at all levels of society, including
-large organizations like universities, companies, and scientific societies
+not only city, state, and national governments, but also 
+universities, companies, and scientific societies
 like ACM.
 
+% A broad scientific consensus warns us that human
+% emissions of carbon dioxide and other greenhouse gases are warming the earth.
+% % (Figure \ref{fig:warming}).
+% This is not a problem to be left to future
+% generations: we are up against it now. The UN's Intergovernmental
+% Panel on Climate Change (IPCC) says a 40\% decrease in
+% emissions by 2030 is needed to avoid irreversible damage~\cite{IPCC-report}.
+% % (Figure~\ref{fig:IPCC-cover}).
+% Achieving reductions on this scale will
+% require urgent and sustained commitment at all levels of society, including
+% large organizations like universities, companies, and scientific societies
+% like ACM.
+
 Indeed, scientific societies arguably have an especially important role to
-play, since one of their main activities is organizing international
-conferences, which produce significant greenhouse gas emissions,
-particularly from air travel. \mwh{The previous is the wrong
-  transition, since again it seems to suggest that travel is a big
-  deal in terms of the overall problem. The sentence also fails to
-  convey the other reason scientific societies should be involved: As
-  role models and advocates for general change.}  A single round-trip flight from
-Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon
-dioxide (\COtwoE) per passenger~\cite{carboncalculator}\iflater\bcp{Trim
-  citation?}\fi---a substantial fraction of the {\em total} yearly emissions
-for an average resident of the US (16.5 tons) or Europe (7
-tons)~\cite{emissions}.  \iflater\bcp{For comparison, ...  (put in some
-  figures about the footprints of a few other things that people might think
-  of reducing---driving, eating meat, upgrading their fridge...)}\fi{}
-Moreover, these emissions have no near-term 
-technological fix, since jet fuel is difficult to replace with
-renewable energy sources~\cite{elec-air}. 
+play, since, for many of their members, travel to conferences may be a
+substantial or even dominant part of their individual contribution to
+climate change.  A single round-trip flight from Philadelphia to Paris emits
+the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
+passenger~\cite{carboncalculator}\iflater\bcp{Trim citation?}\fi---a
+significant fraction of the {\em total} yearly emissions for an average
+resident of the US (16.5 tons) or Europe (7 tons)~\cite{emissions}.
+\iflater\bcp{For comparison, ...  (put in some figures about the footprints
+  of a few other things that people might think of reducing---driving,
+  eating meat, upgrading their fridge...)}\fi{} Moreover, these emissions
+have no near-term technological fix, since jet fuel is difficult to replace
+with renewable energy sources~\cite{elec-air}.
 % \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
@@
 
-Conferences obviously also create significant value.  Science is a
-fundamentally social process, and the conference system accelerates research
-through high-bandwidth interaction, direct dissemination of results, network
-building, and serendipitous cross-fertilization.
+On the other hand, conferences obviously also create significant value.
+Science is a fundamentally social process, and the conference system
+accelerates research through high-bandwidth interaction, direct
+dissemination of results, network building, and serendipitous
+cross-fertilization.
 
-This leads to a puzzle: How do we weigh the environmental harm caused by
+This presents a puzzle: How do we weigh the environmental harm caused by
 conference travel against the social benefits of the scientific progress
@@
 magic of markets.''  By mapping choices and preferences in disparate domains
-onto a common scale (money), economic markets can be viewed as massively
+onto a common scale (i.e., money), economic markets function as massively
 distributed computations that optimize costs and benefits across vastly
@@
   to go much earlier, and the earlier text reworked around them.}
+\bcp{I tried rewording the opening of the second paragraph to address this.}
 
@@
 
-In the short term, one possible response to both questions is to use {\em
-  carbon offsets}, which allow organizations (and individuals) to
-voluntarily impose a price on their own emissions~\cite{CarbonOFfsetReport}.
+One possible short-term response to both questions is to use {\em
+  carbon offsets} to voluntarily impose a price on our own
+emissions~\cite{CarbonOFfsetReport}.
 %
-In a nutshell, the idea is for a conference to raise funds from
-registration fees or corporate sponsors and donate them, via a carbon offset
-{\em vendor}, to a {\em provider} who will use them to support an activity
+In a nutshell, the idea is that a conference raise funds from
+registration fees or corporate sponsors and donates them, via a carbon offset
+{\em vendor}, to a {\em provider} who  uses them to support an activity
 (e.g., installing methane capture devices on landfill sites or buying
@@
 %
-A more fundamental problem is that it is not straightforward to decide how
-much one should pay to offset a ton of carbon: current prices vary widely
+A more fundamental problem is that it is hard to decide how
+much one should pay to offset a ton of carbon: current prices vary
+significantly 
 around an average of around \$15 per ton, depending on the vendor, the
@@
 
-Beyond imposing a price on our carbon footprint (or waiting for one
-to be imposed from outside), a sane response to climate change should
-clearly also include taking steps to reduce the size of this
-footprint.
-\mwh{Don't like previous sentence. It seems to have gotten away from
-  the point of the paper, which is using price to change balance of
-  apples to oranges. Analyzing data is useful for figuring out ACM's
-  internal price, not tweaking around the edges, which is how it reads
-  below.} For
-conferences, there are many options that one might
-contemplate~\cite{ClimateCommitteReport}, from easy 
-but incremental changes like making program committee meetings virtual
-instead of physical and enabling virtual participation via live-streaming,
-all the way to complete re-imagining of conferences and their role in
-science (holding conferences completely in virtual spaces, replacing
-conferences with journals, etc.), with some intriguing hybrid forms in the
-middle (regional conferences, mega-conferences, conferences split into
-two locations on different continents, etc.).  
+Discussions about the future of conferences also require data.  
 %
-However, before we undertake large changes, we need to understand the
+% Beyond imposing a price on our carbon footprint (or waiting for one
+% to be imposed from outside), a sane response to climate change should
+% clearly also include taking steps to reduce the size of this
+% footprint.
+% \mwh{Don't like previous sentence. It seems to have gotten away from
+%   the point of the paper, which is using price to change balance of
+%   apples to oranges. Analyzing data is useful for figuring out ACM's
+%   internal price, not tweaking around the edges, which is how it reads
+%   below.} 
+\bcp{Bad transition here.  The rest of this section needs reorg.}
+For conferences, there are many options that one might
+contemplate~\cite{ClimateCommitteReport}, from easy-but-incremental changes
+like making program committee meetings virtual instead of physical and
+enabling virtual participation via live-streaming, all the way to complete
+re-imagining of conferences and their role in science (holding conferences
+completely in virtual spaces, replacing conferences with journals, etc.),
+with some intriguing hybrid forms in the middle (regional conferences,
+mega-conferences, conferences held simultaneously at two locations on
+different continents, etc.).
+%
+But before undertaking significant changes, we need to understand the
 situation better.  For most of these ideas, it is hard to confidently

### 2019-06-11 main.tex
Fiddling

@@
 
-\SECTION{Putting a Price on Emissions}
+\SECTION{Putting a Price on Carbon}
 
@@
 To be effective in the long term, a carbon price needs to be uniform across
-society---i.e., it should be imposed by governments.  Indeed, many
-countries\bcp{e.g.?} and states (e.g., California) have already adopted
-carbon pricing policies; more will follow.
+society---i.e., it should be imposed by governments.  Indeed, some 40
+countries and more than 20 cities and states have already adopted
+carbon pricing policies~\cite{pricingcarbon}; more will follow.
 %
@@
 
-Anticipating this era, what should we rationally do now to ensure that ACM,
-and computer-science research more broadly, continue to thrive in a world
+Anticipating this era, what should we rationally do now to ensure that 
+computer-science research continues to thrive in a world
 where air travel is more expensive?  Believing that it needs to come quickly
@@
 
-\SECTION{Carbon Offsets}
+\SECTION{Starting Now}
 
@@
 
-Carbon offsets do come with some significant concerns.
+Carbon offsets also raise some significant concerns.
+% %
+% A superficial one is that most organizations do not currently have
+% accounting procedures in place to deal with them, and the piecemeal
+% political and bureaucratic work of designing and implementing a separate
+% policy for each organization can consume significant energy.
 %
-One superficial impediment is that most organizations do not currently have
-accounting procedures in place to deal with them, and the piecemeal
-political and bureaucratic work of designing and implementing a separate
-policy for each organization can consume significant energy.
-%
-Another point to be careful of is that the long-term climate benefit of some
-offsetting strategies (for example, paying people to plant trees) is
-controversial.
+One is that the long-term climate benefit of some offsetting strategies (for
+example, paying people to plant trees) is controversial: it is important to
+choose reputable vendors and projects (such as methane capture) whose
+net benefit is clear.
 %
@@
 much one should pay to offset a ton of carbon: current prices vary
-significantly 
 around an average of around \$15 per ton, depending on the vendor, the
@@
 %
-However, while acknowledging these concerns, we believe that voluntary
+While acknowledging these concerns, we believe that voluntary
 carbon offsetting can be an effective near-term tool for raising awareness
@@
 
-\SECTION{Measure, Analyze, Optimize}
+\SECTION{Grounding the Discussion}
 
@@
 
-Discussions about the future of conferences also require data.  
+What, concretely, would it mean to ``optimize'' the carbon footprint of
+conferences?
 %
@@
 %   below.} 
-\bcp{Bad transition here.  The rest of this section needs reorg.}
-For conferences, there are many options that one might
-contemplate~\cite{ClimateCommitteReport}, from easy-but-incremental changes
-like making program committee meetings virtual instead of physical and
-enabling virtual participation via live-streaming, all the way to complete
-re-imagining of conferences and their role in science (holding conferences
-completely in virtual spaces, replacing conferences with journals, etc.),
-with some intriguing hybrid forms in the middle (regional conferences,
-mega-conferences, conferences held simultaneously at two locations on
-different continents, etc.).
+One might contemplate a wide array of options~\cite{ClimateCommitteReport},
+from easy-but-incremental changes like making program committee meetings
+virtual instead of physical and enabling virtual participation via
+live-streaming, all the way to complete re-imagining of conferences and
+their role in science (holding conferences completely in virtual spaces,
+replacing conferences with journals, etc.), with some intriguing hybrid
+forms in the middle (regional conferences, mega-conferences, conferences
+held simultaneously at two locations on different continents, etc.).
 %
-But before undertaking significant changes, we need to understand the
-situation better.  For most of these ideas, it is hard to confidently
-predict either the potential benefits (in terms of reduced emissions) or the
-potential costs (to scientific exchange, career development of young
-scientists, etc.).  Indeed, a preliminary analysis of registration data from
+How much of a difference would these make to emissions?
+
+% But before undertaking significant changes, we need to understand the
+% situation better.  For most of these ideas, it is hard to confidently
+% predict either the potential benefits (in terms of reduced emissions) or the
+% potential costs (to scientific exchange, career development of young
+% scientists, etc.).  
+A preliminary analysis of registration data from
 a number of SIGPLAN conferences paints a rather complicated picture.
@@
 analysis.}
+
 All parts of ACM should be doing this kind of analysis.  (The acm-climate
@@
 
-\SECTION{Deeper Changes}
+\SECTION{Plan Ahead}
 
-Ultimately, responding to the inevitability of expensive carbon may involve
-rethinking the whole basis of how academic computer science is done.
+Looking further ahead, the expectation of expensive carbon invites a deeper
+rethinking of how academic computer science is done.  
+%
 \bcp{

### 2019-06-17 main.tex
CP

@@
 % \end{figure}
+
+\begin{itemize}
+\item Accountability.
+  \begin{itemize}
+  \item Proposal: Every ACM-sponsored conference should be required to
+  publicly report its carbon footprint (total and per-participant).
+  (Including PC meetings, participant travel, maybe hotel?).
+  \item ACM needs to provide tools to help conferences do this.
+  \item Benefits:
+  \begin{itemize}
+  \item Encourages people to pay attention, which will lead to some changes
+  in behavior all by itself. 
+  \item Generates data that will be needed for grounding discussions of how
+  optimize footprint.
+  \item Indeed, optimizing footprint is not so easy.  (Describe some of our
+  data.)  More data and thorough analysis is needed.
+  \end{itemize}
+  \end{itemize}
+\item Pricing carbon.
+  \begin{itemize}
+  \item To motivate more painful changes and force people to make decisions,
+  we need a price on carbon.
+  \item Ultimately, we want/expect governments to do this.  But we don't need to
+  wait. 
+  \item Proposal: ACM should apply downward pressure on conferences by
+  charging conferences a surcharge based on their carbon footprint.
+  \item This money can be used in many ways: buying carbon offsets,
+  supporting ``green research'', supporting virtualizing conferences
+  (livestreaming, etc.), supporting a transition to a low-carbon / few
+  conferences business model for ACM.
+  \end{itemize}
+\end{itemize}
 

### 2019-06-17 main.tex
CP
New outline

@@
 
-\begin{itemize}
-\item Accountability.
-  \begin{itemize}
-  \item Proposal: Every ACM-sponsored conference should be required to
-  publicly report its carbon footprint (total and per-participant).
-  (Including PC meetings, participant travel, maybe hotel?).
-  \item ACM needs to provide tools to help conferences do this.
-  \item Benefits:
-  \begin{itemize}
-  \item Encourages people to pay attention, which will lead to some changes
-  in behavior all by itself. 
-  \item Generates data that will be needed for grounding discussions of how
-  optimize footprint.
-  \item Indeed, optimizing footprint is not so easy.  (Describe some of our
-  data.)  More data and thorough analysis is needed.
-  \end{itemize}
-  \end{itemize}
-\item Pricing carbon.
-  \begin{itemize}
-  \item To motivate more painful changes and force people to make decisions,
-  we need a price on carbon.
-  \item Ultimately, we want/expect governments to do this.  But we don't need to
-  wait. 
-  \item Proposal: ACM should apply downward pressure on conferences by
-  charging conferences a surcharge based on their carbon footprint.
-  \item This money can be used in many ways: buying carbon offsets,
-  supporting ``green research'', supporting virtualizing conferences
-  (livestreaming, etc.), supporting a transition to a low-carbon / few
-  conferences business model for ACM.
-  \end{itemize}
-\end{itemize}
-
 A broad scientific consensus warns that human
@@
 % (Figure \ref{fig:warming}).
-This is not a problem we can leave to future
+This is not a problem that can be left to future
 generations: The UN's Intergovernmental
 Panel on Climate Change (IPCC) says a 40\% decrease in
-emissions by 2030 is needed to avoid irreversible damage~\cite{IPCC-report}.
+emissions is needed by 2030 to avoid irreversible damage~\cite{IPCC-report}.
 % (Figure~\ref{fig:IPCC-cover}).
@@
 
-Indeed, scientific societies arguably have an especially important role to
+In fact, scientific societies arguably have an especially important role to
 play, since, for many of their members, travel to conferences may be a
@@
 
-On the other hand, conferences obviously also create significant value.
-Science is a fundamentally social process, and the conference system
-accelerates research through high-bandwidth interaction, direct
-dissemination of results, network building, and serendipitous
-cross-fertilization.
+How should ACM respond to these facts?  We have two concrete proposals.
 
-This presents a puzzle: How do we weigh the environmental harm caused by
+\SECTION{Mandate public accounting of carbon footprints.}  
+\bcp{The section headers could use some wordsmithing!}
+
+  \begin{itemize}
+  \item Proposal: Every ACM-sponsored conference should be required to
+  publicly report its carbon footprint (total and per-participant).
+  (Including PC meetings, participant travel, maybe hotel?).
+  \item ACM needs to provide tools to help conferences do this.  (SIGPLAN
+  built one.)
+  \item Benefits:
+  \begin{itemize}
+  \item Encourages people to pay attention, which will lead to some changes
+  in behavior all by itself. 
+  \item Generates data that will be needed for grounding discussions of how
+  optimize footprint.  E.g., SIGPLAN has mostly gone to virtual PC
+  meetings.  Livestreaming.
+  \end{itemize}
+\end{itemize}
+
+These are things that we can and should be doing now.  Indeed, many parts
+of ACM are already doing them.  (In SIGPLAN, we have... and the experience
+has been...)
+
+However, going beyond these simple steps is going to be much more
+challenging.  One reason is that it isn't actually very clear what changes
+will help significantly.  (Describe some of our data.)  More data and
+thorough analysis is needed.
+
+Another reason why going beyond simple measures is hard is that, although
+conferences damage the climate, they also create significant value.  Science
+is a fundamentally social process, and the conference system accelerates
+research through high-bandwidth interaction, direct dissemination of
+results, network building, and serendipitous cross-fertilization.
+
+This presents a puzzle: How can we weigh the environmental harm caused by
 conference travel against the social benefits of the scientific progress
-that comes from conferences?  It seems an extremely difficult question to
-answer: the harm and the compensating benefits cannot be directly
-compared---they are apples and oranges.
-
-\SECTION{Putting a Price on Carbon}
+that comes from conferences?  The question is difficult because the harm and
+the compensating benefits cannot be directly compared---they are apples and
+oranges.
 
@@
   https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
-  for lots of good phrasing and citations.}  Exactly how the price of carbon
+  for lots of good phrasing and citations.}
+
+\SECTION{Put a Price on Carbon.}
+
+  \begin{itemize}
+  \item To motivate more painful changes and force people to make difficult
+  choices, we need a price on carbon.
+  \item Ultimately, governments should do this.  But we don't need to wait:
+  as several companies have shown, even an artificial ``internal price'' on
+  carbon creates a significant price signal that shapes behavior.
+  \item Our second (more ambitious) proposal is: ACM should begin requiring
+  conferences to pay a surcharge based on their carbon footprint.  It should
+  be low at the beginning and increase steadily (and predictably) year on year.
+  \item Benefit: This will create concrete downward pressure on conferences
+  to decrease their footprints.  They can then choose (on the basis of
+  careful analysis of data and weighing of benefits) how best to do so.
+  \item The money raised in this way can be used in many ways: buying carbon
+  offsets, supporting ``green computing research'', virtualizing conferences
+  (livestreaming, etc.), supporting a transition to a low-carbon / few
+  conferences business model for ACM.
+  \end{itemize}
+
+(Some kind of closing.)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+
+\section*{Old text (that may be useful to mine)}
+
+\SECTION{Putting a Price on Carbon}
+
+Exactly how the price of carbon
 should be set is a challenging policy question, but it has a clear goal: It
@@
 % \end{itemize}
+
+\bcp{New outline:
+\begin{itemize}
+\item Accountability.
+  \begin{itemize}
+  \item Proposal: Every ACM-sponsored conference should be required to
+  publicly report its carbon footprint (total and per-participant).
+  (Including PC meetings, participant travel, maybe hotel?).
+  \item ACM needs to provide tools to help conferences do this.  (SIGPLAN
+  built one.)
+  \item Benefits:
+  \begin{itemize}
+  \item Encourages people to pay attention, which will lead to some changes
+  in behavior all by itself. 
+  \item Generates data that will be needed for grounding discussions of how
+  optimize footprint.  E.g., SIGPLAN has mostly gone to virtual PC
+  meetings.  Livestreaming.
+  \item Indeed, optimizing footprint is not so easy.  (Describe some of our
+  data.)  More data and thorough analysis is needed.
+  \item Segue: Colocation, etc.
+  \end{itemize}
+  \end{itemize}
+\item Pricing carbon.
+  \begin{itemize}
+  \item To motivate more painful changes and force people to make decisions,
+  we need a price on carbon.
+  \item Ultimately, we want/expect governments to do this.  But we don't need to
+  wait. 
+  \item Proposal: ACM should apply downward pressure on conferences by
+  charging conferences a surcharge based on their carbon footprint.
+  \item This money can be used in many ways: buying carbon offsets,
+  supporting ``green research'', supporting virtualizing conferences
+  (livestreaming, etc.), supporting a transition to a low-carbon / few
+  conferences business model for ACM.
+  \end{itemize}
+\end{itemize}
+}
 

### 2019-06-17 main.tex
More

@@
 challenging.  One reason is that it isn't actually very clear what changes
-will help significantly.  (Describe some of our data.)  More data and
-thorough analysis is needed.
+will help significantly.  (Describe some of our data.  Show the two
+diagrams.  Draw some conclusions / discuss why it's hard to draw
+conclusions.)  More data and more thorough analysis is needed.
 
-Another reason why going beyond simple measures is hard is that, although
-conferences damage the climate, they also create significant value.  Science
-is a fundamentally social process, and the conference system accelerates
-research through high-bandwidth interaction, direct dissemination of
-results, network building, and serendipitous cross-fertilization.
+But there is another, deep reason why going beyond simple measures is hard:
+while conferences damage the climate, they also create significant value!
+Science is a fundamentally social process, and the conference system
+accelerates research through high-bandwidth interaction, direct
+dissemination of results, network building, and serendipitous
+cross-fertilization.
 
@@
 
-Fortunately, there is a well-established mechanism for comparing apples and
+Fortunately, we have a well-established mechanism for comparing apples and
 oranges---indeed, for comparing almost anything!  Economists call it ``the

### 2019-06-18 main.tex
Work.

@@
 
-How should ACM respond to these facts?  We have two concrete proposals.
+How should ACM respond to these facts?  We have two concrete
+proposals. First, all ACM conferences should publicly account for CO2e
+spent as a result of hosting them. Second, ACM should embrace putting
+a \emph{price} on carbon in conference budgets, to create pressure on
+organizers to reduce its consumption.
 
@@
 
-  \begin{itemize}
-  \item Proposal: Every ACM-sponsored conference should be required to
-  publicly report its carbon footprint (total and per-participant).
-  (Including PC meetings, participant travel, maybe hotel?).
-  \item ACM needs to provide tools to help conferences do this.  (SIGPLAN
-  built one.)
-  \item Benefits:
-  \begin{itemize}
-  \item Encourages people to pay attention, which will lead to some changes
-  in behavior all by itself. 
-  \item Generates data that will be needed for grounding discussions of how
-  optimize footprint.  E.g., SIGPLAN has mostly gone to virtual PC
-  meetings.  Livestreaming.
-  \end{itemize}
-\end{itemize}
+Our first proposal is simple: Every ACM-sponsored conference should
+publicly report its carbon footprint, total and per-participant. A
+conference's footprint will be dominated by participants' air travel,
+but the accounting should not be limited to that: It should also
+include air travel to in-person PC meetings, and estimate hotel and
+food CO2e costs.
 
-These are things that we can and should be doing now.  Indeed, many parts
-of ACM are already doing them.  (In SIGPLAN, we have... and the experience
-has been...)
+We hope that a public accounting will nudge conference organizers and
+attendees to change their behavior. As an analogy: Chain restaurants
+in the US are required by law to post calorie counts of food items on
+menu boards, and studies show that enlightened customers order, on
+average, up to 50 fewer calories a day, as a
+result~\cite{menu}.
 
-However, going beyond these simple steps is going to be much more
-challenging.  One reason is that it isn't actually very clear what changes
-will help significantly.  (Describe some of our data.  Show the two
-diagrams.  Draw some conclusions / discuss why it's hard to draw
-conclusions.)  More data and more thorough analysis is needed.
+ACM can provide tools to help organizers account for this
+footprint. We recently worked with the ACM SGB to develop an air
+travel-focused CO2e calculator. One can input a CSV with registration
+data, and the calculator will compute the estimated CO2e cost of air
+travel for those sufficiently far away. This calculator could be
+expanded to support related costs.
 
-But there is another, deep reason why going beyond simple measures is hard:
-while conferences damage the climate, they also create significant value!
+SIGPLAN has been thinking about how to reduce CO2e usage over the past
+couple of years, informed by an accounting of CO2e expenditures. Doing
+so has led several conference organizers \mwh{name which conferences}
+to switch from holding in-person PC meetings to holding the meetings
+on-line instead. \mwh{Mention Jens survey here.} It has also prompted
+adding support for conference livestreaming (to go with talk
+recording) to permit remote participation.
+
+Most recently, we have been investigating how to purchase \emph{carbon
+  offsets} to defray CO2e costs.
+%
+A carbon offset is sold by a vendor who uses the raised funds to
+finance an activity that permanently avoids emitting an amount of
+greenhouse gases. That the CO2e emission avoidance is truth and not
+appearance can be certified by a watchdog organization. Typically,
+planting trees is not a certifiable activity (since it's hard to
+guarantee the trees won't be cut down), but installing methane capture
+devices on landfill sites or buying fuel-efficient stoves to replace
+open-fire cooking in poor communities can be.
+
+Many organizations (including companies such as Google, Dell,
+Microsoft, General Motors, Delta Airlines, Lyft, and Expedia, numerous
+universities, and a number of academic societies) now buy carbon
+offsets in order to reduce their overall footprint. ACM conferences
+could do the same, and the purchases would be included in their public
+accounting. We have been investigating this possibility for the past
+year. ACM has now agreed that conference attendees can opt to purchase
+an offset during registration, and ACM will use the designated funds
+to purchase an offset. Even getting to this stage required some
+investigation of the legality of ACM making the purchase on behalf of
+the registrant~\cite{CarbonOFfsetReport}. Ideally, the purchase of
+offsets would be made even simpler, but doing so requires further
+thought.
+
+\SECTION{Put a Price on Carbon.}
+
+Publicly accounting for CO2e expenditures will incentivize eliminating
+CO2e-consuming activities whose corresponding value to the conference
+is likely to be low. But a public accounting is unlikely to
+incentivize major CO2e reductions, since these will require
+significant changes to conference system itself, which may negatively
+impact the value that conferences provide. 
+%
 Science is a fundamentally social process, and the conference system
@@
 cross-fertilization.
+%
+Organizers and attendees will be reticent to harm the scientific
+benefits they have all regularly experienced.
 
-This presents a puzzle: How can we weigh the environmental harm caused by
-conference travel against the social benefits of the scientific progress
-that comes from conferences?  The question is difficult because the harm and
-the compensating benefits cannot be directly compared---they are apples and
-oranges.
+(Mike: I think we can significantly trim the text below. Instead, we can
+talk here about a carbon tax as the next natural step. It can be
+related to the sugar tax that some municipalities have imposed; when
+public calorie counts are not enough to discourage use, add a tax
+instead. The arc from the bullet list can come earlier, basically.)
+
+Thus we have the question of apples and oranges: How can we weigh the
+environmental harm caused by conference travel against the social
+benefits of the scientific progress that comes from conferences?
+
+% The question is difficult because the harm and
+% the compensating benefits cannot be easily compared---they are apples and
+% oranges.
 
@@
 
-\SECTION{Put a Price on Carbon.}
-
   \begin{itemize}
@@
   careful analysis of data and weighing of benefits) how best to do so.
-  \item The money raised in this way can be used in many ways: buying carbon
-  offsets, supporting ``green computing research'', virtualizing conferences
+\item The money raised in this way can be used in many ways: buying
+  carbon offsets (Mike: could include text from below on this),
+  supporting ``green computing research'', virtualizing conferences
   (livestreaming, etc.), supporting a transition to a low-carbon / few
@@
   \end{itemize}
+
+(Mike: Need to include somewhere above:)
+
+However, going beyond these simple steps is going to be much more
+challenging.  One reason is that it isn't actually very clear what changes
+will help significantly.  (Describe some of our data.  Show the two
+diagrams.  Draw some conclusions / discuss why it's hard to draw
+conclusions.)  More data and more thorough analysis is needed.
 

### 2019-06-18 main.tex
Fleshed out further.

@@
 
-Publicly accounting for CO2e expenditures will incentivize eliminating
-CO2e-consuming activities whose corresponding value to the conference
-is likely to be low. But a public accounting is unlikely to
-incentivize major CO2e reductions, since these will require
-significant changes to conference system itself, which may negatively
-impact the value that conferences provide. 
+Publicly accounting for CO2e expenditures will incentivize some CO2e
+reductions, but probably not major changes to conference system
+itself, since doing so may negatively impact the value that
+conferences provide.
 %
 Science is a fundamentally social process, and the conference system
-accelerates research through high-bandwidth interaction, direct
-dissemination of results, network building, and serendipitous
+accelerates scientific research through high-bandwidth interaction,
+direct dissemination of results, network building, and serendipitous
 cross-fertilization.
 %
-Organizers and attendees will be reticent to harm the scientific
-benefits they have all regularly experienced.
+Organizers and attendees will be reticent to make changes that could
+harm the scientific benefits they have all regularly experienced.
+%
+Nevertheless, it seems clear that the status quo is unsustainable, so
+hard choices will have to be made. How to motivate organizers to make
+them?
 
-(Mike: I think we can significantly trim the text below. Instead, we can
-talk here about a carbon tax as the next natural step. It can be
-related to the sugar tax that some municipalities have imposed; when
-public calorie counts are not enough to discourage use, add a tax
-instead. The arc from the bullet list can come earlier, basically.)
+This dilemma is the one that all of society faces, in a microcosm. To
+address it, many policy experts advocate some form of carbon tax (or
+\emph{price}), which adds a near-term, monetary cost to greenhouse gas
+emissions~\cite{carbonprice}. Doing so makes manifest today the hidden
+environmental cost of CO2e emissions on tomorrow, thus incentivizing
+CO2e-reducing changes without specifying exactly which, thus allowing
+for creative and efficient responses. As an analogy, some
+municipalities such as New York city and Philadelphia have imposed a
+per-calorie tax on soft drinks, with some studies finding that doing
+so indeed reduced consumption~\cite{sodatax}.
 
-Thus we have the question of apples and oranges: How can we weigh the
-environmental harm caused by conference travel against the social
-benefits of the scientific progress that comes from conferences?
+In the best case, governments would impose a CO2e tax uniformly, and
+all CO2e-intensive activities would have to pay it. But ACM could send
+a strong message about the importance of this issue by acting now. In
+particular, as several companies have shown, even an artificial ``internal price'' on
+  carbon creates a significant price signal that shapes behavior.
+%
+  Thus, our second (more ambitious) proposal is: ACM should begin requiring
+  conferences to pay a surcharge based on their carbon footprint.  It should
+  be low at the beginning and increase steadily (and predictably) year
+  on year.
+%
+Conference organizers can choose (on the basis of
+  careful analysis of data and weighing of benefits) how best to do
+  so.
 
-% The question is difficult because the harm and
-% the compensating benefits cannot be easily compared---they are apples and
-% oranges.
-
-Fortunately, we have a well-established mechanism for comparing apples and
-oranges---indeed, for comparing almost anything!  Economists call it ``the
-magic of markets.''  By mapping choices and preferences in disparate domains
-onto a common scale (i.e., money), economic markets function as massively
-distributed computations that optimize costs and benefits across vastly
-different activities.\bcp{Give a simple, concrete example.}  For this
-reason, many policy experts advocate addressing climate change through some
-form of carbon pricing, which places a monetary cost on greenhouse gas
-emissions, sending a clear market signal that changes aggregate
-behavior.\bcp{See
-  https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
-  for lots of good phrasing and citations.}
-
-  \begin{itemize}
-  \item To motivate more painful changes and force people to make difficult
-  choices, we need a price on carbon.
-  \item Ultimately, governments should do this.  But we don't need to wait:
-  as several companies have shown, even an artificial ``internal price'' on
-  carbon creates a significant price signal that shapes behavior.
-  \item Our second (more ambitious) proposal is: ACM should begin requiring
-  conferences to pay a surcharge based on their carbon footprint.  It should
-  be low at the beginning and increase steadily (and predictably) year on year.
-  \item Benefit: This will create concrete downward pressure on conferences
-  to decrease their footprints.  They can then choose (on the basis of
-  careful analysis of data and weighing of benefits) how best to do so.
-\item The money raised in this way can be used in many ways: buying
-  carbon offsets (Mike: could include text from below on this),
-  supporting ``green computing research'', virtualizing conferences
+ What should ACM do with the funds collected? The money raised in
+  this way can be used in many ways: buying carbon offsets, supporting
+  ``green computing research'', virtualizing conferences
   (livestreaming, etc.), supporting a transition to a low-carbon / few
   conferences business model for ACM.
-  \end{itemize}
 
-(Mike: Need to include somewhere above:)
+Once again, ACM should provide support to conference organizers to
+prepare for a future of higher-cost conferences.
 
-However, going beyond these simple steps is going to be much more
-challenging.  One reason is that it isn't actually very clear what changes
-will help significantly.  (Describe some of our data.  Show the two
-diagrams.  Draw some conclusions / discuss why it's hard to draw
-conclusions.)  More data and more thorough analysis is needed.
+(Mike: Here we can talk about data analysis, and how it could provide
+help with this. We've started, and see that the situation is not very
+clear. Describe some of our data.  Show the two diagrams.  Draw some
+conclusions / discuss why it's hard to draw conclusions.
+
+\SECTION{Conclusion}
 

### 2019-06-18 main.tex
Nit

@@
 
-Our first proposal is simple: Every ACM-sponsored conference should
-publicly report its carbon footprint, total and per-participant. A
-conference's footprint will be dominated by participants' air travel,
-but the accounting should not be limited to that: It should also
-include air travel to in-person PC meetings, and estimate hotel and
-food CO2e costs.
+Our first proposal is simple: Every ACM-sponsored conference should publicly
+report its carbon footprint (both aggregate and per-participant). Most
+conferences' footprints will be dominated by their participants' air travel,
+but the accounting should not be limited to that: it should also include air
+travel to in-person PC meetings and estimated hotel and food CO2e footprints.
 
-We hope that a public accounting will nudge conference organizers and
+A public accounting will nudge conference organizers and
 attendees to change their behavior. As an analogy: Chain restaurants

### 2019-06-19 main.tex
Checkpoint

--- 2019-06-18 main.tex
@@
 
-In fact, scientific societies arguably have an especially important role to
+Indeed, scientific societies arguably have an especially important role to
 play, since, for many of their members, travel to conferences may be a
@@
 How should ACM respond to these facts?  We have two concrete
-proposals. First, all ACM conferences should publicly account for CO2e
-spent as a result of hosting them. Second, ACM should embrace putting
-a \emph{price} on carbon in conference budgets, to create pressure on
-organizers to reduce its consumption.
+proposals. First, all ACM conferences should publicly {\em account} for CO2e
+emitted as a result of putting them on. Second, ACM should put a
+\emph{price} on carbon in conference budgets, to create pressure on
+organizers to reduce their footprints.
 
-\SECTION{Mandate public accounting of carbon footprints.}  
-\bcp{The section headers could use some wordsmithing!}
+\SECTION{Mandate public accountability.}  
+% \bcp{The section headers could use some wordsmithing!}
 
 Our first proposal is simple: Every ACM-sponsored conference should
-publicly report its carbon footprint, total and per-participant. A
-conference's footprint will be dominated by participants' air travel,
-but the accounting should not be limited to that: It should also
-include air travel to in-person PC meetings, and estimate hotel and
-food CO2e costs.
+calculate and publicly report its carbon footprint (both aggregate and
+per-participant). Most conferences' footprints will be dominated by
+participants air travel, but the accounting should go beyond this to include
+air travel to in-person PC meetings and estimated hotel and food CO2e
+footprints.
 
-We hope that a public accounting will nudge conference organizers and
-attendees to change their behavior. As an analogy: Chain restaurants
-in the US are required by law to post calorie counts of food items on
-menu boards, and studies show that enlightened customers order, on
-average, up to 50 fewer calories a day, as a
+ACM can provide tools to measure emissions. As a start, SIGPLAN (the special
+interest group on programming languages) recently worked with ACM's SIG
+Governing Board to develop an air-travel-focused CO2e calculator for
+conferences. Users can upload conference registration data, and the
+calculator will estimate the CO2e cost of air travel for those sufficiently
+far away. This calculator could be expanded to account for other conference
+emissions.
+
+Public accounting will nudge conference organizers and attendees to change
+their behavior. By analogy, chain restaurants in the US are required by law
+to post calorie counts of food items on menu boards, and studies show that
+enlightened customers order, on average, up to 50 fewer calories a day as a
 result~\cite{menu}.
 
-ACM can provide tools to help organizers account for this
-footprint. We recently worked with the ACM SGB to develop an air
-travel-focused CO2e calculator. One can input a CSV with registration
-data, and the calculator will compute the estimated CO2e cost of air
-travel for those sufficiently far away. This calculator could be
-expanded to support related costs.
-
-SIGPLAN has been thinking about how to reduce CO2e usage over the past
-couple of years, informed by an accounting of CO2e expenditures. Doing
-so has led several conference organizers \mwh{name which conferences}
-to switch from holding in-person PC meetings to holding the meetings
-on-line instead. \mwh{Mention Jens survey here.} It has also prompted
-adding support for conference livestreaming (to go with talk
-recording) to permit remote participation.
+SIGPLAN has been considering how to reduce CO2e usage over the past couple
+of years, informed by an accounting of its CO2e expenditures. This
+discussion has led the organizers of some of its flagship
+conferences\mwh{name which conferences}\bcp{POPL and ICFP, not PLDI; I don't
+  remember where SPLASH stands} to switch from in-person to on-line program
+committee meetings, joining a trend among other SIGs. \mwh{Mention Jens
+  survey here.} It has also prompted conferences to increase their
+investments in livestreaming and talk recording to permit remote
+participation.
 
@@
 
-This dilemma is the one that all of society faces, in a microcosm. To
-address it, many policy experts advocate some form of carbon tax (or
-\emph{price}), which adds a near-term, monetary cost to greenhouse gas
-emissions~\cite{carbonprice}. Doing so makes manifest today the hidden
-environmental cost of CO2e emissions on tomorrow, thus incentivizing
-CO2e-reducing changes without specifying exactly which, thus allowing
-for creative and efficient responses. As an analogy, some
+This dilemma is a microcosm of the one that all of society faces. To address
+it, many policy experts advocate some form of ``carbon pricing'' (either
+taxes or cap-and-trade schemes), which imposes a concrete, immediate cost on
+greenhouse gas emissions~\cite{carbonprice}. Doing so makes manifest the
+hidden environmental cost of CO2e emissions, thus incentivizing
+CO2e-reducing changes without specifying exactly which, allowing for
+creative and efficient responses. (Continuing the junk-food analogy, some
 municipalities such as New York city and Philadelphia have imposed a
-per-calorie tax on soft drinks, with some studies finding that doing
-so indeed reduced consumption~\cite{sodatax}.
+per-calorie tax on soft drinks, with some studies finding that doing so
+indeed reduced consumption~\cite{sodatax}.)
 
-In the best case, governments would impose a CO2e tax uniformly, and
-all CO2e-intensive activities would have to pay it. But ACM could send
-a strong message about the importance of this issue by acting now. In
-particular, as several companies have shown, even an artificial ``internal price'' on
-  carbon creates a significant price signal that shapes behavior.
+Ideally, at some point, governments will impose carbon pricing uniformly,
+and all CO2e-intensive activities will have to pay it. But ACM can send a
+strong message about the importance of this issue by acting now. In
+particular, as several companies have shown, even an artificial ``internal
+price'' on carbon creates a significant price signal that shapes behavior.
 %
-  Thus, our second (more ambitious) proposal is: ACM should begin requiring
-  conferences to pay a surcharge based on their carbon footprint.  It should
-  be low at the beginning and increase steadily (and predictably) year
-  on year.
+Thus, our second (more ambitious) proposal is: ACM should impose a a
+surcharge on conferences based on their carbon footprint.  The charge should
+be low at the beginning and increase steadily (and predictably) year on
+year.
 %
-Conference organizers can choose (on the basis of
-  careful analysis of data and weighing of benefits) how best to do
-  so.
+Conference organizers can then choose (on the basis of careful analysis of
+data and weighing of benefits) how best to balance their budgets---whether
+by decreasing emissions, increasing registration fees, soliciting additional
+corporate sponsorship, or some combination.
 
- What should ACM do with the funds collected? The money raised in
-  this way can be used in many ways: buying carbon offsets, supporting
-  ``green computing research'', virtualizing conferences
-  (livestreaming, etc.), supporting a transition to a low-carbon / few
-  conferences business model for ACM.
-
-Once again, ACM should provide support to conference organizers to
-prepare for a future of higher-cost conferences.
+What should ACM do with the funds collected from this surcharge?  One can
+imagine many good uses, including buying carbon
+offsets~\cite{CarbonOFfsetReport}, supporting ``green'' computing
+research\bcp{can we mention some specifics?}, and defraying the costs of
+virtualizing conferences (livestreaming, etc.).
 
@@
 
-(Some kind of closing.)
+The climate crisis is too important to ignore and too urgent to leave to
+world leaders, to address at their own pace: rather, organizations at every
+scale must confront their own contributions, raise awareness among their
+members, and begin establishing new ways of doing business in the
+lower-carbon future that is coming.  ACM should lead the way!
 

### 2019-06-19 main.tex
Done with my pass

@@
 
-Most recently, we have been investigating how to purchase \emph{carbon
-  offsets} to defray CO2e costs.
+Another simple, short-term expedient for lowering net the emissions
+footprint of conferences is purchasing \emph{carbon
+  offsets}~\cite{CarbonOFfsetReport}.
 %
-A carbon offset is sold by a vendor who uses the raised funds to
-finance an activity that permanently avoids emitting an amount of
-greenhouse gases. That the CO2e emission avoidance is truth and not
-appearance can be certified by a watchdog organization. Typically,
-planting trees is not a certifiable activity (since it's hard to
-guarantee the trees won't be cut down), but installing methane capture
-devices on landfill sites or buying fuel-efficient stoves to replace
-open-fire cooking in poor communities can be.
+A carbon offset is sold by a vendor who uses the raised funds to finance an
+activity that permanently avoids emitting some amount of greenhouse gases.
+The veracity and permanence of this CO2e emission avoidance is certified by
+a watchdog organization. (Typically, planting trees is not a certifiable
+activity, since it's hard to guarantee the trees won't be cut down, but
+installing methane capture devices on landfill sites or buying
+fuel-efficient stoves to replace open-fire cooking in poor communities can
+be.)
+%
+Many organizations, including companies such as Google, Dell, Microsoft,
+General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
+and some academic societies, now buy carbon offsets in order to reduce their
+net footprint. ACM conferences can do the same, and the purchases can be
+included in the public accounting we are proposing.  In fact, ACM recently
+decided to allow conferences to allow attendees to pay for an offset during
+registration; the conference then uses the designated funds to purchase the
+offset from a qualified vendor.  Some conferences are also experimenting
+with corporate sponsorship of bulk offsets for the whole conference.  (It
+should be noted that carbon offsets are a controversial tool.  Besides
+issues with verifiability and permanence, some observers worry
+that---especially given their currently quite low cost---they will encourage
+complacency, leading organizations to postpone making the difficult changes
+that will certainly be required to fully decarbonize society.  We agree that
+offsets are at best a short-term measure, to be considered in parallel with,
+not instead of, more difficult cutting.)
 
-Many organizations (including companies such as Google, Dell,
-Microsoft, General Motors, Delta Airlines, Lyft, and Expedia, numerous
-universities, and a number of academic societies) now buy carbon
-offsets in order to reduce their overall footprint. ACM conferences
-could do the same, and the purchases would be included in their public
-accounting. We have been investigating this possibility for the past
-year. ACM has now agreed that conference attendees can opt to purchase
-an offset during registration, and ACM will use the designated funds
-to purchase an offset. Even getting to this stage required some
-investigation of the legality of ACM making the purchase on behalf of
-the registrant~\cite{CarbonOFfsetReport}. Ideally, the purchase of
-offsets would be made even simpler, but doing so requires further
-thought.
+Beyond encouraging conferences to take these sorts of simple steps, public
+accounting of conference emissions footprints will generate valuable data
+that future organizers can use to optimize their own footprints.
+%
+\bcp{Talk about the little bit of data gathering that we've done.  Make the
+  point that, if the public accounting of conference carbon footprints is
+  detailed enough (e.g., including an origin city for each participant,
+  rather than just aggregate statistics on average per-participant
+  footprint), and if it's publicized in a centralized place and in a uniform
+  format, then it becomes a resource for all sorts of useful analysis.}
 
@@
 
-Publicly accounting for CO2e expenditures will incentivize some CO2e
-reductions, but probably not major changes to conference system
-itself, since doing so may negatively impact the value that
-conferences provide.
+However, while publicly accounting for CO2e expenditures will incentivize
+some CO2e reductions, it will probably not, by itself, induce major changes
+to the conference system.  There are good reasons for this, since such
+changes will undoubtedly impact the value that conferences provide.
 %
@@
 harm the scientific benefits they have all regularly experienced.
-%
-Nevertheless, it seems clear that the status quo is unsustainable, so
-hard choices will have to be made. How to motivate organizers to make
-them?
+
+Indeed, a preliminary analysis that we have done with registration data for
+recent SIGPLAN conferences raises as many questions as it answers.
+\bcp{Talk about our data analysis, and how it could provide help with
+  this. We've started, and see that the situation is not very
+  clear. Describe some of our data.  Show the two diagrams.  Draw some
+  conclusions / discuss why it's hard to draw conclusions.  (This would be a
+  good place to refer to~\cite{ClimateCommitteReport}---we can enumerate
+  some of the alternatives that one might have hoped the data would help us
+  choose between.)}
+
+\bcp{Tricky transition here.  I think the point we want to make is that
+  these decisions are genuinely hard and painful, which is exactly what our
+  data shows.}
+
+Nevertheless, it seems clear that the status quo is unsustainable, so hard
+choices will have to be made. How do we motivate organizers to make them?
 
@@
 research\bcp{can we mention some specifics?}, and defraying the costs of
-virtualizing conferences (livestreaming, etc.).
-
-(Mike: Here we can talk about data analysis, and how it could provide
-help with this. We've started, and see that the situation is not very
-clear. Describe some of our data.  Show the two diagrams.  Draw some
-conclusions / discuss why it's hard to draw conclusions.
+virtualizing conferences (livestreaming, etc.).  \bcp{Expand this paragraph
+  a bit.}
 
@@
 
-The climate crisis is too important to ignore and too urgent to leave to
-world leaders, to address at their own pace: rather, organizations at every
-scale must confront their own contributions, raise awareness among their
-members, and begin establishing new ways of doing business in the
-lower-carbon future that is coming.  ACM should lead the way!
+\bcp{I wish we could put a paragraph here about the long-term picture and
+  the need to rethink scientific exchange and ACM's business model.  That
+  will please our grumpy reviewer, and I think it's actually a point worth
+  making.}
+
+The climate crisis is too important to ignore and too urgent to leave for
+world leaders to address at their own pace: Organizations at every scale
+must confront their own contributions, raise awareness among their members,
+and begin establishing new ways of doing business in the lower-carbon future
+that is coming.  ACM should lead the way!
 

### 2019-06-19 main.tex
Some more work on transitions

@@
 
-ACM can provide tools to measure emissions. As a start, SIGPLAN (the special
-interest group on programming languages) recently worked with ACM's SIG
-Governing Board to develop an air-travel-focused CO2e calculator for
-conferences. Users can upload conference registration data, and the
-calculator will estimate the CO2e cost of air travel for those sufficiently
-far away. This calculator could be expanded to account for other conference
-emissions.
-
 Public accounting will nudge conference organizers and attendees to change
@@
 
+ACM can provide tools for gathering and publicizing this data. As a start,
+SIGPLAN (the special interest group on programming languages) recently
+worked with ACM's SIG Governing Board to develop an air-travel-focused CO2e
+calculator for conferences. Users can upload conference registration data,
+and the calculator will estimate the CO2e cost of air travel for those
+sufficiently far away. This calculator could be expanded to account for
+other conference emissions.
+
 \SECTION{Put a Price on Carbon.}
@@
 
-Indeed, a preliminary analysis that we have done with registration data for
-recent SIGPLAN conferences raises as many questions as it answers.
-\bcp{Talk about our data analysis, and how it could provide help with
-  this. We've started, and see that the situation is not very
-  clear. Describe some of our data.  Show the two diagrams.  Draw some
-  conclusions / discuss why it's hard to draw conclusions.  (This would be a
-  good place to refer to~\cite{ClimateCommitteReport}---we can enumerate
-  some of the alternatives that one might have hoped the data would help us
-  choose between.)}
+\begin{figure}[t]
+\centering
+\includegraphics[width=6in]{SIGPLAN-confs.png}
+\includegraphics[width=6in]{ParticipantsOriginAll.png}
+\caption{Top: carbon footprint per participant for travel to recent SIGPLAN
+  conferences\iflater\bcp{Get better file from the other paper, or get
+    better version of this one from Crista.}\fi.  Bottom: Origin of
+  participants for each conference.  \bcp{Explain in more detail, once we
+    have Crista's new figures.}} 
+\label{fig:confs}
+\end{figure}
 
-\bcp{Tricky transition here.  I think the point we want to make is that
-  these decisions are genuinely hard and painful, which is exactly what our
-  data shows.}
+The registration data we've gathered from SIGPLAN conferences illustrates
+some of the difficulties.  
+%
+\bcp{Explain the two diagrams.  Discuss how they can be used to draw
+  contradictory conclusions.  E.g., one of the many footprint-reduction
+  measures that have been discussed in SIGPLAN~\cite{ClimateCommitteReport}
+  is locating conferences near their historical ``center of mass.''  Looking
+  at the top diagram, one might thus be tempted to locate all conferences in
+  the northeast US or western Europe.  But the bottom figure shows that much
+  of the participation in each instance of a conference is local; always
+  holding a conference at its ``center of mass'' will damage local
+  communities everywhere else.  (Maybe this would be a good place to refer
+  to~\cite{ClimateCommitteReport}---we can enumerate some of the
+  alternatives that one might have hoped the data would help us choose
+  between.)}  
+%
+\bcp{Maybe we can replace the bottom part of the figure by some clearer way
+  of visualizing the observation that much of the attendance at each
+  conference is local?  E.g., what if, instead of color-coding the different
+  sections of the bars by continent, we instead color-coded them by distance
+  traveled (in buckets of 500 miles, say)?}
+
+\bcp{Summary: The data shows why these decisions are genuinely hard and
+  painful.}
 

### 2019-06-20 main.tex
Small changes.

@@
 
-How should ACM respond to these facts?  We have two concrete
+How should ACM respond to these facts? We are the members of SIGPLAN's
+(the special interest group on programming languages) Climate
+Committee,\footnote{https://www.sigplan.org/Resources/Climate/} and we
+have been researching this 
+question for the past couple of years. Our work has led to two concrete
 proposals. First, all ACM conferences should publicly {\em account} for CO2e
@@
 
-Another simple, short-term expedient for lowering net the emissions
-footprint of conferences is purchasing \emph{carbon
+SIGPLAN has also been exploring how to mitigate a conference's CO2e emissions
+by purchasing \emph{carbon
   offsets}~\cite{CarbonOFfsetReport}.
@@
 included in the public accounting we are proposing.  In fact, ACM recently
-decided to allow conferences to allow attendees to pay for an offset during
+decided that conferences could offer attendees the option to pay for an offset during
 registration; the conference then uses the designated funds to purchase the
 offset from a qualified vendor.  Some conferences are also experimenting
-with corporate sponsorship of bulk offsets for the whole conference.  (It
+with corporate sponsorship of bulk offsets for the whole conference.\footnote{It
 should be noted that carbon offsets are a controversial tool.  Besides
 issues with verifiability and permanence, some observers worry
-that---especially given their currently quite low cost---they will encourage
+that---especially given their currently artificially-low cost---they will encourage
 complacency, leading organizations to postpone making the difficult changes
-that will certainly be required to fully decarbonize society.  We agree that
-offsets are at best a short-term measure, to be considered in parallel with,
-not instead of, more difficult cutting.)
+that will be required to sustainably decarbonize society.  We believe that
+offsets are a short-term measure, to be considered in parallel with,
+not instead of, more difficult cutting.}
 
@@
 ACM can provide tools for gathering and publicizing this data. As a start,
-SIGPLAN (the special interest group on programming languages) recently
+SIGPLAN recently
 worked with ACM's SIG Governing Board to develop an air-travel-focused CO2e
@@
 
-However, while publicly accounting for CO2e expenditures will incentivize
+While publicly accounting for CO2e expenditures will incentivize
 some CO2e reductions, it will probably not, by itself, induce major changes
-to the conference system.  There are good reasons for this, since such
-changes will undoubtedly impact the value that conferences provide.
+to the conference system. 
 %
@@
 by decreasing emissions, increasing registration fees, soliciting additional
-corporate sponsorship, or some combination.
+corporate sponsorship, or some combination. \mwh{Problem: We showed
+  just above that it's hard to figure out what to do. Come back to
+  this point here and say that the surcharge will spur bold choices?
+  Maybe say that these choices won't necessarily be bad, but could be
+  really great?}
 
@@
 imagine many good uses, including buying carbon
-offsets~\cite{CarbonOFfsetReport}, supporting ``green'' computing
+offsets, supporting ``green'' computing
 research\bcp{can we mention some specifics?}, and defraying the costs of

### 2019-06-20 main.tex
Little things

@@
 included in the public accounting we are proposing.  In fact, ACM recently
-decided that conferences could offer attendees the option to pay for an offset during
-registration; the conference then uses the designated funds to purchase the
-offset from a qualified vendor.  Some conferences are also experimenting
-with corporate sponsorship of bulk offsets for the whole conference.\footnote{It
-should be noted that carbon offsets are a controversial tool.  Besides
-issues with verifiability and permanence, some observers worry
-that---especially given their currently artificially-low cost---they will encourage
-complacency, leading organizations to postpone making the difficult changes
-that will be required to sustainably decarbonize society.  We believe that
-offsets are a short-term measure, to be considered in parallel with,
-not instead of, more difficult cutting.}
+decided that conferences could offer attendees the option to pay for an
+offset during registration; the conference then uses the designated funds to
+purchase the offset from a qualified vendor.  Some conferences are also
+experimenting with corporate sponsorship of bulk offsets for the whole
+conference.\footnote{We acknowledge that carbon offsets are a
+  controversial tool\bcp{Cite the thing the reviewer pointed us to, as well
+    as the recent controversial UN report.}.  Besides issues with
+  verifiability and permanence, some observers worry that---especially given
+  their currently artificially-low cost---they will encourage complacency,
+  leading organizations to postpone making the difficult changes that will
+  be required to sustainably decarbonize society.  We believe that offsets
+  are a short-term measure, to be considered in parallel with, not instead
+  of, more difficult cutting.}
 
@@
 by decreasing emissions, increasing registration fees, soliciting additional
-corporate sponsorship, or some combination. \mwh{Problem: We showed
+corporate sponsorship, or by other, more creative, means. \mwh{Problem: We showed
   just above that it's hard to figure out what to do. Come back to
@@
   will please our grumpy reviewer, and I think it's actually a point worth
-  making.}
+  making.\cite{NCN}}
 

### 2019-06-20 main.tex
Better version of the alternative origin chart
Switched to using sigplan-confs2
Added distances chart

@@
 \centering
-\includegraphics[width=6in]{SIGPLAN-confs.png}
-\includegraphics[width=6in]{ParticipantsOriginAll.png}
+\includegraphics[width=6in]{SIGPLAN-confs2.png}
+\includegraphics[width=6in]{ParticipantsOriginAll2.png}
+\includegraphics[width=6in]{Distances.png}
 \caption{Top: carbon footprint per participant for travel to recent SIGPLAN
   conferences\iflater\bcp{Get better file from the other paper, or get
-    better version of this one from Crista.}\fi.  Bottom: Origin of
+    better version of this one from Crista.}\fi.  Middle: Origin of
   participants for each conference. \bcp{Explain in more detail, once we
-    have Crista's new figures.}} 
+    have Crista's new figures.} Bottom: Distances traveled by participants, one way, in Km; middle dashed line is the median travel distance.}
 \label{fig:confs}

### 2019-06-21 main.tex
Checkpoint

--- 2019-06-20 main.tex
@@
 
-\title{\Huge Conferences in an Era of Expensive Carbon}
+\title{\Huge Conferences in the Era of Expensive Carbon}
 % \title{\Huge ACM Conferences and Cost of Carbon}
@@
 
-How should ACM respond to these facts? We are the members of SIGPLAN's
-(the special interest group on programming languages) Climate
-Committee,\footnote{https://www.sigplan.org/Resources/Climate/} and we
-have been researching this 
-question for the past couple of years. Our work has led to two concrete
-proposals. First, all ACM conferences should publicly {\em account} for CO2e
-emitted as a result of putting them on. Second, ACM should put a
-\emph{price} on carbon in conference budgets, to create pressure on
-organizers to reduce their footprints.
+How should ACM respond to these facts?  Two years ago, SIGPLAN (ACM's
+Special Interest Group on Programming Languages) convened an ad hoc Climate
+Committee\footnote{https://www.sigplan.org/Resources/Climate/} to research
+this question.  After investigating many
+options~\cite{ClimateCommitteReport} and discussing intensively with SIGPLAN
+members and conference organizers, ACM staff, and other SIG leaders, we have
+two concrete proposals. First, all ACM conferences should publicly {\em
+  account} for CO2e emitted as a result of putting them on. Second, ACM
+should put a \emph{price} on carbon in conference budgets, to create
+pressure on organizers to reduce their footprints.
 
 \SECTION{Mandate public accountability.}
-% \bcp{The section headers could use some wordsmithing!}
 
 Our first proposal is simple: Every ACM-sponsored conference should
-calculate and publicly report its carbon footprint (both aggregate and
-per-participant). Most conferences' footprints will be dominated by
-participants air travel, but the accounting should go beyond this to include
-air travel to in-person PC meetings and estimated hotel and food CO2e
-footprints.
+calculate and publicly report its carbon footprint. Most conferences'
+footprints will be dominated by participants air travel, but the accounting
+should go beyond this to include air travel to in-person PC meetings and
+estimated hotel and food CO2e footprints.
 
@@
 of years, informed by an accounting of its CO2e expenditures. This
-discussion has led the organizers of some of its flagship
-conferences\mwh{name which conferences}\bcp{POPL and ICFP, not PLDI; I don't
-  remember where SPLASH stands} to switch from in-person to on-line program
-committee meetings, joining a trend among other SIGs. \mwh{Mention Jens
-  survey here.} It has also prompted conferences to increase their
-investments in livestreaming and talk recording to permit remote
-participation.
+discussion has led the organizers of two of its flagship conferences (POPL
+and ICFP) to switch from in-person to on-line program committee meetings,
+joining a trend among other SIGs (\bcp{ZZZ} out of \bcp{WWW} report that
+their flagship meetings now have virtual PC meetings). It has also prompted
+conferences to increase their investments in livestreaming and talk
+recording to permit remote participation.
 
-SIGPLAN has also been exploring how to mitigate a conference's CO2e emissions
-by purchasing \emph{carbon
-  offsets}~\cite{CarbonOFfsetReport}.
+SIGPLAN has also explored how to mitigate a conference's CO2e emissions by
+purchasing \emph{carbon offsets}~\cite{CarbonOFfsetReport}.
 %
@@
 experimenting with corporate sponsorship of bulk offsets for the whole
-conference.\footnote{We acknowledge that carbon offsets are a
-  controversial tool\bcp{Cite the thing the reviewer pointed us to, as well
-    as the recent controversial UN report.}.  Besides issues with
+conference.\footnote{We acknowledge that carbon offsets are a controversial
+  tool~\cite{anderson2012inconvenient,
+  carbon-offsets-are-not-our-get-out-jail-free-card}.  Besides issues with
   verifiability and permanence, some observers worry that---especially given
@@
 accounting of conference emissions footprints will generate valuable data
-that future organizers can use to optimize their own footprints.
-%
-\bcp{Talk about the little bit of data gathering that we've done.  Make the
-  point that, if the public accounting of conference carbon footprints is
-  detailed enough (e.g., including an origin city for each participant,
-  rather than just aggregate statistics on average per-participant
-  footprint), and if it's publicized in a centralized place and in a uniform
-  format, then it becomes a resource for all sorts of useful analysis.}
+that future organizers can use to optimize their own footprints.  Emissions
+data should be published both in aggregate form (total footprint for the
+conference and per-participant footprint), for quick reference, and in a raw
+form (a suitably anonymized record for each individual attending each
+conference, including originating city), and it should be made available in
+a central place and in a uniform format, to facilitate analysis.
 
 ACM can provide tools for gathering and publicizing this data. As a start,
-SIGPLAN recently
-worked with ACM's SIG Governing Board to develop an air-travel-focused CO2e
-calculator for conferences. Users can upload conference registration data,
-and the calculator will estimate the CO2e cost of air travel for those
-sufficiently far away. This calculator could be expanded to account for
-other conference emissions.
+SIGPLAN recently worked with ACM's SIG Governing Board to develop an
+air-travel-focused CO2e calculator for conferences. Users can upload
+conference registration data, and the calculator will estimate the CO2e cost
+of air travel for those sufficiently far away. This calculator could be
+expanded to account for other forms of conference emissions.
 
@@
 \centering
-\includegraphics[width=6in]{SIGPLAN-confs.png}
-\includegraphics[width=6in]{ParticipantsOriginAll.png}
-\caption{Top: carbon footprint per participant for travel to recent SIGPLAN
+\includegraphics[width=6in]{SIGPLAN-confs2.png}
+\includegraphics[width=6in]{ParticipantsOriginAll2.png}
+% \includegraphics[width=6in]{Distances.png}
+\caption{Top: Carbon footprint per participant for travel to recent SIGPLAN
   conferences\iflater\bcp{Get better file from the other paper, or get
-    better version of this one from Crista.}\fi.  Bottom: Origin of
-  participants for each conference.  \bcp{Explain in more detail, once we
-    have Crista's new figures.}} 
+    better version of this one from Crista.}\fi.  (Small dots represent
+  \bcp{XXX} tons of CO2e per participant; the largest dot represents
+  \bcp{YYY} tons per participant.) Bottom: Breakdown of 
+  continent-of-origin for participants in each conference. 
+% Bottom: Distances traveled by participants, one way, in Km; middle dashed
+% line is the median travel distance. 
+}
 \label{fig:confs}
@@
 
-The registration data we've gathered from SIGPLAN conferences illustrates
+Registration data from recent SIGPLAN conferences illustrates
 some of the difficulties.  

### 2019-06-21 main.tex
A mostly finished first draft

@@
 
-Our first proposal is simple: Every ACM-sponsored conference should
-calculate and publicly report its carbon footprint. Most conferences'
+Our first proposal is simple: {\bf Every ACM-sponsored conference should
+calculate and publicly report its carbon footprint.} Most conferences'
 footprints will be dominated by participants air travel, but the accounting
@@
 and ICFP) to switch from in-person to on-line program committee meetings,
-joining a trend among other SIGs (\bcp{ZZZ} out of \bcp{WWW} report that
-their flagship meetings now have virtual PC meetings). It has also prompted
-conferences to increase their investments in livestreaming and talk
-recording to permit remote participation.
+joining a trend among other SIGs, most of which now use virtual PC meetings
+for some or all of their flagship conferences. It has also prompted
+conferences to increase their investments in livestreaming and video
+recording to support remote participation.
 
@@
 experimenting with corporate sponsorship of bulk offsets for the whole
-conference.\footnote{We acknowledge that carbon offsets are a controversial
+conference.\footnote{We acknowledge that carbon offsets are a somewhat
+  controversial 
   tool~\cite{anderson2012inconvenient,
@@
   verifiability and permanence, some observers worry that---especially given
-  their currently artificially-low cost---they will encourage complacency,
+  their currently artificially low cost---they will encourage complacency,
   leading organizations to postpone making the difficult changes that will
   be required to sustainably decarbonize society.  We believe that offsets
-  are a short-term measure, to be considered in parallel with, not instead
-  of, more difficult cutting.}
+  should be considered as a useful short-term expedient, to be considered in
+  parallel with, not instead of, more difficult cutting.}
 
@@
 
-Registration data from recent SIGPLAN conferences illustrates
-some of the difficulties.  
+The two visualizations of registration data from some recent conferences in
+Figure~\ref{fig:confs} illustrate some of the difficulties.  The top diagram
+shows an estimated per-participant CO2e footprint for each instance of the
+four major SIGPLAN conferences over the past ten years (excluding a few for
+which we had trouble getting data), with larger dots representing more
+emissions.  Eyeballing this diagram, it appears that carbon-conscious
+conference planners should hold all of these conferences either in the
+northeast US or in western Europe every year.  But the bottom visualization
+tells a different story.  Here, the horizontal colored bar at the bottom
+represents the continent on which each conference was held, and each
+vertical bar gives a breakdown of the participants in that conference,
+colored according to the continent they gave as their home address.  Again,
+from a superficial glance at the colors, it is clear that, though a minority
+come from far away, the majority of participants in each case are from the
+region where the conference is being held---that is, there is a large amount
+of local participation.\iflater\bcp{That point could be strengthened by
+  adding some numbers (just in the text, not a figure) about what percentage
+  of conference attendees attend even one additional instance of the same
+  conference...}\fi{} From this, one might conclude that always locating
+conferences in the same one or two places would significantly damage the
+research community by discouraging participation from other parts of the
+world. 
 %
-\bcp{Explain the two diagrams.  Discuss how they can be used to draw
-  contradictory conclusions.  E.g., one of the many footprint-reduction
-  measures that have been discussed in SIGPLAN~\cite{ClimateCommitteReport}
-  is locating conferences near their historical ``center of mass.''  Looking
-  at the top diagram, one might thus be tempted to locate all conferences in
-  the northeast US or western Europe.  But the bottom figure shows that much
-  of the participation in each instance of a conference is local; always
-  holding a conference at its ``center of mass'' will damage local
-  communities everywhere else.  (Maybe this would be a good place to refer
-  to~\cite{ClimateCommitteReport}---we can enumerate some of the
-  alternatives that one might have hoped the data would help us choose
-  between.)}  
 %
-\bcp{Maybe we can replace the bottom part of the figure by some clearer way
-  of visualizing the observation that much of the attendance at each
-  conference is local?  E.g., what if, instead of color-coding the different
-  sections of the bars by continent, we instead color-coded them by distance
-  traveled (in buckets of 500 miles, say)?}
+% \bcp{Explain the two diagrams.  Discuss how they can be used to draw
+%   contradictory conclusions.  E.g., one of the many footprint-reduction
+%   measures that have been discussed in SIGPLAN~\cite{ClimateCommitteReport}
+%   is locating conferences near their historical ``center of mass.''  Looking
+%   at the top diagram, one might thus be tempted to locate all conferences in
+%   the northeast US or western Europe.  But the bottom figure shows that much
+%   of the participation in each instance of a conference is local; always
+%   holding a conference at its ``center of mass'' will damage local
+%   communities everywhere else.  (Maybe this would be a good place to refer
+%   to~\cite{ClimateCommitteReport}---we can enumerate some of the
+%   alternatives that one might have hoped the data would help us choose
+%   between.)}  
+% %
+% \bcp{Maybe we can replace the bottom part of the figure by some clearer way
+%   of visualizing the observation that much of the attendance at each
+%   conference is local?  E.g., what if, instead of color-coding the different
+%   sections of the bars by continent, we instead color-coded them by distance
+%   traveled (in buckets of 500 miles, say)?}
+%
+These disparate conclusions suggest that reducing conference emissions will
+require genuinely hard painful compromises.  The impulse to avoid or
+postpone thinking about it is entirely understandable!
 
-\bcp{Summary: The data shows why these decisions are genuinely hard and
-  painful.}
-
-Nevertheless, it seems clear that the status quo is unsustainable, so hard
-choices will have to be made. How do we motivate organizers to make them?
+However, the present trajectory of world emissions is unsustainable, and
+hard choices will have to be made if ACM is to play its part by reducing its
+own emissions. How do we motivate organizers to make them?
 
@@
 taxes or cap-and-trade schemes), which imposes a concrete, immediate cost on
-greenhouse gas emissions~\cite{carbonprice}. Doing so makes manifest the
-hidden environmental cost of CO2e emissions, thus incentivizing
-CO2e-reducing changes without specifying exactly which, allowing for
-creative and efficient responses. (Continuing the junk-food analogy, some
-municipalities such as New York city and Philadelphia have imposed a
-per-calorie tax on soft drinks, with some studies finding that doing so
-indeed reduced consumption~\cite{sodatax}.)
+emissions~\cite{carbonprice}. Doing so makes manifest the hidden
+environmental cost of CO2e emissions, thus incentivizing CO2e-reducing
+changes without specifying exactly which, allowing for creative and
+efficient responses. (Continuing the junk-food analogy, some municipalities
+such as New York city and Philadelphia have imposed a per-calorie tax on
+soft drinks, with some studies finding that doing so indeed reduced
+consumption~\cite{sodatax}.)
 
@@
 %
-Thus, our second (more ambitious) proposal is: ACM should impose a a
-surcharge on conferences based on their carbon footprint.  The charge should
-be low at the beginning and increase steadily (and predictably) year on
-year.
+Thus, our second proposal is that {\bf ACM should impose a a surcharge on
+  conferences based on their carbon footprint.}  The charge should be low at
+the beginning and increase steadily (and predictably) year on year.
 %
@@
 data and weighing of benefits) how best to balance their budgets---whether
-by decreasing emissions, increasing registration fees, soliciting additional
-corporate sponsorship, or by other, more creative, means. \mwh{Problem: We showed
-  just above that it's hard to figure out what to do. Come back to
-  this point here and say that the surcharge will spur bold choices?
-  Maybe say that these choices won't necessarily be bad, but could be
-  really great?}
+by decreasing per-participant emissions, decreasing (physical)
+participation, increasing registration fees, soliciting additional corporate
+sponsorship, or other, more creative, means.  \iflater\mwh{Problem: We
+  showed just above that it's hard to figure out what to do. Come back to
+  this point here and say that the surcharge will spur bold choices?  Maybe
+  say that these choices won't necessarily be bad, but could be really
+  great?}\fi
 
 What should ACM do with the funds collected from this surcharge?  One can
-imagine many good uses, including buying carbon
-offsets, supporting ``green'' computing
-research\bcp{can we mention some specifics?}, and defraying the costs of
-virtualizing conferences (livestreaming, etc.).  \bcp{Expand this paragraph
-  a bit.}
+imagine many good uses, including carbon offsetting, supporting ``green''
+computing research\iflater\bcp{can we mention some specifics?}\fi, and
+defraying the costs of virtualizing conferences (livestreaming, etc.).
+\iflater\bcp{Expand this paragraph a bit.}\fi
 
 \SECTION{Conclusion}
+
+% Ultimately, the carbon cost of air travel may get so high, and individual
+% carbon budgets so low, that people simply stop flying except on rare
+% occasions.
 

### 2019-06-21 main.tex
Add crista numbers

@@
   conferences\iflater\bcp{Get better file from the other paper, or get
-    better version of this one from Crista.}\fi.  (Small dots represent
-  \bcp{XXX} tons of CO2e per participant; the largest dot represents
-  \bcp{YYY} tons per participant.) Bottom: Breakdown of 
-  continent-of-origin for participants in each conference. 
+    better version of this one from Crista.}\fi.  (The smallest dot (ICFP
+  14, in Gothenburg, Sweden) represents .9 tons of CO2e per participant; the
+  largest dot (ICFP 16, in Nara Japan) represents \bcp{1.94} tons per
+  participant.) Bottom: Breakdown of continent-of-origin for participants in
+  each conference.
 % Bottom: Distances traveled by participants, one way, in Km; middle dashed

### 2019-06-23 main.tex
Complete draft

@@
 Special Interest Group on Programming Languages) convened an ad hoc Climate
-Committee\footnote{https://www.sigplan.org/Resources/Climate/} to research
-this question.  After investigating many
-options~\cite{ClimateCommitteReport} and discussing intensively with SIGPLAN
-members and conference organizers, ACM staff, and other SIG leaders, we have
-two concrete proposals. First, all ACM conferences should publicly {\em
-  account} for CO2e emitted as a result of putting them on. Second, ACM
-should put a \emph{price} on carbon in conference budgets, to create
-pressure on organizers to reduce their footprints.
+Committee\cite{ClimateCommittee} to research this question.  After
+investigating many options~\cite{ClimateCommitteReport} and discussing
+intensively with SIGPLAN members and conference organizers, ACM staff, and
+other SIG leaders, we have two concrete proposals. First, all ACM
+conferences should publicly {\em account} for CO2e emitted as a result of
+putting them on. Second, ACM should put a \emph{price} on carbon in
+conference budgets, to create pressure on organizers to reduce their
+footprints.
 
@@
 Our first proposal is simple: {\bf Every ACM-sponsored conference should
-calculate and publicly report its carbon footprint.} Most conferences'
+publicly report its carbon footprint.} Most conferences'
 footprints will be dominated by participants air travel, but the accounting
@@
 
-SIGPLAN has been considering how to reduce CO2e usage over the past couple
-of years, informed by an accounting of its CO2e expenditures. This
-discussion has led the organizers of two of its flagship conferences (POPL
-and ICFP) to switch from in-person to on-line program committee meetings,
-joining a trend among other SIGs, most of which now use virtual PC meetings
-for some or all of their flagship conferences. It has also prompted
-conferences to increase their investments in livestreaming and video
-recording to support remote participation.
+As a case in point, SIGPLAN has been considering how to reduce CO2e usage
+over the past couple of years, informed by an accounting of its own CO2e
+expenditures. This discussion has led the organizers of two of its flagship
+conferences (POPL and ICFP) to switch from in-person to on-line program
+committee meetings, joining a trend among other SIGs, most of which now use
+virtual PC meetings for some or all of their flagship conferences. It has
+also prompted several conferences to increase their investments in
+livestreaming and video recording to support remote participation.
 
@@
 Beyond encouraging conferences to take these sorts of simple steps, public
-accounting of conference emissions footprints will generate valuable data
-that future organizers can use to optimize their own footprints.  Emissions
-data should be published both in aggregate form (total footprint for the
-conference and per-participant footprint), for quick reference, and in a raw
+accounting of conference emissions will generate data that future organizers
+can use.  To this end, emissions data should be published both in aggregate
+form (total and per-participant footprint), for quick reference, and in raw
 form (a suitably anonymized record for each individual attending each
@@
 
-ACM can provide tools for gathering and publicizing this data. As a start,
-SIGPLAN recently worked with ACM's SIG Governing Board to develop an
+ACM should provide tools for gathering and publicizing this data. As a
+start, SIGPLAN recently worked with ACM's SIG Governing Board to develop an
 air-travel-focused CO2e calculator for conferences. Users can upload
@@
 
-While publicly accounting for CO2e expenditures will incentivize
-some CO2e reductions, it will probably not, by itself, induce major changes
-to the conference system. 
+While publicly accounting for CO2e expenditures will incentivize some CO2e
+reductions, it is not likely, by itself, to induce major changes.
 %
@@
 %
-Organizers and attendees will be reticent to make changes that could
-harm the scientific benefits they have all regularly experienced.
+Organizers and attendees will be reticent to make changes that might impact
+these benefits.
 
@@
 The two visualizations of registration data from some recent conferences in
-Figure~\ref{fig:confs} illustrate some of the difficulties.  The top diagram
-shows an estimated per-participant CO2e footprint for each instance of the
-four major SIGPLAN conferences over the past ten years (excluding a few for
-which we had trouble getting data), with larger dots representing more
-emissions.  Eyeballing this diagram, it appears that carbon-conscious
-conference planners should hold all of these conferences either in the
-northeast US or in western Europe every year.  But the bottom visualization
-tells a different story.  Here, the horizontal colored bar at the bottom
-represents the continent on which each conference was held, and each
-vertical bar gives a breakdown of the participants in that conference,
-colored according to the continent they gave as their home address.  Again,
-from a superficial glance at the colors, it is clear that, though a minority
-come from far away, the majority of participants in each case are from the
-region where the conference is being held---that is, there is a large amount
-of local participation.\iflater\bcp{That point could be strengthened by
-  adding some numbers (just in the text, not a figure) about what percentage
-  of conference attendees attend even one additional instance of the same
+Figure~\ref{fig:confs} illustrate this difficulty.  The top diagram shows an
+estimated per-participant CO2e footprint for each instance of the four major
+SIGPLAN conferences over the past ten years (excluding a few for which we
+had trouble getting data), with larger dots representing more emissions.
+Eyeballing this diagram, it appears that carbon-conscious conference
+planners should hold all of these conferences either in the northeast US or
+in western Europe every year.  But the bottom visualization tells a
+different story.  Here, the horizontal colored bar at the bottom represents
+the continent on which each conference was held, and each vertical bar gives
+a breakdown of the participants in that conference, colored according to the
+continent they gave as their home address.  Again, from a superficial glance
+at the colors, it is clear that, though a minority come from far away, the
+majority of participants in each case are from the region where the
+conference is being held---that is, there is a large amount of local
+participation.\iflater\bcp{That point could be strengthened by adding some
+  numbers (just in the text, not a figure) about what percentage of
+  conference attendees attend even one additional instance of the same
   conference...}\fi{} From this, one might conclude that always locating
@@
 research community by discouraging participation from other parts of the
-world. 
+world---indeed, perhaps conferences should move around as much as
+possible. 
 %
@@
 These disparate conclusions suggest that reducing conference emissions will
-require genuinely hard painful compromises.  The impulse to avoid or
-postpone thinking about it is entirely understandable!
+require genuinely painful compromises.  The impulse to avoid thinking about
+it is entirely understandable!
 
@@
 
+One potential sticking point is that some strategies for reducing carbon
+footprints---such as encouraging some (or all\cite{NCN}!)  participants to
+attend remotely---will also reduce conference revenues; in the limit, this
+could significantly impact the income stream of ACM itself, which could make
+emissions reduction politically unpalatable unless ACM's conference-focused
+business model is also adjusted.  Planning for this eventuality should start
+now.
+
+% \SECTION{Prepare for Deeper Changes}
+
+% Our two concrete proposals aim to make conferences much more
+% carbon-efficient, allowing ACM to play its part in the 40\% reduction in
+% emissions that is needed over the next decade.  But this is just a first
+% step.  Looking further ahead, the challenge will be to eliminate {\em all}
+% greenhouse emissions by around 2050 and then move into global negative
+% emissions during the second half of the century.
+
+% \bcp{I wish we could put a paragraph here about the long-term picture and
+%   the need to rethink scientific exchange and ACM's business model.  That
+%   will please our grumpy reviewer, and I think it's actually a point worth
+%   making.\cite{NCN}}
+
 \SECTION{Conclusion}
@@
 % occasions.
-
-\bcp{I wish we could put a paragraph here about the long-term picture and
-  the need to rethink scientific exchange and ACM's business model.  That
-  will please our grumpy reviewer, and I think it's actually a point worth
-  making.\cite{NCN}}
 

### 2019-06-23 main.tex
Nit

@@
 %
-Thus, our second proposal is that {\bf ACM should impose a a surcharge on
+Thus, our second proposal is that {\bf ACM should impose a surcharge on
   conferences based on their carbon footprint.}  The charge should be low at

### 2019-06-23 reviews-plus-responses.txt
More polishing

   06-Mar-2019

   Dear Dr. Pierce:

   Manuscript ID CACM-18-12-3311 entitled "ACM and Climate Change" which you submitted to the Communications of the ACM, has been reviewed.  The comments from reviewer(s) are included at the bottom of this letter.

   In view of the comments from the reviewer(s), I must decline the manuscript for publication in Communications of the ACM at this time.  However, a new manuscript may be submitted which takes into consideration these comments.

   Please note that resubmitting your manuscript does not guarantee eventual acceptance, and that your resubmission will be subject to re-review by the reviewer(s) before a decision is rendered.

   You will be unable to make your revisions on the originally submitted version of your manuscript.  Instead, revise your own copy of the manuscript and resubmit upon completion.

   Once you have revised your manuscript, go to https://mc.manuscriptcentral.com/cacm and login to your Author Center.  Click on "Manuscripts with Decisions," and then click on "Create a Resubmission" located next to the manuscript number.  Then, follow the steps for resubmitting your manuscript.

   Because we are trying to facilitate timely publication of manuscripts submitted to the Communications of the ACM, your revised manuscript should be uploaded as soon as possible.  If it is not possible for you to submit your revision within a reasonable amount of time, we will consider your paper as a new submission. Please see deadlines on your ManuscriptCentral Author Dashboard. If you need a slight extension, please contact the Editor-in-Chief.

   I look forward to a resubmission.

   Sincerely,
   Andrew A. Chien
   William Eckhardt Professor, The University of Chicago
   Editor-in-Chief, Communications of the ACM

   Co-Chair: Co-Chair, Viewpoints
   Comments to the Author:
   I concur with the associate editor's recommendation.  The piece addresses an important issue, but can be strengthened by considering the reviewers' comments and resubmitting.

   Associate Editor: Bishop, Judith
   Comments to the Author:
   Both eminent reviewers agree that the fundamental issue of the current emphasis on conferences as publication venues was side-stepped. It has to be made a priority, or else such a Viewpoint, published by the ACM, is misleading and counter-productive for combating climate change.

We didn't end up putting this front and center, but we do make a pretty
strong point about it in the penultimate paragraph.

   I am very keen to publish a Viewpoint from your committee, but trust that you can give due attention to the reviews before a re-submission. You might also like to look at the editorial in the latest issue of CACM which indicates how important this matter is to the Association.

   Reviewer: 1

   Recommendation: Minor Revision

   Comments:

   This short paper clearly identifies the issues and possible solutions to reduce the carbon footprint of ACM conferences. It is well written and clear.

   Regarding suggested solutions, here is my feedback:

   - Choose conference locations … centre of mass …”. This would result into re-enforcing communities that are already dominating and preventing others, often from countries with less research funding, to participate. 
   - Colocation: It is difficult for many people to stay away for extended periods of time and therefore to attend two or more conferences in a sequence. 
   - Conference taking place in more than one location: One of the main — if not the main benefits— of conferences is networking. This would result into a split of research communities according to geographical location. 
   - “Journal-first initiatives, which make conference attendance
   optional, can also help.” I don’t understand this point. This initiative actually increase traveling to conferences, though it has many benefits. 
   - “questioning some fundamental assumptions …” This is actually a very good point. The current emphasis on conferences as publication venues is wrong, inconsistent with other fields, and this has directly led to an inflation of conferences and traveling. I would pursue this avenue in priority. What is the really role of conferences in CS, do we need all of them?

   Additional Questions:

   Reviewer: 2

   Recommendation: Major Revision

   Comments:
   This paper argues that ACM conferences represent a significant carbon footprint, which cannot easily be reduced without substantially changing how conferences are run, as there are no low-carbon alternatives for long distance travel. The paper sets out some suggestions for what to do about this, and invites a broader discussion.

   I completely agree with the argument that the ACM conference model needs to be re-thought, and hence I strongly support the publication of a viewpoint on this topic, to challenge the community to grapple with this challenge.

   Unfortunately, the current manuscript suffers from many flaws, which need to be addressed before publication. Most importantly, the urgency of the problem set out in the first few paragraphs is somewhat undermined by a series of weak, incrementalist suggestions, none of which really get to the heart of the problem. I'm reminded of the countless articles that exhort people to deal with climate change by changing their lightbulbs and recycling more. Quite frankly, I thought we'd moved on beyond trying to solve climate change by encouraging individuals to make different choice. Such suggestions don't work, because the systemic conditions that drive those choices continually reward the behaviours we are asking people to stop doing. Put simply, ACM members fly all over the world to conferences because it's essential to their careers. You cannot stop that without a deeper systemic analysis of the perverse incentives that drive those behaviours.

   The "solutions" presented in this article all smack of timid incrementalism. If you read the IPCC reports carefully, it becomes clear that the problem isn't finding ways to cut emissions on the activities we currently carry out. The real challenge is to completely eliminate all CO2 emissions by around 2050-2060, and then moving into global negative emissions during the second half of the century. Every tonne of CO2 emitted between now and then matters. Small reductions in emissions won't get us where we need to be.

   So let's look at the suggestions made in this paper, starting with carbon offsets. Carbon offsets are nothing short of a scam. Climate scientist Kevin Anderson describes it as "worse than doing nothing", as it almost certainly causes an increase in global emissions. See for example, Anderson, 2012, The inconvenient truth of carbon offsets. Nature, 484, 7 (05 April 2012). Building these into ACM conference fees means we'll all be paying to make the problem worse. I don't believe offsets have any place in a serious discussion of action on climate change.

We respectfully disagree here.

   Then there's the suggestion to explore remote attendance. It's naive to suggest this without discussing the revenue model for ACM conferences. Conference chairs are strongly motivated by ACM’s conference budget process to maximize onsite attendance (full registration fees), and ACM's policies currently penalize authors who can’t attend to present their work, for example with the threat of having their papers excluded from the proceedings. Further, it’s highly likely that many ACM members only renew each year to get discounted conference registration. ACM will not support attempts to reduce conference attendance while its revenue model depends on maximizing such attendance. So a serious look at ACM's revenue model and conference planning policies are needed here.

   A notable absence in the list of suggestions is a complete switch away from conferences as the default venue for peer reviewed publication in computer science. Very few other fields use conferences in the same way we do - most fields use journal publication exclusively, with no peer reviewed conferences at all - conferences are then used just for the networking role, with typically a single large conference once per year in each field. Doing this would entirely eliminate the need to travel to many different conferences over the course of each year (and incidentally, would also solve many other problems associated with conference peer review - see DOI:10.1145/1531793.1531815). The article does mention journal-first tracks, which are a baby step in this direction. But if we're serious about tackling climate change, sooner or later we have to find a way to completely decouple peer reviewed publication from the need to travel.

   The ACM is also uniquely placed to explore technologies that would help other scholarly communities cut their travel carbon footprints. For example, the ACM could work with a number of conference venues around the world to build remote connection facilities into convention centres to enable federated conferences, where participants only travel to a local centre, but still gain many of the intangible benefits of networking with their global community. Such technologies must move beyond the idea of webcasting and remote talk presentation, to technologies that support the informal, unplanned interactions that constitute the main benefit of conference attendance.

   So, overall, while I welcome the publication of a viewpoint on this, I would prefer to see something that really gets to grips with the scale and urgency of the problem, and the systemic incentives that drive the ACM community to run and attend ever more conferences. The current article merely pushes dubious and ineffective non-solutions.

   In addition to these comments, the figures in the paper are all problematic:

   Figure 1 is a neat visualization, but without a proper key, it's fairly meaningless. Several other time-series representations of the same data from the IPCC reports would be much better choices.

   Figure 2 seems like a waste of space - do we really need to see a book cover?

   Figure 3 is interesting, but again lacks a key, so it's not clear what the colours represent. And even with a key, the coloured dots will be hard to see. Surely there's a better way to visualize this?

   Additional Questions:

### 2019-06-23 main.tex
More polishing

@@
 participation, increasing registration fees, soliciting additional corporate
-sponsorship, or other, more creative, means.  \iflater\mwh{Problem: We
-  showed just above that it's hard to figure out what to do. Come back to
-  this point here and say that the surcharge will spur bold choices?  Maybe
-  say that these choices won't necessarily be bad, but could be really
-  great?}\fi
+sponsorship, or other means.  \iflater\mwh{Problem: We showed just above
+  that it's hard to figure out what to do. Come back to this point here and
+  say that the surcharge will spur bold choices?  Maybe say that these
+  choices won't necessarily be bad, but could be really great?}\fi
 
@@
 imagine many good uses, including carbon offsetting, supporting ``green''
-computing research\iflater\bcp{can we mention some specifics?}\fi, and
+computing research (for example, what if ACM funded a major cross-cutting
+research initiative specifically aimed at understanding how to best replace
+or approximate the socializing and networking aspects of conferences in a
+virtual setting?\iflater\bcp{I'm tempted to make this into a third
+  proposal!}\fi) \iflater\bcp{can we mention some more specifics?}\fi, and
 defraying the costs of virtualizing conferences (livestreaming, etc.).
@@
 
-One potential sticking point is that some strategies for reducing carbon
-footprints---such as encouraging some (or all\cite{NCN}!)  participants to
-attend remotely---will also reduce conference revenues; in the limit, this
-could significantly impact the income stream of ACM itself, which could make
-emissions reduction politically unpalatable unless ACM's conference-focused
-business model is also adjusted.  Planning for this eventuality should start
-now.
+Beyond encouraging conferences to trim their emissions, putting a concrete,
+visible price on 
+carbon will stimulate creative rethinking of the conference model
+itself---for example, seriously considering alternatives such as
+rapid-turnaround journal-only publishing models, once-a-year
+megaconferences, and entirely virtual conferences~\cite{NCN}.
+%
+(A potential sticking point here is that some of these strategies will also
+significantly reduce conference revenues; in the limit, this may impact the
+income stream of ACM itself, which could make emissions reduction
+politically unpalatable unless ACM's conference-focused business model is
+also adjusted.  Planning for this eventuality should begin now!)
 

### 2019-06-24 main.tex
Pass.

@@
 publicly report its carbon footprint.} Most conferences'
-footprints will be dominated by participants air travel, but the accounting
+footprints will be dominated by participants' air travel, but the accounting
 should go beyond this to include air travel to in-person PC meetings and
@@
 committee meetings, joining a trend among other SIGs, most of which now use
-virtual PC meetings for some or all of their flagship conferences. It has
+virtual PC meetings for their flagship conferences. SIGPLAN's accounting has
 also prompted several conferences to increase their investments in
@@
   should be considered as a useful short-term expedient, to be considered in
-  parallel with, not instead of, more difficult cutting.}
+  parallel with, not instead of, more difficult cutting.} \mwh{Should
+  we mention the FCRC numbers here?}
 
@@
 %
-Organizers and attendees will be reticent to make changes that might impact
-these benefits.
+Organizers and attendees will be reticent to make particular changes
+that might threaten these benefits.
 
@@
 
-The two visualizations of registration data from some recent conferences in
-Figure~\ref{fig:confs} illustrate this difficulty.  The top diagram shows an
+To illustrate the challenge, consider the goal of choosing a conference
+location that minimizes CO2e emissions from participants' travel. 
+Using recent conference registration data,
+Figure~\ref{fig:confs} relates a conference's location 
+to where its participants are coming from.
+%
+The top diagram shows an
 estimated per-participant CO2e footprint for each instance of the four major
@@
 changes without specifying exactly which, allowing for creative and
-efficient responses. (Continuing the junk-food analogy, some municipalities
+efficient responses. Continuing the junk-food analogy, some municipalities
 such as New York city and Philadelphia have imposed a per-calorie tax on
 soft drinks, with some studies finding that doing so indeed reduced
-consumption~\cite{sodatax}.)
+consumption~\cite{sodatax}.
 
@@
 particular, as several companies have shown, even an artificial ``internal
-price'' on carbon creates a significant price signal that shapes behavior.
+price'' on carbon sends a significant price signal that shapes behavior.
 %
 Thus, our second proposal is that {\bf ACM should impose a surcharge on
-  conferences based on their carbon footprint.}  The charge should be low at
-the beginning and increase steadily (and predictably) year on year.
+  conferences based on their carbon footprint.}  The charge should be low to
+start and increase steadily (and predictably) year on year.
 %
@@
 What should ACM do with the funds collected from this surcharge?  One can
-imagine many good uses, including carbon offsetting, supporting ``green''
-computing research (for example, what if ACM funded a major cross-cutting
+imagine many good uses, including carbon offsetting, 
+defraying the costs of virtualizing conferences (livestreaming, etc.),
+and supporting ``green''
+computing research. For an example of the last: what if ACM funded a major cross-cutting
 research initiative specifically aimed at understanding how to best replace
 or approximate the socializing and networking aspects of conferences in a
-virtual setting?\iflater\bcp{I'm tempted to make this into a third
-  proposal!}\fi) \iflater\bcp{can we mention some more specifics?}\fi, and
-defraying the costs of virtualizing conferences (livestreaming, etc.).
-\iflater\bcp{Expand this paragraph a bit.}\fi
-
-Beyond encouraging conferences to trim their emissions, putting a concrete,
-visible price on 
-carbon will stimulate creative rethinking of the conference model
+virtual setting? We would hope that the cost of carbon pricing and the funding of
+initiatives such as this will stimulate creative rethinking of the conference model
 itself---for example, seriously considering alternatives such as
@@
 %
-(A potential sticking point here is that some of these strategies will also
+A potential sticking point here is that some of these strategies will also
 significantly reduce conference revenues; in the limit, this may impact the
@@
 politically unpalatable unless ACM's conference-focused business model is
-also adjusted.  Planning for this eventuality should begin now!)
+also adjusted.  Planning for this eventuality should begin now!
 

### 2019-06-24 main.tex
FCRC data

@@
   parallel with, not instead of, more difficult cutting.} \mwh{Should
-  we mention the FCRC numbers here?}
+  we mention the FCRC numbers here? 45 out of 2,700 used CO2e offsets}
 

### 2019-06-24 main.tex
Polishing

--- 2019-06-24 main.tex
@@
 Our first proposal is simple: {\bf Every ACM-sponsored conference should
-publicly report its carbon footprint.} Most conferences'
-footprints will be dominated by participants' air travel, but the accounting
-should go beyond this to include air travel to in-person PC meetings and
-estimated hotel and food CO2e footprints.
-
+publicly report its carbon footprint.} 
+%
 Public accounting will nudge conference organizers and attendees to change
-their behavior. By analogy, chain restaurants in the US are required by law
+their behavior. (By analogy, chain restaurants in the US are required by law
 to post calorie counts of food items on menu boards, and studies show that
 enlightened customers order, on average, up to 50 fewer calories a day as a
-result~\cite{menu}.
+result~\cite{menu}.)
 
@@
 experimenting with corporate sponsorship of bulk offsets for the whole
-conference.\footnote{We acknowledge that carbon offsets are a somewhat
+conference.\footnote{We acknowledge that carbon offsets are a 
   controversial 
@@
   carbon-offsets-are-not-our-get-out-jail-free-card}.  Besides issues with
-  verifiability and permanence, some observers worry that---especially given
-  their currently artificially low cost---they will encourage complacency,
+  verifiability and permanence, some observers worry that---especially at
+  their current artificially low cost---they will encourage complacency,
   leading organizations to postpone making the difficult changes that will
@@
   should be considered as a useful short-term expedient, to be considered in
-  parallel with, not instead of, more difficult cutting.} \mwh{Should
-  we mention the FCRC numbers here?}
+  parallel with, not instead of, more difficult cutting.}
+% \mwh{Should
+% we mention the FCRC numbers here? 45 out of 2,700 used CO2e offsets}
+% That's getting too much down into the details, I think -- also, these
+% numbers reflect a hastily organized, preliminary version of what we really
+% want to do.
 
@@
 conference, including originating city), and it should be made available in
-a central place and in a uniform format, to facilitate analysis.
+a central place and in a uniform format, to facilitate analysis.  (Most
+conferences footprints will be dominated by participants' air travel, but the
+data gathered should go beyond this to include air travel to in-person PC
+meetings and estimated hotel and food CO2e footprints.)
 
 ACM should provide tools for gathering and publicizing this data. As a
-start, SIGPLAN recently worked with ACM's SIG Governing Board to develop an
+first step, SIGPLAN recently worked with ACM's SIG Governing Board to develop an
 air-travel-focused CO2e calculator for conferences. Users can upload
@@
 What should ACM do with the funds collected from this surcharge?  One can
-imagine many good uses, including carbon offsetting, 
-defraying the costs of virtualizing conferences (livestreaming, etc.),
-and supporting ``green''
-computing research. For an example of the last: what if ACM funded a major cross-cutting
-research initiative specifically aimed at understanding how to best replace
-or approximate the socializing and networking aspects of conferences in a
-virtual setting? We would hope that the cost of carbon pricing and the funding of
-initiatives such as this will stimulate creative rethinking of the conference model
+imagine many good uses, including carbon offsetting, defraying the costs of
+virtualizing conferences (livestreaming, etc.), and supporting ``green''
+computing research. As an example of the last, ACM could help fund a major
+cross-cutting research initiative specifically aimed at understanding how to
+best replace or approximate the socializing and networking aspects of
+conferences in a virtual setting?
+
+Beyond simple belt-tightening, carbon pricing and the funding of initiatives
+such as this should stimulate creative rethinking of the conference model
 itself---for example, seriously considering alternatives such as
@@
 %
-A potential sticking point here is that some of these strategies will also
+A potential sticking point, however, is that some of these strategies will also
 significantly reduce conference revenues; in the limit, this may impact the

### 2019-06-24 main.tex
Polishing

@@
   14, in Gothenburg, Sweden) represents .9 tons of CO2e per participant; the
-  largest dot (ICFP 16, in Nara Japan) represents \bcp{1.94} tons per
+  largest dot (ICFP 16, in Nara Japan) represents 1.94 tons per
   participant.) Bottom: Breakdown of continent-of-origin for participants in

### 2019-06-24 main.tex
Small fixes

@@
 Panel on Climate Change (IPCC) says a 40\% decrease in
-emissions is needed by 2030 to avoid irreversible damage~\cite{IPCC-report}.
+emissions is needed by 2030 to avoid irreversible damage~\cite{IPCC18}.
 % (Figure~\ref{fig:IPCC-cover}).
@@
 Special Interest Group on Programming Languages) convened an ad hoc Climate
-Committee\cite{ClimateCommittee} to research this question.  After
+Committee~\cite{ClimateCommittee} to research this question.  After
 investigating many options~\cite{ClimateCommitteReport} and discussing
@@
 other SIG leaders, we have two concrete proposals. First, all ACM
-conferences should publicly {\em account} for CO2e emitted as a result of
+conferences should publicly {\em account} for {\COtwoE} emitted as a result of
 putting them on. Second, ACM should put a \emph{price} on carbon in
@@
 
-As a case in point, SIGPLAN has been considering how to reduce CO2e usage
-over the past couple of years, informed by an accounting of its own CO2e
+As a case in point, SIGPLAN has been considering how to reduce {\COtwoE} usage
+over the past couple of years, informed by an accounting of its own {\COtwoE}
 expenditures. This discussion has led the organizers of two of its flagship
@@
 
-SIGPLAN has also explored how to mitigate a conference's CO2e emissions by
+SIGPLAN has also explored how to mitigate a conference's {\COtwoE} emissions by
 purchasing \emph{carbon offsets}~\cite{CarbonOFfsetReport}.
@@
 activity that permanently avoids emitting some amount of greenhouse gases.
-The veracity and permanence of this CO2e emission avoidance is certified by
+The veracity and permanence of this {\COtwoE} emission avoidance is certified by
 a watchdog organization. (Typically, planting trees is not a certifiable
@@
 % \mwh{Should
-% we mention the FCRC numbers here? 45 out of 2,700 used CO2e offsets}
+% we mention the FCRC numbers here? 45 out of 2,700 used {\COtwoE} offsets}
 % That's getting too much down into the details, I think -- also, these
@@
 data gathered should go beyond this to include air travel to in-person PC
-meetings and estimated hotel and food CO2e footprints.)
+meetings and estimated hotel and food {\COtwoE} footprints.)
 
@@
 first step, SIGPLAN recently worked with ACM's SIG Governing Board to develop an
-air-travel-focused CO2e calculator for conferences. Users can upload
-conference registration data, and the calculator will estimate the CO2e cost
+air-travel-focused {\COtwoE} calculator for conferences. Users can upload
+conference registration data, and the calculator will estimate the {\COtwoE} cost
 of air travel for those sufficiently far away. This calculator could be
@@
 
-While publicly accounting for CO2e expenditures will incentivize some CO2e
+While publicly accounting for {\COtwoE} expenditures will incentivize some {\COtwoE}
 reductions, it is not likely, by itself, to induce major changes.
@@
     better version of this one from Crista.}\fi.  (The smallest dot (ICFP
-  14, in Gothenburg, Sweden) represents .9 tons of CO2e per participant; the
+  14, in Gothenburg, Sweden) represents .9 tons of {\COtwoE} per participant; the
   largest dot (ICFP 16, in Nara Japan) represents 1.94 tons per
@@
 To illustrate the challenge, consider the goal of choosing a conference
-location that minimizes CO2e emissions from participants' travel. 
+location that minimizes {\COtwoE} emissions from participants' travel. 
 Using recent conference registration data,
@@
 The top diagram shows an
-estimated per-participant CO2e footprint for each instance of the four major
+estimated per-participant {\COtwoE} footprint for each instance of the four major
 SIGPLAN conferences over the past ten years (excluding a few for which we
@@
 emissions~\cite{carbonprice}. Doing so makes manifest the hidden
-environmental cost of CO2e emissions, thus incentivizing CO2e-reducing
+environmental cost of {\COtwoE} emissions, thus incentivizing {\COtwoE}-reducing
 changes without specifying exactly which, allowing for creative and
@@
 Ideally, at some point, governments will impose carbon pricing uniformly,
-and all CO2e-intensive activities will have to pay it. But ACM can send a
+and all {\COtwoE}-intensive activities will have to pay it. But ACM can send a
 strong message about the importance of this issue by acting now. In
@@
 
-\newpage
+% \newpage
 
-\section*{Old text (that may be useful to mine)}
+% \section*{Old text (that may be useful to mine)}
 
-\SECTION{Putting a Price on Carbon}
+% \SECTION{Putting a Price on Carbon}
 
-Exactly how the price of carbon
-should be set is a challenging policy question, but it has a clear goal: It
-should be chosen so that the world arrives at zero emissions within the
-timeframe that climate scientists tell us is safe.
+% Exactly how the price of carbon
+% should be set is a challenging policy question, but it has a clear goal: It
+% should be chosen so that the world arrives at zero emissions within the
+% timeframe that climate scientists tell us is safe.
 
-To be effective in the long term, a carbon price needs to be uniform across
-society---i.e., it should be imposed by governments.  Indeed, some 40
-countries and more than 20 cities and states have already adopted
-carbon pricing policies~\cite{pricingcarbon}; more will follow.
-%
-Bottom line: We are entering an era in which carbon will be expensive.
+% To be effective in the long term, a carbon price needs to be uniform across
+% society---i.e., it should be imposed by governments.  Indeed, some 40
+% countries and more than 20 cities and states have already adopted
+% carbon pricing policies~\cite{pricingcarbon}; more will follow.
+% %
+% Bottom line: We are entering an era in which carbon will be expensive.
 
-Anticipating this era, what should we rationally do now to ensure that 
-computer-science research continues to thrive in a world
-where air travel is more expensive?  Believing that it needs to come quickly
-to avert disaster, what can we do to accelerate the process? \mwh{The
-  question now seems to have changed. The paper started by saying ACM has an
-  ``especially important role to play,'' but now seems to suggest it should
-  be reacting to an inevitable/coming future. These questions probably need
-  to go much earlier, and the earlier text reworked around them.}
-\bcp{I tried rewording the opening of the second paragraph to address this.}
+% Anticipating this era, what should we rationally do now to ensure that 
+% computer-science research continues to thrive in a world
+% where air travel is more expensive?  Believing that it needs to come quickly
+% to avert disaster, what can we do to accelerate the process? \mwh{The
+%   question now seems to have changed. The paper started by saying ACM has an
+%   ``especially important role to play,'' but now seems to suggest it should
+%   be reacting to an inevitable/coming future. These questions probably need
+%   to go much earlier, and the earlier text reworked around them.}
+% \bcp{I tried rewording the opening of the second paragraph to address this.}
 
-\SECTION{Starting Now}
+% \SECTION{Starting Now}
 
-One possible short-term response to both questions is to use {\em
-  carbon offsets} to voluntarily impose a price on our own
-emissions~\cite{CarbonOFfsetReport}.
-%
-In a nutshell, the idea is that a conference raise funds from
-registration fees or corporate sponsors and donates them, via a carbon offset
-{\em vendor}, to a {\em provider} who  uses them to support an activity
-(e.g., installing methane capture devices on landfill sites or buying
-fuel-efficient stoves to replace open-fire cooking in poor communities) that
-is {\em certified} (by yet another organization) to permanently avoid
-emitting an amount of greenhouse gases equivalent to the emissions from
-participants' travel to the conference.
-%
-Many organizations (including companies such as Google, Dell, Microsoft,
-General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
-and a number of academic societies) are already buying carbon offsets, and
-some ACM conferences have already started to experiment with voluntary
-carbon offset fees at registration time.
+% One possible short-term response to both questions is to use {\em
+%   carbon offsets} to voluntarily impose a price on our own
+% emissions~\cite{CarbonOFfsetReport}.
+% %
+% In a nutshell, the idea is that a conference raise funds from
+% registration fees or corporate sponsors and donates them, via a carbon offset
+% {\em vendor}, to a {\em provider} who  uses them to support an activity
+% (e.g., installing methane capture devices on landfill sites or buying
+% fuel-efficient stoves to replace open-fire cooking in poor communities) that
+% is {\em certified} (by yet another organization) to permanently avoid
+% emitting an amount of greenhouse gases equivalent to the emissions from
+% participants' travel to the conference.
+% %
+% Many organizations (including companies such as Google, Dell, Microsoft,
+% General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
+% and a number of academic societies) are already buying carbon offsets, and
+% some ACM conferences have already started to experiment with voluntary
+% carbon offset fees at registration time.
 
-Carbon offsets also raise some significant concerns.
+% Carbon offsets also raise some significant concerns.
+% % %
+% % A superficial one is that most organizations do not currently have
+% % accounting procedures in place to deal with them, and the piecemeal
+% % political and bureaucratic work of designing and implementing a separate
+% % policy for each organization can consume significant energy.
 % %
-% A superficial one is that most organizations do not currently have
-% accounting procedures in place to deal with them, and the piecemeal
-% political and bureaucratic work of designing and implementing a separate
-% policy for each organization can consume significant energy.
-%
-One is that the long-term climate benefit of some offsetting strategies (for
-example, paying people to plant trees) is controversial: it is important to
-choose reputable vendors and projects (such as methane capture) whose
-net benefit is clear.
-%
-A more fundamental problem is that it is hard to decide how
-much one should pay to offset a ton of carbon: current prices vary
-around an average of around \$15 per ton, depending on the vendor, the
-nature of the offset project, the region of the world where it will be
-implemented, etc.  Moreover, in order for offsets (or any other form of a
-carbon pricing) to fully motivate the needed changes in behavior, the price
-may need to go much higher---the true ``social cost'' of a ton of carbon is
-variously estimated at anywhere from \$40 to \$400 per
-ton~\bcp{citation}---and setting it too low may encourage complacency.
-%
-While acknowledging these concerns, we believe that voluntary
-carbon offsetting can be an effective near-term tool for raising awareness
-and beginning to shift organizational behavior because it associates
-emissions with a number, giving decision makers within the organization
-something concrete to optimize.
+% One is that the long-term climate benefit of some offsetting strategies (for
+% example, paying people to plant trees) is controversial: it is important to
+% choose reputable vendors and projects (such as methane capture) whose
+% net benefit is clear.
+% %
+% A more fundamental problem is that it is hard to decide how
+% much one should pay to offset a ton of carbon: current prices vary
+% around an average of around \$15 per ton, depending on the vendor, the
+% nature of the offset project, the region of the world where it will be
+% implemented, etc.  Moreover, in order for offsets (or any other form of a
+% carbon pricing) to fully motivate the needed changes in behavior, the price
+% may need to go much higher---the true ``social cost'' of a ton of carbon is
+% variously estimated at anywhere from \$40 to \$400 per
+% ton~\bcp{citation}---and setting it too low may encourage complacency.
+% %
+% While acknowledging these concerns, we believe that voluntary
+% carbon offsetting can be an effective near-term tool for raising awareness
+% and beginning to shift organizational behavior because it associates
+% emissions with a number, giving decision makers within the organization
+% something concrete to optimize.
 
-\SECTION{Grounding the Discussion}
+% \SECTION{Grounding the Discussion}
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=6in]{SIGPLAN-confs.png}
-\includegraphics[width=6in]{ParticipantsOriginAll.png}
-\caption{Top: carbon footprint per participant for travel to recent SIGPLAN
-  conferences\iflater\bcp{Get better file from the other paper, or get
-    better version of this one from Crista.}\fi Bottom: origin of participants for each conference.} 
-\label{fig:confs}
-\end{figure}
+% \begin{figure}[t]
+% \centering
+% \includegraphics[width=6in]{SIGPLAN-confs.png}
+% \includegraphics[width=6in]{ParticipantsOriginAll.png}
+% \caption{Top: carbon footprint per participant for travel to recent SIGPLAN
+%   conferences\iflater\bcp{Get better file from the other paper, or get
+%     better version of this one from Crista.}\fi Bottom: origin of participants for each conference.} 
+% \label{fig:confs}
+% \end{figure}
 
-What, concretely, would it mean to ``optimize'' the carbon footprint of
-conferences?
-%
-% Beyond imposing a price on our carbon footprint (or waiting for one
-% to be imposed from outside), a sane response to climate change should
-% clearly also include taking steps to reduce the size of this
-% footprint.
-% \mwh{Don't like previous sentence. It seems to have gotten away from
-%   the point of the paper, which is using price to change balance of
-%   apples to oranges. Analyzing data is useful for figuring out ACM's
-%   internal price, not tweaking around the edges, which is how it reads
-%   below.} 
-One might contemplate a wide array of options~\cite{ClimateCommitteReport},
-from easy-but-incremental changes like making program committee meetings
-virtual instead of physical and enabling virtual participation via
-live-streaming, all the way to complete re-imagining of conferences and
-their role in science (holding conferences completely in virtual spaces,
-replacing conferences with journals, etc.), with some intriguing hybrid
-forms in the middle (regional conferences, mega-conferences, conferences
-held simultaneously at two locations on different continents, etc.).
-%
-How much of a difference would these make to emissions?
+% What, concretely, would it mean to ``optimize'' the carbon footprint of
+% conferences?
+% %
+% % Beyond imposing a price on our carbon footprint (or waiting for one
+% % to be imposed from outside), a sane response to climate change should
+% % clearly also include taking steps to reduce the size of this
+% % footprint.
+% % \mwh{Don't like previous sentence. It seems to have gotten away from
+% %   the point of the paper, which is using price to change balance of
+% %   apples to oranges. Analyzing data is useful for figuring out ACM's
+% %   internal price, not tweaking around the edges, which is how it reads
+% %   below.} 
+% One might contemplate a wide array of options~\cite{ClimateCommitteReport},
+% from easy-but-incremental changes like making program committee meetings
+% virtual instead of physical and enabling virtual participation via
+% live-streaming, all the way to complete re-imagining of conferences and
+% their role in science (holding conferences completely in virtual spaces,
+% replacing conferences with journals, etc.), with some intriguing hybrid
+% forms in the middle (regional conferences, mega-conferences, conferences
+% held simultaneously at two locations on different continents, etc.).
+% %
+% How much of a difference would these make to emissions?
 
-% But before undertaking significant changes, we need to understand the
-% situation better.  For most of these ideas, it is hard to confidently
-% predict either the potential benefits (in terms of reduced emissions) or the
-% potential costs (to scientific exchange, career development of young
-% scientists, etc.).  
-A preliminary analysis of registration data from
-a number of SIGPLAN conferences paints a rather complicated picture.
-%
-\bcp{Put the discussion of Figure 1 (and an accompanying figure
-showing distance traveled to several instances of the same conference) here;
-this illustrates the difficulty and the need for careful data gathering and
-analysis.}
+% % But before undertaking significant changes, we need to understand the
+% % situation better.  For most of these ideas, it is hard to confidently
+% % predict either the potential benefits (in terms of reduced emissions) or the
+% % potential costs (to scientific exchange, career development of young
+% % scientists, etc.).  
+% A preliminary analysis of registration data from
+% a number of SIGPLAN conferences paints a rather complicated picture.
+% %
+% \bcp{Put the discussion of Figure 1 (and an accompanying figure
+% showing distance traveled to several instances of the same conference) here;
+% this illustrates the difficulty and the need for careful data gathering and
+% analysis.}
 
-All parts of ACM should be doing this kind of analysis.  (The acm-climate
-mailing list is one place to exchange ideas about this.)
+% All parts of ACM should be doing this kind of analysis.  (The acm-climate
+% mailing list is one place to exchange ideas about this.)
 
-\SECTION{Plan Ahead}
+% \SECTION{Plan Ahead}
 
-Looking further ahead, the expectation of expensive carbon invites a deeper
-rethinking of how academic computer science is done.  
-%
-\bcp{
-\begin{itemize}
-\item Add a bit of meat about how the current economic model may be
-  unsustainable and reworking it will involve substantial organizational
-  changes.  (There may be some good words that we can paraphrase from our
-  second reviewer!)
-\item monitoring and mitigating carbon costs should be an ACM-level activity
-\end{itemize}
-}
+% Looking further ahead, the expectation of expensive carbon invites a deeper
+% rethinking of how academic computer science is done.  
+% %
+% \bcp{
+% \begin{itemize}
+% \item Add a bit of meat about how the current economic model may be
+%   unsustainable and reworking it will involve substantial organizational
+%   changes.  (There may be some good words that we can paraphrase from our
+%   second reviewer!)
+% \item monitoring and mitigating carbon costs should be an ACM-level activity
+% \end{itemize}
+% }
 
-The climate crisis is too large and too urgent to leave to world leaders to
-address at their own pace: organizations at every scale should confront
-their own contributions, raise awareness among their members, and begin
-establishing new ways of doing business in the lower-carbon future that is
-coming.
-% For ACM, this especially means confronting the carbon-intensive
-% nature of our conferences.
-% %
-% You can help accelerate this process in a number of ways: 
+% The climate crisis is too large and too urgent to leave to world leaders to
+% address at their own pace: organizations at every scale should confront
+% their own contributions, raise awareness among their members, and begin
+% establishing new ways of doing business in the lower-carbon future that is
+% coming.
+% % For ACM, this especially means confronting the carbon-intensive
+% % nature of our conferences.
+% % %
+% % You can help accelerate this process in a number of ways: 
+% % \begin{itemize}
+% % \item {\bf Get involved} in raising awareness about this issue in the
+% % organizations you are a part of.  \bcp{Does the rest of this smack of
+% %   incrementalism?  Maybe delete it.  ``For example, ask the organizers of
+% %   the conferences you attend what they are doing about reducing or
+% %   offsetting emissions from participants' travel.''}
+% % \item {\bf Connect} with others that are engaging with these issues---in
+% % your workplace, in your other communities, and within ACM (e.g., on the
+% % {\tt acm-climate} mailing list~\cite{acm-climate}).
+% % \item {\bf Analyze} your organization's carbon-emitting activities to
+% % understand how to reduce its footprint now and prepare for the coming era of
+% % expensive carbon.
+% % \end{itemize}
+
+% \bcp{New outline:
 % \begin{itemize}
-% \item {\bf Get involved} in raising awareness about this issue in the
-% organizations you are a part of.  \bcp{Does the rest of this smack of
-%   incrementalism?  Maybe delete it.  ``For example, ask the organizers of
-%   the conferences you attend what they are doing about reducing or
-%   offsetting emissions from participants' travel.''}
-% \item {\bf Connect} with others that are engaging with these issues---in
-% your workplace, in your other communities, and within ACM (e.g., on the
-% {\tt acm-climate} mailing list~\cite{acm-climate}).
-% \item {\bf Analyze} your organization's carbon-emitting activities to
-% understand how to reduce its footprint now and prepare for the coming era of
-% expensive carbon.
+% \item Accountability.
+%   \begin{itemize}
+%   \item Proposal: Every ACM-sponsored conference should be required to
+%   publicly report its carbon footprint (total and per-participant).
+%   (Including PC meetings, participant travel, maybe hotel?).
+%   \item ACM needs to provide tools to help conferences do this.  (SIGPLAN
+%   built one.)
+%   \item Benefits:
+%   \begin{itemize}
+%   \item Encourages people to pay attention, which will lead to some changes
+%   in behavior all by itself. 
+%   \item Generates data that will be needed for grounding discussions of how
+%   optimize footprint.  E.g., SIGPLAN has mostly gone to virtual PC
+%   meetings.  Livestreaming.
+%   \item Indeed, optimizing footprint is not so easy.  (Describe some of our
+%   data.)  More data and thorough analysis is needed.
+%   \item Segue: Colocation, etc.
 %   \end{itemize}
+%   \end{itemize}
+% \item Pricing carbon.
+%   \begin{itemize}
+%   \item To motivate more painful changes and force people to make decisions,
+%   we need a price on carbon.
+%   \item Ultimately, we want/expect governments to do this.  But we don't need to
+%   wait. 
+%   \item Proposal: ACM should apply downward pressure on conferences by
+%   charging conferences a surcharge based on their carbon footprint.
+%   \item This money can be used in many ways: buying carbon offsets,
+%   supporting ``green research'', supporting virtualizing conferences
+%   (livestreaming, etc.), supporting a transition to a low-carbon / few
+%   conferences business model for ACM.
+%   \end{itemize}
+% \end{itemize}
+% }
 
-\bcp{New outline:
-\begin{itemize}
-\item Accountability.
-  \begin{itemize}
-  \item Proposal: Every ACM-sponsored conference should be required to
-  publicly report its carbon footprint (total and per-participant).
-  (Including PC meetings, participant travel, maybe hotel?).
-  \item ACM needs to provide tools to help conferences do this.  (SIGPLAN
-  built one.)
-  \item Benefits:
-  \begin{itemize}
-  \item Encourages people to pay attention, which will lead to some changes
-  in behavior all by itself. 
-  \item Generates data that will be needed for grounding discussions of how
-  optimize footprint.  E.g., SIGPLAN has mostly gone to virtual PC
-  meetings.  Livestreaming.
-  \item Indeed, optimizing footprint is not so easy.  (Describe some of our
-  data.)  More data and thorough analysis is needed.
-  \item Segue: Colocation, etc.
-  \end{itemize}
-  \end{itemize}
-\item Pricing carbon.
-  \begin{itemize}
-  \item To motivate more painful changes and force people to make decisions,
-  we need a price on carbon.
-  \item Ultimately, we want/expect governments to do this.  But we don't need to
-  wait. 
-  \item Proposal: ACM should apply downward pressure on conferences by
-  charging conferences a surcharge based on their carbon footprint.
-  \item This money can be used in many ways: buying carbon offsets,
-  supporting ``green research'', supporting virtualizing conferences
-  (livestreaming, etc.), supporting a transition to a low-carbon / few
-  conferences business model for ACM.
-  \end{itemize}
-\end{itemize}
-}
-
+\iflater
 \bcp{Things still missing:
   \begin{itemize}
-  \item citations for the claims we make (and the carbon-offsets-are-bad paper
-  mentioned by the reviewer)
-  \item Go through all the comments and ideas below (including the CACM
-  reviews) one last time
+  \item do we need any more citations for the claims we make?
   \item Find a place for this: {\tt acm-climate} mailing list~\cite{acm-climate}
-  \item Think about adding headings, per Bishop's recommendation
-
   \item Write a cover letter
   \begin{itemize}
-  \item We tried hard to remove all incrementalism and make it much more
+  \item We tried hard to remove incrementalism and make it much more
   hard hitting
+  \item In particular, we make two concrete proposals
   \item However, we did keep the suggestion of carbon offsets, hopefully
@@
 }
+\fi
 
@@
 \item 
-Is the citation for emissions by region CO2 or CO2e?
-\item 
 Is there a ``more authoritative'' source for price of carbon offsets than

### 2019-06-25 main.tex
Nits

@@
 a central place and in a uniform format, to facilitate analysis.  (Most
-conferences footprints will be dominated by participants' air travel, but the
+conferences' footprints will be dominated by participants' air travel, but the
 data gathered should go beyond this to include air travel to in-person PC
@@
 
-ACM should provide tools for gathering and publicizing this data. As a
+ACM should provide tools to gather and publicize this data. As a
 first step, SIGPLAN recently worked with ACM's SIG Governing Board to develop an

### 2019-06-27 main.tex
Clarification
de-Macronized the second proposal

@@
 experimenting with corporate sponsorship of bulk offsets for the whole
-conference.\footnote{We acknowledge that carbon offsets are a 
+conference.  
+%
+(We acknowledge that carbon offsets are a 
   controversial 
@@
   should be considered as a useful short-term expedient, to be considered in
-  parallel with, not instead of, more difficult cutting.}
+  parallel with, not instead of, more difficult cutting.)
 % \mwh{Should
@@
   participant.) Bottom: Breakdown of continent-of-origin for participants in
-  each conference.
+  each conference.  Colored bars represent percentages of participants whose
+  home city is in each continent.
 % Bottom: Distances traveled by participants, one way, in Km; middle dashed
@@
 virtualizing conferences (livestreaming, etc.), and supporting ``green''
-computing research. As an example of the last, ACM could help fund a major
+computing research. As an example of the last, ACM could help fund a 
 cross-cutting research initiative specifically aimed at understanding how to
 best replace or approximate the socializing and networking aspects of
-conferences in a virtual setting?
+conferences in a virtual setting.  But before any of this, the first concern
+should be to find ways of {\em reducing} the finanial burden of such a
+surcharge on those who may be disproportionately affected by it---students
+without grant support, participants from developing areas, etc.  That is,
+wealthier participants should subsidize the carbon footprints of less
+fortunate ones.
 

### 2019-06-30 main.tex
Small changes. Added a bit more structure.

@@
 
-\newcommand{\SECTION}{\paragraph*}
+\newcommand{\SECTION}{\section*}
+\newcommand{\PARAGRAPH}{\paragraph*}
 
@@
 
-\SECTION{Mandate public accountability.}
+\SECTION{Mandate public accountability}
 
@@
 Public accounting will nudge conference organizers and attendees to change
-their behavior. (By analogy, chain restaurants in the US are required by law
+their behavior. By analogy, chain restaurants in the US are required by law
 to post calorie counts of food items on menu boards, and studies show that
 enlightened customers order, on average, up to 50 fewer calories a day as a
-result~\cite{menu}.)
+result~\cite{menu}.  
+%MWH: I don't know why this should be a parenthetical.
+Why would the analogy be an aside?
+It is meant to connect the reader's understanding to what we are proposing, explciitly; it is meant to be read, not skipped over.
 
@@
 
+\PARAGRAPH{Carbon Offsets}
+%
 SIGPLAN has also explored how to mitigate a conference's {\COtwoE} emissions by
@@
 be.)
-%
+
 Many organizations, including companies such as Google, Dell, Microsoft,
@@
 conference.
-%
-(We acknowledge that carbon offsets are a 
+
+We acknowledge that carbon offsets are a 
   controversial 
@@
   should be considered as a useful short-term expedient, to be considered in
-  parallel with, not instead of, more difficult cutting.)
+  parallel with, not instead of, more difficult cutting.
+%MWH: It's a bit odd to have two lengthy parenthetical asides in the same
+paragraph. I think this one is on point, so need not be an aside.
 % \mwh{Should
@@
 
+\PARAGRAPH{Data gathering}
+%
 Beyond encouraging conferences to take these sorts of simple steps, public
@@
 
-\SECTION{Put a Price on Carbon.}
+\SECTION{Put a Price on Carbon}
 
+\PARAGRAPH{Hard choices}
+%
 While publicly accounting for {\COtwoE} expenditures will incentivize some {\COtwoE}
@@
 
+\PARAGRAPH{A cost for carbon}
+%
 This dilemma is a microcosm of the one that all of society faces. To address
@@
 consumption~\cite{sodatax}.
+%MWH: The above is not a parenthetical, so the earlier junkfood analogy should not be, either
 
@@
 
+\PARAGRAPH{Precipitating change}
+ %
 What should ACM do with the funds collected from this surcharge?  One can
@@
 conferences in a virtual setting.  But before any of this, the first concern
-should be to find ways of {\em reducing} the finanial burden of such a
+should be to find ways of {\em reducing} the financial burden of such a
 surcharge on those who may be disproportionately affected by it---students
@@
 and begin establishing new ways of doing business in the lower-carbon future
-that is coming.  ACM should lead the way!
+that is coming.  ACM should play its part!
 

### 2019-06-30 main.tex
Polishing

@@
 A broad scientific consensus warns that human
-emissions of carbon dioxide and other greenhouse gases are warming the earth.
+emissions of greenhouse gases are warming the earth.
 % (Figure \ref{fig:warming}).
@@
 Achieving reductions on this scale will
-require urgent and sustained commitment at all levels of society, including
-not only city, state, and national governments, but also 
-universities, companies, and scientific societies
-like ACM.
+require urgent and sustained commitment at all levels of society---not only
+national, state, and city governments, but also 
+universities, companies, and scientific societies.
 
@@
 the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
-passenger~\cite{carboncalculator}\iflater\bcp{Trim citation?}\fi---a
-significant fraction of the {\em total} yearly emissions for an average
-resident of the US (16.5 tons) or Europe (7 tons)~\cite{emissions}.
-\iflater\bcp{For comparison, ...  (put in some figures about the footprints
-  of a few other things that people might think of reducing---driving,
-  eating meat, upgrading their fridge...)}\fi{} Moreover, these emissions
-have no near-term technological fix, since jet fuel is difficult to replace
-with renewable energy sources~\cite{elec-air}.
+passenger\iflater\bcp{Citation?}\fi, which is a significant fraction of the
+{\em total} yearly emissions for an average resident of the US (16.5 tons)
+or Europe (7 tons)~\cite{emissions}.  \iflater\bcp{For comparison, ...  (put
+  in some figures about the footprints of a few other things that people
+  might think of reducing---driving, eating meat, upgrading their
+  fridge...)}\fi{} Moreover, these emissions have no near-term technological
+fix, since jet fuel is difficult to replace with renewable energy
+sources~\cite{elec-air}.
 % \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
@@
 
-How should ACM respond to these facts?  Two years ago, SIGPLAN (ACM's
-Special Interest Group on Programming Languages) convened an ad hoc Climate
-Committee~\cite{ClimateCommittee} to research this question.  After
+How should ACM respond to these facts?  Two years ago, ACM's Special
+Interest Group on Programming Languages (SIGPLAN) convened an ad hoc Climate
+Committee to research this question~\cite{ClimateCommittee}.  After
 investigating many options~\cite{ClimateCommitteReport} and discussing
-intensively with SIGPLAN members and conference organizers, ACM staff, and
-other SIG leaders, we have two concrete proposals. First, all ACM
-conferences should publicly {\em account} for {\COtwoE} emitted as a result of
-putting them on. Second, ACM should put a \emph{price} on carbon in
-conference budgets, to create pressure on organizers to reduce their
-footprints.
+intensively with members, conference organizers, ACM staff, and leaders of
+other SIGs, we are putting forward two concrete proposals. First, all ACM
+conferences should publicly {\em account} for the {\COtwoE} emitted as a
+result of putting them on. Second, ACM \iflater\bcp{Are we really
+  talking about ACM, in the first instance?  Or is the proposal really that
+  individual SIGs should do this, until there is enough buy-in to do it at
+  the ACM level?  The first is easier to say, but the second is the only
+  plausible implementation path.  If there's space for at least a sentence
+  or two on this (and other practicalities of implementation), it might be
+  good.}\fi should put a \emph{price} on 
+carbon in conference budgets, to create pressure on organizers to reduce
+their footprints.
 
@@
 
-Our first proposal is simple: {\bf Every ACM-sponsored conference should
+Our first proposal is modest: {\bf Every ACM-sponsored conference should
 publicly report its carbon footprint.} 
@@
 result~\cite{menu}.  
-%MWH: I don't know why this should be a parenthetical.
-Why would the analogy be an aside?
-It is meant to connect the reader's understanding to what we are proposing, explciitly; it is meant to be read, not skipped over.
+%MWH: I don't know why this should be a parenthetical.  Why would the
+%analogy be an aside? It is meant to connect the reader's understanding to
+%what we are proposing, explciitly; it is meant to be read, not skipped
+%over. 
+% BCP: In general, parenthetical remarks are not meant to be skipped over:
+% they are meant to be read as asides.  I think this reads better in parens,
+% but I won't fight for it.
 
-As a case in point, SIGPLAN has been considering how to reduce {\COtwoE} usage
-over the past couple of years, informed by an accounting of its own {\COtwoE}
-expenditures. This discussion has led the organizers of two of its flagship
-conferences (POPL and ICFP) to switch from in-person to on-line program
-committee meetings, joining a trend among other SIGs, most of which now use
-virtual PC meetings for their flagship conferences. SIGPLAN's accounting has
-also prompted several conferences to increase their investments in
+As a case in point, SIGPLAN has been considering how to reduce emissions
+over the past couple of years, informed by an accounting of its own
+{\COtwoE} expenditures. This discussion has led the organizers of two of its
+flagship conferences (POPL and ICFP) to switch from in-person to on-line
+program committee meetings (joining a trend among other SIGs, most of which
+already use virtual PC meetings for their flagship conferences). This
+discussion has also prompted several conferences to increase investments in
 livestreaming and video recording to support remote participation.
 
-\PARAGRAPH{Carbon Offsets}
+\PARAGRAPH{Carbon offsets}
 %
-SIGPLAN has also explored how to mitigate a conference's {\COtwoE} emissions by
-purchasing \emph{carbon offsets}~\cite{CarbonOFfsetReport}.
+SIGPLAN has also explored how to mitigate conference emissions by purchasing
+\emph{carbon offsets}~\cite{CarbonOFfsetReport}.
 %
-A carbon offset is sold by a vendor who uses the raised funds to finance an
-activity that permanently avoids emitting some amount of greenhouse gases.
-The veracity and permanence of this {\COtwoE} emission avoidance is certified by
-a watchdog organization. (Typically, planting trees is not a certifiable
+A carbon offset is sold by a vendor who uses the funds to finance an
+activity that permanently removes or avoids emitting some amount of
+greenhouse gases.  The veracity and permanence of this activity is certified
+by a watchdog organization. (Typically, planting trees is not a certifiable
 activity, since it's hard to guarantee the trees won't be cut down, but
@@
 net footprint. ACM conferences can do the same, and the purchases can be
-included in the public accounting we are proposing.  In fact, ACM recently
+mentioned in the public accounting we are proposing.  In fact, ACM recently
 decided that conferences could offer attendees the option to pay for an
-offset during registration; the conference then uses the designated funds to
-purchase the offset from a qualified vendor.  Some conferences are also
-experimenting with corporate sponsorship of bulk offsets for the whole
-conference.
+offset during registration.  Other scientific societies are going further,
+rolling carbon offsets into the registration fees for all
+participants.\iflater\bcp{In a later draft, we may be able to say that ICML,
+NIPS, etc. are a case in point.  Ask Alexandre Lacoste about this.}\fi
 
@@
   their current artificially low cost---they will encourage complacency,
-  leading organizations to postpone making the difficult changes that will
-  be required to sustainably decarbonize society.  We believe that offsets
+  leading organizations to postpone facing the difficult changes that will
+  be required to sustainably decarbonize society.  Offsets
   should be considered as a useful short-term expedient, to be considered in
   parallel with, not instead of, more difficult cutting.
-%MWH: It's a bit odd to have two lengthy parenthetical asides in the same
-paragraph. I think this one is on point, so need not be an aside.
+% MWH: It's a bit odd to have two lengthy parenthetical asides in the same
+% paragraph. I think this one is on point, so need not be an aside.
+% BCP: The whole paragraph was in parens, before.  However, the new paragraph
+% header does a decent job of "closing the aside," so I agree the parens can
+% go.
+
 % \mwh{Should
@@
 %
-Beyond encouraging conferences to take these sorts of simple steps, public
-accounting of conference emissions will generate data that future organizers
-can use.  To this end, emissions data should be published both in aggregate
-form (total and per-participant footprint), for quick reference, and in raw
-form (a suitably anonymized record for each individual attending each
-conference, including originating city), and it should be made available in
-a central place and in a uniform format, to facilitate analysis.  (Most
+Beyond encouraging conferences to trim their emissions, public 
+accounting will generate useful data for future organizers.  To this end,
+emissions data should be published both in aggregate 
+form (total and per-participant footprints), for quick reference, and in raw
+form, with an anonymized record for each participant in each
+conference, including their originating city.  The data should be made
+available in 
+a central place, in a uniform format.  Most
 conferences' footprints will be dominated by participants' air travel, but the
 data gathered should go beyond this to include air travel to in-person PC
-meetings and estimated hotel and food {\COtwoE} footprints.)
+meetings and estimated hotel and food {\COtwoE} footprints.
 
-ACM should provide tools to gather and publicize this data. As a
-first step, SIGPLAN recently worked with ACM's SIG Governing Board to develop an
-air-travel-focused {\COtwoE} calculator for conferences. Users can upload
-conference registration data, and the calculator will estimate the {\COtwoE} cost
-of air travel for those sufficiently far away. This calculator could be
-expanded to account for other forms of conference emissions.
+ACM should develop tools to gather and publicize this data. As a first step,
+SIGPLAN recently built an air-travel-focused {\COtwoE} calculator for
+conferences\iflater~\cite{carboncalculator}\bcp{Trim citation?  Is it even
+  the right one? Link seems
+  dead!}\fi. Users can upload conference registration data, and the
+calculator will estimate the {\COtwoE} cost of air travel for those
+sufficiently far away. This calculator could be expanded to account for
+other forms of conference emissions.
 
@@
 
+While public accounting of {\COtwoE} expenditures will incentivize some 
+reductions, it is not likely, by itself, to induce major changes.
+
 \PARAGRAPH{Hard choices}
-%
-While publicly accounting for {\COtwoE} expenditures will incentivize some {\COtwoE}
-reductions, it is not likely, by itself, to induce major changes.
-%
+
 Science is a fundamentally social process, and the conference system
@@
 %
-Organizers and attendees will be reticent to make particular changes
-that might threaten these benefits.
+Organizers and attendees will be reticent to make changes that might
+threaten these benefits.
 
@@
 
-To illustrate the challenge, consider the goal of choosing a conference
-location that minimizes {\COtwoE} emissions from participants' travel. 
-Using recent conference registration data,
-Figure~\ref{fig:confs} relates a conference's location 
-to where its participants are coming from.
+To illustrate the challenges, consider the problem of choosing a conference
+location that minimizes emissions from participants' travel.  Using recent
+registration data from four SIGPLAN conferences, Figure~\ref{fig:confs}
+shows two ways of looking at the relation between locations and emissions. 
 %
 The top diagram shows an
-estimated per-participant {\COtwoE} footprint for each instance of the four major
-SIGPLAN conferences over the past ten years (excluding a few for which we
+estimated per-participant {\COtwoE} footprint for each instance of each
+conference over the past ten years (excluding a few for which we
 had trouble getting data), with larger dots representing more emissions.
-Eyeballing this diagram, it appears that carbon-conscious conference
-planners should hold all of these conferences either in the northeast US or
-in western Europe every year.  But the bottom visualization tells a
-different story.  Here, the horizontal colored bar at the bottom represents
+Eyeballing this diagram, it seems that carbon-conscious 
+organizers should hold all of these conferences in either the northeast US or
+western Europe every year.  But the bottom visualization tells a
+different story.  The horizontal colored bar at the bottom represents
 the continent on which each conference was held, and each vertical bar gives
 a breakdown of the participants in that conference, colored according to the
-continent they gave as their home address.  Again, from a superficial glance
+continent of their work address.  From a superficial glance
 at the colors, it is clear that, though a minority come from far away, the
-majority of participants in each case are from the region where the
-conference is being held---that is, there is a large amount of local
-participation.\iflater\bcp{That point could be strengthened by adding some
+majority of participants in each conference are local to the region where the
+conference is being held.\iflater\bcp{That point could be strengthened by adding some
   numbers (just in the text, not a figure) about what percentage of
   conference attendees attend even one additional instance of the same
-  conference...}\fi{} From this, one might conclude that always locating
-conferences in the same one or two places would significantly damage the
+  conference...}\fi{} This suggests that locating
+conferences every year in the same one or two places would significantly
+damage the 
 research community by discouraging participation from other parts of the
-world---indeed, perhaps conferences should move around as much as
-possible. 
+world.  Indeed, one might conclude that, from the point of view of
+strengthening the research community, conferences should move around as 
+much as possible.
 %
@@
 %
-These disparate conclusions suggest that reducing conference emissions will
+These disparate perspectives suggest that reducing conference emissions will
 require genuinely painful compromises.  The impulse to avoid thinking about
@@
 
-However, the present trajectory of world emissions is unsustainable, and
-hard choices will have to be made if ACM is to play its part by reducing its
-own emissions. How do we motivate organizers to make them?
+However, the present trajectory of world emissions is unsustainable: hard
+choices will have to be made, and soon, if ACM is to play its part by
+reducing its own emissions. How do we motivate organizers to face these
+choices?
 
@@
 %
-This dilemma is a microcosm of the one that all of society faces. To address
-it, many policy experts advocate some form of ``carbon pricing'' (either
-taxes or cap-and-trade schemes), which imposes a concrete, immediate cost on
+This dilemma is a microcosm of one faced by all of society. To address it,
+many policy experts advocate some form of carbon pricing (either taxes or
+cap-and-trade schemes), imposing a concrete, immediate cost on
 emissions~\cite{carbonprice}. Doing so makes manifest the hidden
-environmental cost of {\COtwoE} emissions, thus incentivizing {\COtwoE}-reducing
-changes without specifying exactly which, allowing for creative and
-efficient responses. Continuing the junk-food analogy, some municipalities
-such as New York city and Philadelphia have imposed a per-calorie tax on
-soft drinks, with some studies finding that doing so indeed reduced
-consumption~\cite{sodatax}.
+environmental cost of emissions, incentivizing
+{\COtwoE}-reducing changes without specifying exactly which, and thus
+allowing for creative and efficient responses. Continuing the junk-food
+analogy, some 
+municipalities such as New York city and Philadelphia have imposed a
+per-calorie tax on soft drinks, with some studies finding that doing so
+indeed reduced consumption~\cite{sodatax}.
 %MWH: The above is not a parenthetical, so the earlier junkfood analogy should not be, either
@@
   conferences based on their carbon footprint.}  The charge should be low to
-start and increase steadily (and predictably) year on year.
+start and increase steadily and predictably, year on year.
 %
-Conference organizers can then choose (on the basis of careful analysis of
-data and weighing of benefits) how best to balance their budgets---whether
+Conference organizers can then choose how best to balance their
+budgets---whether 
 by decreasing per-participant emissions, decreasing (physical)
@@
 best replace or approximate the socializing and networking aspects of
-conferences in a virtual setting.  But before any of this, the first concern
+conferences in a virtual setting.  But the very first concern
 should be to find ways of {\em reducing} the financial burden of such a
@@
 wealthier participants should subsidize the carbon footprints of less
-fortunate ones.
+wealthy ones.
 
-Beyond simple belt-tightening, carbon pricing and the funding of initiatives
-such as this should stimulate creative rethinking of the conference model
+Beyond simple belt-tightening, carbon pricing will stimulate creative
+rethinking of the conference model 
 itself---for example, seriously considering alternatives such as
-rapid-turnaround journal-only publishing models, once-a-year
+rapid-turnaround, journal-only publishing models, once-a-year
 megaconferences, and entirely virtual conferences~\cite{NCN}.
 %
-A potential sticking point, however, is that some of these strategies will also
-significantly reduce conference revenues; in the limit, this may impact the
-income stream of ACM itself, which could make emissions reduction
-politically unpalatable unless ACM's conference-focused business model is
-also adjusted.  Planning for this eventuality should begin now!
+A potential sticking point is that some of these will significantly reduce
+conference revenues, in turn impacting the income stream of ACM itself; this
+could make emissions reduction politically problematic unless ACM's
+conference-focused business model is also adjusted.  Planning for this
+eventuality should begin now!
 
@@
 must confront their own contributions, raise awareness among their members,
-and begin establishing new ways of doing business in the lower-carbon future
-that is coming.  ACM should play its part!
+and establish new ways of doing business in the lower-carbon future that is
+coming at us.  We in ACM must play our part.
 

### 2019-07-01 response.txt
Add response to reviews

Thank you for the very helpful comments!

The main feeling among the reviewers and AE was that the article did not go far enough -- that its proposals were too timid and incremental.  We took this response to heart and essentially rewrote the whole thing from top to bottom, removing most discussion of short-term mitigations and replacing it with two concrete and specific proposals for actions that ACM can and should take on a large scale -- one reasonably easy, one much more ambitious.  We have also removed the uninformative figures and completely revamped the remaining figure to make it (we hope) much more interesting.

The new version does still diverge from the recommendations of the reviewers in two significant respects:

1) Reviewer 2 felt that the main focus of the article should be on the need to move past conferences altogether.  While we agree that this is where things may need to move in the medium to long term, it is very far from where the community is now, and we feel this is for good reason: CS conferences offer a great deal of value in addition to the environmental harm that their emissions create, and it is far from clear at this point that simply deleting conferences would be a net benefit to society.  The discussion about balancing benefits and harms is going to be a long and painful one, and we do not feel ready to prejudge the outcome at this point.  What is needed is to begin the discussion, and we hope that the new version of the article helps do this.

We did add a paragraph at the end acknowledging reviewer 2's observation that ACM's business model relies heavily on conference revenue and therefore needs rethinking.

2) Reviewer 2 also strongly objected to our even mentioning carbon offsets as a potential mitigation strategy.  After much discussion among ourselves, we decided not to do this: The reviewer's position is reasonable (and held by lots of reasonable and well-informed people), but not universal, and we respectfully continue to believe (and assert in the article) that carefully chosen offsets offer a useful, short-term way to mitigate the harm we are doing right now, buying the world (and ACM) a little more time to make the deeper adjustments that we 100% agree are necessary.  However, we have greatly shrunk the discussion of offsets, and we've included a new paragraph explicitly mentioning the dangers of bogus offsets and the greater danger of encouraging complacency.  We hope this will be viewed as a reasonable compromise.

### 2019-07-01 main.tex
Add response to reviews

@@
   conferences\iflater\bcp{Get better file from the other paper, or get
-    better version of this one from Crista.}\fi.  (The smallest dot (ICFP
+    better version of this one from Crista.}\fi.  The smallest dot (ICFP
   14, in Gothenburg, Sweden) represents .9 tons of {\COtwoE} per participant; the
   largest dot (ICFP 16, in Nara Japan) represents 1.94 tons per
-  participant.) Bottom: Breakdown of continent-of-origin for participants in
+  participant. Bottom: Breakdown of continent-of-origin for participants in
   each conference.  Colored bars represent percentages of participants whose

### 2019-07-16 main.tex
Working on Richard comments

@@
 % \end{figure}
+
+\bcp{Richard Kim sent a bunch of great comments!  In particular...
+  \begin{itemize}
+  \item (a) You might say something more about how ACM is in a good position
+  to set a trend -- something about taking leadership among scientific
+  societies or being on the forefront. There's a little bit about this at
+  the end, but it might be developed more into a sentence or paragraph. You
+  probably want to phrase this as leadership as opposed to say, sticking
+  one's neck out, but there could be something about how computer science &
+  tech folks are current and forward-looking, and it's consistent that we
+  not be behind the leading edge.
+  \item (b) You might say more how ACM needs to prepare itself in light of
+  the trends that are coming -- carbon pricing is good organizational sense!
+  There's some of this around ACM's potentially needing to adjust its
+  business model, but it falls short of making a *business case for carbon
+  pricing*, which is what I guess I'm recommending.
+  \item Also, in addition to mentioning that Microsoft and other tech
+  companies set a price on carbon, you might also note here that Exxon and
+  other energy companies set an internal price on carbon as well, as a means
+  of risk assessment. In other words, even the oil companies know what's
+  coming.
+  \end{itemize}
+}
 
@@
 
+\bcp{See comments at this point about companies (MS, Google?) that are
+  already doing this.}
+% You probably know about Microsoft using an internal price for carbon  - it's mentioned here:
+% https://unfccc.int/climate-action/momentum-for-change/financing-for-climate-friendly/microsoft-global-carbon-fee
+% And this report is also instructive:
+% https://download.microsoft.com/download/0/A/B/0AB2FDD7-BDD9-4E23-AF6B-9417A8691CF5/Microsoft%20Carbon%20Fee%20Impact.pdf
+
 \PARAGRAPH{Precipitating change}
@@
 
-The climate crisis is too important to ignore and too urgent to leave for
+The climate crisis is too urgent to leave for
 world leaders to address at their own pace: Organizations at every scale

### 2019-08-22 main.tex
Note
COmments

--- 2019-07-01 main.tex
@@
 \newif\ifdraft\drafttrue    % comments / todos for now
-\newif\iflater\laterfalse  % comments / todos for final version
+\newif\iflater\latertrue    % comments / todos for final version
 
@@
 % \end{figure}
+
+\bcp{Consider all of Mike's cuts.}
+
+\bcp{Take out details of registration data.}
 
@@
 
+\bcp{Close with a discussion-starter based on the second reviewer's last
+  review.}
+
 \SECTION{Conclusion}
@@
 and establish new ways of doing business in the lower-carbon future that is
-coming at us.  We in ACM must play our part.
+coming at us.  We in ACM must step up and play our part.
 
@@
 our carbon report?
-\end{itemize}
 \item 
 Check where we've used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
+\end{itemize}
 \fi

### 2019-08-22 main.tex
Tweaks

@@
 \caption{Top: Carbon footprint per participant for travel to recent SIGPLAN
-  conferences\iflater\bcp{Get better file from the other paper, or get
-    better version of this one from Crista.}\fi.  The smallest dot (ICFP
+  conferences.  The smallest dot (ICFP
   14, in Gothenburg, Sweden) represents .9 tons of {\COtwoE} per participant; the
-  largest dot (ICFP 16, in Nara Japan) represents 1.94 tons per
+  largest dot (ICFP 16, in Nara, Japan) represents 1.94 tons per
   participant. Bottom: Breakdown of continent-of-origin for participants in

### 2019-09-20 main.tex
Some BCP comments

--- 2019-07-16 main.tex
@@
 Committee to research this question~\cite{ClimateCommittee}.  After
-investigating many options~\cite{ClimateCommitteReport} and discussing
-intensively with members, conference organizers, ACM staff, and leaders of
-other SIGs, we are putting forward two concrete proposals. First, all ACM
+investigating many options~\cite{ClimateCommitteReport} and intensive and
+wide-ranging discussions, we are putting forward two concrete
+proposals. First, all ACM 
 conferences should publicly {\em account} for the {\COtwoE} emitted as a
-result of putting them on. Second, ACM \iflater\bcp{Are we really
+result of putting them on---in particular, from travel to
+the conference. Second, ACM \iflater\bcp{Are we really
   talking about ACM, in the first instance?  Or is the proposal really that
@@
 Our first proposal is modest: {\bf Every ACM-sponsored conference should
-publicly report its carbon footprint.} 
+publicly report its carbon footprint}.
+Critically, this accounting should include the emissions generated by participants' travel to and from each conference, as this will constitute the bulk of the conference's carbon footprint.
 %
@@
 Many organizations, including companies such as Google, Dell, Microsoft,
-General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
-and some academic societies, now buy carbon offsets in order to reduce their
+General Motors, Delta Airlines, Lyft, and Expedia, universities,
+and academic societies, now buy carbon offsets in order to reduce their
 net footprint. ACM conferences can do the same, and the purchases can be
@@
   the right one? Link seems
-  dead!}\fi. Users can upload conference registration data, and the
+  dead!  But Richard says ``As a reader, my first question is, how do I get
+  access to this calculator? Can you cite this calculator in some way? It
+  would be great to give a citation to where people can learn about it or
+  use it, even if just, "To learn more about the calculator, e-mail
+  _________."''}\fi. Users can upload conference registration data, and the
 calculator will estimate the {\COtwoE} cost of air travel for those
@@
   each conference.  Colored bars represent percentages of participants whose
-  home city is in each continent.
+  home city is in each continent.  \iflater\bcp{Richard: I find this a
+    little hard to follow. Maybe you could simply label the different groups
+    of conferences by their location? I think the "(NA)", "(EU)", and
+    "(IAS)" are a bit hard to read and decode at first, so that I didn't
+    understand where each conference was located at first glance. (Adrianne
+    had the same response when I showed her.) Maybe just put text under
+    each, maybe with a horizontal bracket, that says "North America,"
+    "Europe," and "Rest of world," with the bracket and/or words color-coded
+    as well.}\fi 
 % Bottom: Distances traveled by participants, one way, in Km; middle dashed
@@
 
+\bcp{Richard: I get a little lost in this paragraph.
+Partly because I found the figure a bit confusing, partly because of the length of the paragraph.
+I think one fix might be to frontload your conclusion a bit with a topic sentence.
+E.g., start the paragraph with something like:
+"For example, choosing a conference location with the goal of minimizing CO2e emissions from participants' travel results in trade-offs in who participates."}
 To illustrate the challenges, consider the problem of choosing a conference
@@
 research community by discouraging participation from other parts of the
-world.  Indeed, one might conclude that, from the point of view of
+world.\bcp{Richard: The damage to the research community is a little unclear to me on first read.
+I would reword to something like "that always locating conferences in the same one or two places would significantly damage the heterogeneity of the research community by discouraging participation from other parts of the world".
+Or perhaps some better world, but more specificity of the damage.}  Indeed, one might conclude that, from the point of view of
 strengthening the research community, conferences should move around as 

### 2019-09-20 main.tex

@@
 }
+
+\bcp{Consider all of Mike's cuts.}
+
+\bcp{Take out details of registration data.}
 

### 2019-10-15 main.tex
Fix

@@
   probably want to phrase this as leadership as opposed to say, sticking
-  one's neck out, but there could be something about how computer science &
+  one's neck out, but there could be something about how computer science and
   tech folks are current and forward-looking, and it's consistent that we
@@
   would be great to give a citation to where people can learn about it or
-  use it, even if just, "To learn more about the calculator, e-mail
-  _________."''}\fi. Users can upload conference registration data, and the
+  use it, even if just, "To learn more about the calculator, e-mail xxx."''}\fi. 
+Users can upload conference registration data, and the
 calculator will estimate the {\COtwoE} cost of air travel for those

### 2019-10-15 main.tex
Comments

@@
  would be great to give a citation to where people can learn about it or
-  use it, even if just, "To learn more about the calculator, e-mail xxx."''}\fi. 
+ use it, even if just, "To learn more about the calculator, e-mail xxx"''}\fi. 
 Users can upload conference registration data, and the

### 2019-10-15 main.tex
CP

@@
 }
-
-\bcp{Consider all of Mike's cuts.}
-
-\bcp{Take out details of registration data.}
 
@@
 Many organizations, including companies such as Google, Dell, Microsoft,
-General Motors, Delta Airlines, Lyft, and Expedia, universities,
-and academic societies, now buy carbon offsets in order to reduce their
+General Motors, Delta Airlines, Lyft, and Expedia, as well as universities,
+academic societies, and even energy companies such as Exxon, now use
+carbon offsets in order to reduce their 
 net footprint. ACM conferences can do the same, and the purchases can be
@@
 other forms of conference emissions.
+
+We acknowledge that there is some the reputational risk to ACM in taking the
+step of publicizing its carbon footprint: the numbers are likely to be high,
+and may be used to criticize both the ACM and the broader academic
+community.  Nevertheless, making this information available is a crucial
+first step: we cannot manage what we have not measured.
 
@@
 
-% \SECTION{Prepare for Deeper Changes}
+\PARAGRAPH{What should the price be?}
 
-% Our two concrete proposals aim to make conferences much more
-% carbon-efficient, allowing ACM to play its part in the 40\% reduction in
-% emissions that is needed over the next decade.  But this is just a first
-% step.  Looking further ahead, the challenge will be to eliminate {\em all}
-% greenhouse emissions by around 2050 and then move into global negative
-% emissions during the second half of the century.
-
-% \bcp{I wish we could put a paragraph here about the long-term picture and
-%   the need to rethink scientific exchange and ACM's business model.  That
-%   will please our grumpy reviewer, and I think it's actually a point worth
-%   making.\cite{NCN}}
-
-\bcp{Close with a discussion-starter based on the second reviewer's last
-  review.}
+A key issue in implementation of this proposal will be how to set a price
+that reflects the true social cost of carbon, without unduly harming the
+community.
+%
+Initial data gathering will play a role, and a steady and predictable annual
+ramp on the price is a necessary component, but setting both the initial
+price and the slope of the ramp will likely be difficult political
+decisions.
+%
+One measurable target, in line with the latest evidence from climate
+science, would be to halve emissions every decade from now on, following the
+``Carbon Law'' proposed in a recent paper~\cite{Rockstrom}.  \bcp{Read the
+  paper and cite it!}
+% https://doi.org/10.1126/science.aah3443
 
@@
 
-The climate crisis is too urgent to leave for
-world leaders to address at their own pace: Organizations at every scale
-must confront their own contributions, raise awareness among their members,
-and establish new ways of doing business in the lower-carbon future that is
-coming at us.  We in ACM must step up and play our part.
+The climate crisis is too urgent to leave for world leaders to address at
+their own pace: Organizations at every scale, including ACM, must confront
+their own contributions, raise awareness among their members, and establish
+new ways of doing business in the lower-carbon future that is coming at us.
 

### 2019-10-15 main.tex
Possible cuts. Some trim.

--- 2019-10-15 main.tex
@@
   probably want to phrase this as leadership as opposed to say, sticking
-  one's neck out, but there could be something about how computer science and
+  one's neck out, but there could be something about how computer science \&
   tech folks are current and forward-looking, and it's consistent that we
@@
 publicly report its carbon footprint}.
-Critically, this accounting should include the emissions generated by participants' travel to and from each conference, as this will constitute the bulk of the conference's carbon footprint.
+\mwh{CUT: Critically, this accounting should include the emissions generated by participants' travel to and from each conference, as this will constitute the bulk of the conference's carbon footprint.}
 %
@@
 flagship conferences (POPL and ICFP) to switch from in-person to on-line
-program committee meetings (joining a trend among other SIGs, most of which
-already use virtual PC meetings for their flagship conferences). This
+program committee meetings (joining a trend among other SIGs). This
 discussion has also prompted several conferences to increase investments in
@@
   tool~\cite{anderson2012inconvenient,
- carbon-offsets-are-not-our-get-out-jail-free-card}.  Besides issues with
+  carbon-offsets-are-not-our-get-out-jail-free-card}.  \mwh{CUT: Besides issues with
   verifiability and permanence, some observers worry that---especially at
@@
   leading organizations to postpone facing the difficult changes that will
- be required to sustainably decarbonize society.  Offsets
- should be considered as a useful short-term expedient, to be considered in
- parallel with, not instead of, more difficult cutting.
+  be required to sustainably decarbonize society.} Offsets are a
+useful short-term expedient, not a replacement for more difficult
+cutting. 
 % MWH: It's a bit odd to have two lengthy parenthetical asides in the same
@@
   would be great to give a citation to where people can learn about it or
- use it, even if just, "To learn more about the calculator, e-mail xxx"''}\fi. 
-Users can upload conference registration data, and the
+  use it, even if just, "To learn more about the calculator, e-mail
+  \_\_\_\_\_\_\_\_\_."''}\fi. Users can upload conference registration data, and the
 calculator will estimate the {\COtwoE} cost of air travel for those

### 2019-10-15 main.tex
CP

@@
 {\em total} yearly emissions for an average resident of the US (16.5 tons)
-or Europe (7 tons)~\cite{emissions}.  \iflater\bcp{For comparison, ...  (put
- in some figures about the footprints of a few other things that people
- might think of reducing---driving, eating meat, upgrading their
- fridge...)}\fi{} Moreover, these emissions have no near-term technological
+or Europe (7 tons)~\cite{emissions}.  
+Moreover, these emissions have no near-term technological
 fix, since jet fuel is difficult to replace with renewable energy
@@
 result of putting them on---in particular, from travel to
-the conference. Second, ACM \iflater\bcp{Are we really
- talking about ACM, in the first instance?  Or is the proposal really that
- individual SIGs should do this, until there is enough buy-in to do it at
- the ACM level?  The first is easier to say, but the second is the only
- plausible implementation path.  If there's space for at least a sentence
- or two on this (and other practicalities of implementation), it might be
- good.}\fi should put a \emph{price} on 
+the conference. Second, ACM should put a \emph{price} on 
 carbon in conference budgets, to create pressure on organizers to reduce

### 2019-10-15 main.tex
Another one

--- 2019-10-15 main.tex
@@
 Beyond encouraging conferences to trim their emissions, public 
-accounting will generate useful data for future organizers.  To this end,
+accounting will generate useful data for future organizers.  \mwh{CUT: To this end,
 emissions data should be published both in aggregate 
@@
 form, with an anonymized record for each participant in each
-conference, including their originating city.  The data should be made
+conference, including their originating city.}  The data should be made
 available in 

### 2019-10-15 main.tex
CP

--- 2019-10-15 main.tex
@@
 publicly report its carbon footprint}.  
-\mwh{CUT: Critically, this accounting should include the emissions generated by participants' travel to and from each conference, as this will constitute the bulk of the conference's carbon footprint.}
 %
@@
   tool~\cite{anderson2012inconvenient,
-  carbon-offsets-are-not-our-get-out-jail-free-card}.  \mwh{CUT: Besides issues with
-  verifiability and permanence, some observers worry that---especially at
-  their current artificially low cost---they will encourage complacency,
-  leading organizations to postpone facing the difficult changes that will
-  be required to sustainably decarbonize society.} Offsets are a
+  carbon-offsets-are-not-our-get-out-jail-free-card}.  Offsets are a
 useful short-term expedient, not a replacement for more difficult

### 2019-10-15 main.tex
Small
CP

--- 2019-10-15 main.tex
@@
 Beyond encouraging conferences to trim their emissions, public 
-accounting will generate useful data for future organizers.  To this end,
-emissions data should be published both in aggregate 
-form (total and per-participant footprints), for quick reference, and in raw
-form, with an anonymized record for each participant in each
-conference, including their originating city.  The data should be made
+accounting will generate useful data for future organizers.  
+The data should be made
 available in 
@@
 
-ACM should develop tools to gather and publicize this data. As a first step,
+ACM should develop tools to gather and publicize this data. 
 SIGPLAN recently built an air-travel-focused {\COtwoE} calculator for
@@
 reductions, it is not likely, by itself, to induce major changes.
-
-\PARAGRAPH{Hard choices}
-
+%
 Science is a fundamentally social process, and the conference system
@@
 \end{figure}
+
+\PARAGRAPH{Hard choices}
 

### 2019-10-15 main.tex
Change to match blog

--- 2019-10-15 main.tex
@@
 
+\PARAGRAPH{Hard choices}
+
 \bcp{Richard: I get a little lost in this paragraph.
@@
 %
+Eyeballing this diagram, it seems that carbon-conscious organizers
+should hold all of these conferences in either the northeast US or
+western Europe every year.
+%
+The second diagram tells a different story. The horizontal colored bar
+at the bottom represents the continent on which each conference was
+held, and each vertical bar gives a breakdown of the participants in
+that conference, colored according to the continent of their work
+address. From a superficial glance at the colors, it is clear that,
+though a minority come from far away, the majority of participants are
+local to the region where each conference is being held. This suggests
+that locating conferences every year in the same one or two places
+would significantly damage the research community by discouraging
+participation from other parts of the world. Indeed, one might
+conclude that, from the point of view of strengthening the research
+community, conferences should move around as much as possible. These
+disparate perspectives suggest that reducing conference emissions will
+require genuinely painful compromises. The impulse to avoid thinking
+about it is entirely understandable!
+\mwh{OLD:
 The top diagram shows an
@@
 research community by discouraging participation from other parts of the
-world.\bcp{Richard: The damage to the research community is a little unclear to me on first read.
+world.}\bcp{Richard: The damage to the research community is a little unclear to me on first read.
 I would reword to something like "that always locating conferences in the same one or two places would significantly damage the heterogeneity of the research community by discouraging participation from other parts of the world".

### 2019-10-15 main.tex

@@
 Beyond encouraging conferences to trim their emissions, public 
-accounting will generate useful data for future organizers.  \mwh{CUT: To this end,
-emissions data should be published both in aggregate 
-form (total and per-participant footprints), for quick reference, and in raw
-form, with an anonymized record for each participant in each
-conference, including their originating city.}  The data should be made
+accounting will generate useful data for future organizers.  
+The data should be made
 available in 

### 2019-10-15 main.tex
CP

@@
 
-\bcp{Richard Kim sent a bunch of great comments!  In particular...
-  \begin{itemize}
-  \item (a) You might say something more about how ACM is in a good position
-  to set a trend -- something about taking leadership among scientific
-  societies or being on the forefront. There's a little bit about this at
-  the end, but it might be developed more into a sentence or paragraph. You
-  probably want to phrase this as leadership as opposed to say, sticking
-  one's neck out, but there could be something about how computer science and
- tech folks are current and forward-looking, and it's consistent that we
- not be behind the leading edge.
- \item (b) You might say more how ACM needs to prepare itself in light of
- the trends that are coming -- carbon pricing is good organizational sense!
- There's some of this around ACM's potentially needing to adjust its
- business model, but it falls short of making a *business case for carbon
- pricing*, which is what I guess I'm recommending.
- \item Also, in addition to mentioning that Microsoft and other tech
- companies set a price on carbon, you might also note here that Exxon and
- other energy companies set an internal price on carbon as well, as a means
- of risk assessment. In other words, even the oil companies know what's
- coming.
- \end{itemize}
-}
-
 A broad scientific consensus warns that human
@@
 the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
-passenger\iflater\bcp{Citation?}\fi, which is a significant fraction of the
+passenger\iflater\bcp{Citation?}\fi, a significant fraction of the
 {\em total} yearly emissions for an average resident of the US (16.5 tons)

### 2019-10-15 main.tex
CP

@@
 
-\bcp{Richard: I get a little lost in this paragraph.
-Partly because I found the figure a bit confusing, partly because of the length of the paragraph.
-I think one fix might be to frontload your conclusion a bit with a topic sentence.
-E.g., start the paragraph with something like:
-"For example, choosing a conference location with the goal of minimizing CO2e emissions from participants' travel results in trade-offs in who participates."}
+% \bcp{Richard: I get a little lost in this paragraph. Partly because I
+% found the figure a bit confusing, partly because of the length of the
+% paragraph. I think one fix might be to frontload your conclusion a bit
+% with a topic sentence. E.g., start the paragraph with something like: "For
+% example, choosing a conference location with the goal of minimizing CO2e
+% emissions from participants' travel results in trade-offs in who
+% participates."}
+
 To illustrate the challenges, consider the problem of choosing a conference

### 2019-10-15 main.tex
More small cuts.

@@
 %
-Eyeballing this diagram, it seems that carbon-conscious organizers
-should hold all of these conferences in either the northeast US or
-western Europe every year.
-%
-The second diagram tells a different story. The horizontal colored bar
-at the bottom represents the continent on which each conference was
-held, and each vertical bar gives a breakdown of the participants in
-that conference, colored according to the continent of their work
-address. From a superficial glance at the colors, it is clear that,
-though a minority come from far away, the majority of participants are
-local to the region where each conference is being held. This suggests
-that locating conferences every year in the same one or two places
-would significantly damage the research community by discouraging
-participation from other parts of the world. Indeed, one might
-conclude that, from the point of view of strengthening the research
-community, conferences should move around as much as possible. These
-disparate perspectives suggest that reducing conference emissions will
-require genuinely painful compromises. The impulse to avoid thinking
-about it is entirely understandable!
-\mwh{OLD:
 The top diagram shows an
@@
 research community by discouraging participation from other parts of the
-world.}\bcp{Richard: The damage to the research community is a little unclear to me on first read.
+world.\bcp{Richard: The damage to the research community is a little unclear to me on first read.
 I would reword to something like "that always locating conferences in the same one or two places would significantly damage the heterogeneity of the research community by discouraging participation from other parts of the world".
@@
 
-Ideally, at some point, governments will impose carbon pricing uniformly,
-and all {\COtwoE}-intensive activities will have to pay it. But ACM can send a
-strong message about the importance of this issue by acting now. In
-particular, as several companies have shown, even an artificial ``internal
-price'' on carbon sends a significant price signal that shapes behavior.
-%
 Thus, our second proposal is that {\bf ACM should impose a surcharge on
@@
   choices won't necessarily be bad, but could be really great?}\fi
+
+Ideally, at some point, governments will impose carbon pricing uniformly,
+and all {\COtwoE}-intensive activities will have to pay it. But ACM can send a
+strong message about the importance of this issue by acting now.
 
@@
 
-Beyond simple belt-tightening, carbon pricing will stimulate creative
-rethinking of the conference model 
+Carbon pricing will stimulate creative rethinking of the conference model 
 itself---for example, seriously considering alternatives such as

### 2019-10-15 main.tex
CP

@@
 Thus, our second proposal is that {\bf ACM should impose a surcharge on
-  conferences based on their carbon footprint.}  The charge should be low to
-start and increase steadily and predictably, year on year.
+  conferences based on their carbon footprint.}  The charge should start low 
+and increase steadily and predictably, year on year.
 %
@@
 participation, increasing registration fees, soliciting additional corporate
-sponsorship, or other means.  \iflater\mwh{Problem: We showed just above
-  that it's hard to figure out what to do. Come back to this point here and
-  say that the surcharge will spur bold choices?  Maybe say that these
-  choices won't necessarily be bad, but could be really great?}\fi
+sponsorship, or other means.
 
@@
 
-\bcp{See comments at this point about companies (MS, Google?) that are
-  already doing this.}
+% \bcp{See comments at this point about companies (MS, Google?) that are
+%   already doing this.}
 % You probably know about Microsoft using an internal price for carbon  - it's mentioned here:
@@
 science, would be to halve emissions every decade from now on, following the
-``Carbon Law'' proposed in a recent paper~\cite{Rockstrom}.  \bcp{Read the
-  paper and cite it!}
-% https://doi.org/10.1126/science.aah3443
+``Carbon Law'' proposed in a recent paper~\cite{Rockstrom1269}.  \bcp{Read the
+  paper and cite it!
+  https://doi.org/10.1126/science.aah3443
+  }
 

### 2019-12-10 main.tex
CP
Note

@@
 
-How should ACM respond to these facts?  Two years ago, ACM's Special
+How should ACM respond to these facts?  Two years ago, \mwh{Three, now?} ACM's Special
 Interest Group on Programming Languages (SIGPLAN) convened an ad hoc Climate
@@
 and all {\COtwoE}-intensive activities will have to pay it. But ACM can send a
-strong message about the importance of this issue by acting now.
+strong message about the importance of this issue---and get ahead of the
+coming changes---by acting now.
 
@@
 could make emissions reduction politically problematic unless ACM's
-conference-focused business model is also adjusted.  Planning for this
-eventuality should begin now!
+conference-focused business model is also adjusted.
 

### 2019-12-21 main.tex
A few nites

@@
 
-\newcommand{\bcp}[1]{\ifdraft{\bf [bcp: #1]}\fi}
-\newcommand{\mwh}[1]{\ifdraft{\bf [mwh: #1]}\fi}
+\usepackage{xcolor}
+\definecolor{darkred}{rgb}{0.7, 0.0, 0.0}
+\newcommand{\bcp}[1]{\ifdraft{\bf\textcolor{darkred}{[\bf bcp: #1]}}\fi}
+\newcommand{\mwh}[1]{\ifdraft{\bf\textcolor{darkred}{[\bf mwh: #1]}}\fi}
 \newcommand{\COtwoE}{CO$_2$e}
@@
 
-\title{\Huge Conferences in the Era of Expensive Carbon}
+\title{\Huge Conferences in an Era of Expensive Carbon}
 % \title{\Huge ACM Conferences and Cost of Carbon}
@@
 the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
-passenger\iflater\bcp{Citation?}\fi, a significant fraction of the
+passenger\iflater\bcp{Citation?  Yes, to the same carbon calculator that we
+  recommend elsewhere!}\fi, a significant fraction of the
 {\em total} yearly emissions for an average resident of the US (16.5 tons)
@@
 
-How should ACM respond to these facts?  Two years ago, \mwh{Three, now?} ACM's Special
+How should ACM respond to these facts?  In 2016, ACM's Special
 Interest Group on Programming Languages (SIGPLAN) convened an ad hoc Climate
@@
 offset during registration.  Other scientific societies are going further,
-rolling carbon offsets into the registration fees for all
-participants.\iflater\bcp{In a later draft, we may be able to say that ICML,
-NIPS, etc. are a case in point.  Ask Alexandre Lacoste about this.}\fi
+rolling a carbon offset into the registration fee for every participant.
 

### 2019-12-22 main.tex
Polishing

@@
 % (Figure \ref{fig:warming}).
-This is not a problem that can be left to future
-generations: The UN's Intergovernmental
+This is not a problem for future
+generations, but a present-day emergency: The UN's Intergovernmental
 Panel on Climate Change (IPCC) says a 40\% decrease in
@@
 % (Figure~\ref{fig:IPCC-cover}).
-Achieving reductions on this scale will
+Reductions on this scale will
 require urgent and sustained commitment at all levels of society---not only
@@
 Indeed, scientific societies arguably have an especially important role to
-play, since, for many of their members, travel to conferences may be a
+play since, for many of their members, travel to conferences represents
 substantial or even dominant part of their individual contribution to
 climate change.  A single round-trip flight from Philadelphia to Paris emits
-the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
-passenger\iflater\bcp{Citation?  Yes, to the same carbon calculator that we
-  recommend elsewhere!}\fi, a significant fraction of the
+the equivalent of about 1.8 tons of carbon dioxide (\COtwoE) per
+passenger~\cite{carboncalculator}, a significant fraction of the
 {\em total} yearly emissions for an average resident of the US (16.5 tons)
@@
 Moreover, these emissions have no near-term technological
-fix, since jet fuel is difficult to replace with renewable energy
+fix, as jet fuel is difficult to replace with renewable energy
 sources~\cite{elec-air}.
@@
 Committee to research this question~\cite{ClimateCommittee}.  After
-investigating many options~\cite{ClimateCommitteReport} and intensive and
-wide-ranging discussions, we are putting forward two concrete
-proposals. First, all ACM
+investigating many options~\cite{ClimateCommitteReport}, we are putting
+forward two concrete proposals. First, all ACM
 conferences should publicly {\em account} for the {\COtwoE} emitted as a
@@
 the conference. Second, ACM should put a \emph{price} on
-carbon in conference budgets, to create pressure on organizers to reduce
+carbon in conference budgets, creating a steady pressure on organizers to reduce
 their footprints.
@@
 publicly report its carbon footprint}.
-%
+
 Public accounting will nudge conference organizers and attendees to change
@@
 As a case in point, SIGPLAN has been considering how to reduce emissions
-over the past couple of years, informed by an accounting of its own
+over the past few years, informed by an accounting of its own
 {\COtwoE} expenditures. This discussion has led the organizers of two of its
 flagship conferences (POPL and ICFP) to switch from in-person to on-line
-program committee meetings (joining a trend among other SIGs). This
-discussion has also prompted several conferences to increase investments in
+program committee meetings, joining a trend among other SIGs, and
+has prompted several conferences to increase investments in
 livestreaming and video recording to support remote participation.
@@
 %
-A carbon offset is sold by a vendor who uses the funds to finance an
+A carbon offset is sold by a vendor, who uses the funds to finance an
 activity that permanently removes or avoids emitting some amount of
 greenhouse gases.  The veracity and permanence of this activity is certified
-by a watchdog organization. (Typically, planting trees is not a certifiable
-activity, since it's hard to guarantee the trees won't be cut down, but
+by a watchdog organization. (E.g., planting trees is typically not a certifiable
+activity, since it's hard to guarantee they won't be cut down, but
 installing methane capture devices on landfill sites or buying
@@
 academic societies, and even energy companies such as Exxon, now use
-carbon offsets in order to reduce their
-net footprint. ACM conferences can do the same, and the purchases can be
-mentioned in the public accounting we are proposing.  In fact, ACM recently
-decided that conferences could offer attendees the option to pay for an
-offset during registration.  Other scientific societies are going further,
-rolling a carbon offset into the registration fee for every participant.
+carbon offsets to reduce their
+net footprint. ACM conferences should do the same, and the purchases should be
+included in the public accounting we are proposing.  Indeed, ACM recently
+declared that conferences can offer attendees the option to pay for an
+offset during registration.  Other scientific societies and conferences are
+going further, rolling a carbon offset into the registration fee for every
+participant.
 
@@
   tool~\cite{anderson2012inconvenient,
-  carbon-offsets-are-not-our-get-out-jail-free-card}.  Offsets are a
+  carbon-offsets-are-not-our-get-out-jail-free-card}.  They should be
+regarded as a
 useful short-term expedient, not a replacement for more difficult
-cutting.
-% MWH: It's a bit odd to have two lengthy parenthetical asides in the same
-% paragraph. I think this one is on point, so need not be an aside.
-% BCP: The whole paragraph was in parens, before.  However, the new paragraph
-% header does a decent job of "closing the aside," so I agree the parens can
-% go.
-
-% \mwh{Should
-% we mention the FCRC numbers here? 45 out of 2,700 used {\COtwoE} offsets}
-% That's getting too much down into the details, I think -- also, these
-% numbers reflect a hastily organized, preliminary version of what we really
-% want to do.
+cutting.  \bcp{Needs a bit of smoothing.}
 
@@
 SIGPLAN recently built an air-travel-focused {\COtwoE} calculator for
-conferences\iflater~\cite{carboncalculator}\bcp{Trim citation?  Is it even
- the right one? Link seems
- dead!  But Richard says ``As a reader, my first question is, how do I get
- access to this calculator? Can you cite this calculator in some way? It
- would be great to give a citation to where people can learn about it or
- use it, even if just, "To learn more about the calculator, e-mail xxx"''}\fi.
+conferences\iflater~\cite{ClimateCommittee}.
 Users can upload conference registration data, and the
@@
 majority of participants in each conference are local to the region where the
-conference is being held.\iflater\bcp{That point could be strengthened by adding some
-  numbers (just in the text, not a figure) about what percentage of
-  conference attendees attend even one additional instance of the same
-  conference...}\fi{} This suggests that locating
-conferences every year in the same one or two places would significantly
-damage the
+conference is being held.
+% \iflater\bcp{That point could be strengthened by adding some
+%   numbers (just in the text, not a figure) about what percentage of
+%   conference attendees attend even one additional instance of the same
+%   conference...}\fi{}
+ This suggests that always locating conferences in the
+same one or two places would significantly damage the heterogeneity of the
 research community by discouraging participation from other parts of the
-world.\bcp{Richard: The damage to the research community is a little unclear to me on first read.
-I would reword to something like "that always locating conferences in the same one or two places would significantly damage the heterogeneity of the research community by discouraging participation from other parts of the world".
-Or perhaps some better world, but more specificity of the damage.}  Indeed, one might conclude that, from the point of view of
+world.  Indeed, one might conclude that, from the point of view of
 strengthening the research community, conferences should move around as
-much as possible.
+much as possible!
 %
@@
 One measurable target, in line with the latest evidence from climate
-science, would be to halve emissions every decade from now on, following the
-``Carbon Law'' proposed in a recent paper~\cite{Rockstrom1269}.  \bcp{Read the
-  paper and cite it!
-  https://doi.org/10.1126/science.aah3443
-  }
+science, would be to halve actual emissions every decade from now on,
+following a recently proposed
+``Carbon Law''~\cite{Rockstrom1269}.  \bcp{Read the
+  paper!}
 
@@
 
-The climate crisis is too urgent to leave for world leaders to address at
+The climate crisis is too urgent to leave to world leaders to address at
 their own pace: Organizations at every scale, including ACM, must confront
 their own contributions, raise awareness among their members, and establish
-new ways of doing business in the lower-carbon future that is coming at us.
+new ways of doing business in the lower-carbon future that is now upon us.
+ACM should do its part by mandating public accounting of conference carbon
+footprints and by putting a concrete price on the carbon that they use.
 
@@
   \item do we need any more citations for the claims we make?
-  \item Find a place for this: {\tt acm-climate} mailing list~\cite{acm-climate}
+  \item Find a place for the {\tt acm-climate} mailing list~\cite{acm-climate}
   \item Write a cover letter

### 2019-12-24 main.tex
More polishing

@@
 % (Figure \ref{fig:warming}).
-This is not a problem for future
-generations, but a present-day emergency: The UN's Intergovernmental
+This is a present-day emergency: the UN's Intergovernmental
 Panel on Climate Change (IPCC) says a 40\% decrease in
@@
 % (Figure~\ref{fig:IPCC-cover}).
-Reductions on this scale will
+Reductions on this scale
 require urgent and sustained commitment at all levels of society---not only
@@
 
-Indeed, scientific societies arguably have an especially important role to
-play since, for many of their members, travel to conferences represents
+Indeed, scientific societies have an especially important role to
+play, since, for many members, travel to conferences represents a
 substantial or even dominant part of their individual contribution to
@@
 the equivalent of about 1.8 tons of carbon dioxide (\COtwoE) per
-passenger~\cite{carboncalculator}, a significant fraction of the
+passenger~\cite{carboncalculator}, which is a significant fraction of the
 {\em total} yearly emissions for an average resident of the US (16.5 tons)
@@
 Interest Group on Programming Languages (SIGPLAN) convened an ad hoc Climate
-Committee to research this question~\cite{ClimateCommittee}.  After
+Committee to consider this question~\cite{ClimateCommittee}.  After
 investigating many options~\cite{ClimateCommitteReport}, we are putting
@@
 publicly report its carbon footprint}.
+%
+These reports should be collected in a central place, in a uniform format.
+Most conferences' footprints will be dominated by participants' air travel,
+but the data gathered should go beyond this to include ground
+transportation, travel to in-person PC meetings, and estimated emissions
+from hotel and food.
 
-Public accounting will nudge conference organizers and attendees to change
-their behavior. By analogy, chain restaurants in the US are required by law
-to post calorie counts of food items on menu boards, and studies show that
-enlightened customers order, on average, up to 50 fewer calories a day as a
-result~\cite{menu}.
-%MWH: I don't know why this should be a parenthetical.  Why would the
-%analogy be an aside? It is meant to connect the reader's understanding to
-%what we are proposing, explciitly; it is meant to be read, not skipped
-%over.
-% BCP: In general, parenthetical remarks are not meant to be skipped over:
-% they are meant to be read as asides.  I think this reads better in parens,
-% but I won't fight for it.
+ACM should develop tools to gather and publicize this data.  For example,
+SIGPLAN recently built an air-travel-focused {\COtwoE} calculator for
+conferences\iflater~\cite{ClimateCommittee}.  Users can upload conference
+registration data, and the calculator will estimate the {\COtwoE} cost of
+air travel for those sufficiently far away. This calculator could be
+expanded to account for other forms of conference emissions.
 
-As a case in point, SIGPLAN has been considering how to reduce emissions
-over the past few years, informed by an accounting of its own
-{\COtwoE} expenditures. This discussion has led the organizers of two of its
-flagship conferences (POPL and ICFP) to switch from in-person to on-line
-program committee meetings, joining a trend among other SIGs, and
-has prompted several conferences to increase investments in
-livestreaming and video recording to support remote participation.
+There is some the reputational risk to ACM in taking the step of publicizing
+its carbon footprint: the numbers are likely to be high, and they may be
+used to criticize both the ACM and the broader academic community.  But
+making this information available is a crucial first step: we cannot manage
+what we have not measured.  And ACM is well positioned to develop tools that
+other academic societies can adopt.
 
-\PARAGRAPH{Carbon offsets}
-%
-SIGPLAN has also explored how to mitigate conference emissions by purchasing
-\emph{carbon offsets}~\cite{CarbonOFfsetReport}.
-%
-A carbon offset is sold by a vendor, who uses the funds to finance an
-activity that permanently removes or avoids emitting some amount of
-greenhouse gases.  The veracity and permanence of this activity is certified
-by a watchdog organization. (E.g., planting trees is typically not a certifiable
-activity, since it's hard to guarantee they won't be cut down, but
-installing methane capture devices on landfill sites or buying
-fuel-efficient stoves to replace open-fire cooking in poor communities can
-be.)
+\PARAGRAPH{Easy trimming}
+One effect of public accounting will be to nudge conference organizers and
+attendees to change their behavior. By analogy, chain restaurants in the US
+are now required by law to post calorie counts of food items on menu boards;
+studies show that enlightened customers order, on average, up to 50 fewer
+calories a day~\cite{menu}.
 
-Many organizations, including companies such as Google, Dell, Microsoft,
-General Motors, Delta Airlines, Lyft, and Expedia, as well as universities,
-academic societies, and even energy companies such as Exxon, now use
-carbon offsets to reduce their
-net footprint. ACM conferences should do the same, and the purchases should be
-included in the public accounting we are proposing.  Indeed, ACM recently
-declared that conferences can offer attendees the option to pay for an
-offset during registration.  Other scientific societies and conferences are
-going further, rolling a carbon offset into the registration fee for every
-participant.
+Similarly, SIGPLAN has been considering how to reduce emissions over the
+past few years, informed by an accounting of its own {\COtwoE}
+emissions. This discussion has led the organizers of two flagship
+conferences (POPL and ICFP) to switch from in-person to online program
+committee meetings, joining a trend among other SIGs, and has prompted
+several conferences to increase investments in livestreaming and video
+recording to support remote participation.
 
-We acknowledge that carbon offsets are a
-  controversial
-  tool~\cite{anderson2012inconvenient,
-  carbon-offsets-are-not-our-get-out-jail-free-card}.  They should be
-regarded as a
-useful short-term expedient, not a replacement for more difficult
-cutting.  \bcp{Needs a bit of smoothing.}
-
-\PARAGRAPH{Data gathering}
-%
-Beyond encouraging conferences to trim their emissions, public
-accounting will generate useful data for future organizers.
-The data should be made
-available in
-a central place, in a uniform format.  Most
-conferences' footprints will be dominated by participants' air travel, but the
-data gathered should go beyond this to include air travel to in-person PC
-meetings and estimated hotel and food {\COtwoE} footprints.
-
-ACM should develop tools to gather and publicize this data.
-SIGPLAN recently built an air-travel-focused {\COtwoE} calculator for
-conferences\iflater~\cite{ClimateCommittee}.
-Users can upload conference registration data, and the
-calculator will estimate the {\COtwoE} cost of air travel for those
-sufficiently far away. This calculator could be expanded to account for
-other forms of conference emissions.
-
-We acknowledge that there is some the reputational risk to ACM in taking the
-step of publicizing its carbon footprint: the numbers are likely to be high,
-and may be used to criticize both the ACM and the broader academic
-community.  Nevertheless, making this information available is a crucial
-first step: we cannot manage what we have not measured.
-
-\SECTION{Put a Price on Carbon}
-
-While public accounting of {\COtwoE} expenditures will incentivize some
-reductions, it is not likely, by itself, to induce major changes.
-%
-Science is a fundamentally social process, and the conference system
-accelerates scientific research through high-bandwidth interaction,
-direct dissemination of results, network building, and serendipitous
-cross-fertilization.
-%
-Organizers and attendees will be reticent to make changes that might
-threaten these benefits.
-
-\begin{figure}[t]
+\begin{figure}[t!]
 \centering
@@
   14, in Gothenburg, Sweden) represents .9 tons of {\COtwoE} per participant; the
-  largest dot (ICFP 16, in Nara, Japan) represents 1.94 tons per
+  largest (ICFP 16, in Nara, Japan) represents 1.94 tons per
   participant. Bottom: Breakdown of continent-of-origin for participants in
   each conference.  Colored bars represent percentages of participants whose
-  home city is in each continent.  \iflater\bcp{Richard: I find this a
-    little hard to follow. Maybe you could simply label the different groups
-    of conferences by their location? I think the "(NA)", "(EU)", and
-    "(IAS)" are a bit hard to read and decode at first, so that I didn't
-    understand where each conference was located at first glance. (Adrianne
-    had the same response when I showed her.) Maybe just put text under
-    each, maybe with a horizontal bracket, that says "North America,"
-    "Europe," and "Rest of world," with the bracket and/or words color-coded
-    as well.}\fi
+  home city is in each continent: blue for North America, orange for Europe,
+  green for Asia.
+% \iflater\bcp{Richard: I find this a
+%     little hard to follow. Maybe you could simply label the different groups
+%     of conferences by their location? I think the "(NA)", "(EU)", and
+%     "(IAS)" are a bit hard to read and decode at first, so that I didn't
+%     understand where each conference was located at first glance. (Adrianne
+%     had the same response when I showed her.) Maybe just put text under
+%     each, maybe with a horizontal bracket, that says "North America,"
+%     "Europe," and "Rest of world," with the bracket and/or words color-coded
+%     as well.}\fi
+%
 % Bottom: Distances traveled by participants, one way, in Km; middle dashed
@@
 \PARAGRAPH{Hard choices}
+
+However, while public accounting of {\COtwoE} will encourage easy
+reductions, it is not likely, by itself, to induce major shifts in behavior.
+%
+Science is a fundamentally social process, and the conference system
+accelerates scientific research through high-bandwidth interaction,
+direct dissemination of results, network building, and serendipitous
+cross-fertilization.
+%
+Organizers and attendees will naturally be reluctant to consider changes
+that might threaten these benefits.
 
@@
 location that minimizes emissions from participants' travel.  Using recent
-registration data from four SIGPLAN conferences, Figure~\ref{fig:confs}
+registration data from four \mbox{SIGPLAN} conferences, Figure~\ref{fig:confs}
 shows two ways of looking at the relation between locations and emissions.
@@
 conference over the past ten years (excluding a few for which we
-had trouble getting data), with larger dots representing more emissions.
+had trouble getting data), with larger dots representing higher emissions.
 Eyeballing this diagram, it seems that carbon-conscious
@@
  This suggests that always locating conferences in the
-same one or two places would significantly damage the heterogeneity of the
+same one or two places would significantly impact the diversity of the
 research community by discouraging participation from other parts of the
@@
 %
-These disparate perspectives suggest that reducing conference emissions will
-require genuinely painful compromises.  The impulse to avoid thinking about
-it is entirely understandable!
+These disparate perspectives suggest that reducing conference emissions may
+require genuinely painful compromises.  The impulse to ignore the issue is
+entirely understandable.
 
@@
 
-\PARAGRAPH{A cost for carbon}
+\SECTION{Put a Price on Carbon}
 %
 This dilemma is a microcosm of one faced by all of society. To address it,
-many policy experts advocate some form of carbon pricing (either taxes or
-cap-and-trade schemes), imposing a concrete, immediate cost on
+many policy experts advocate using some form of carbon pricing to impose a
+concrete, immediate cost on
 emissions~\cite{carbonprice}. Doing so makes manifest the hidden
 environmental cost of emissions, incentivizing
-{\COtwoE}-reducing changes without specifying exactly which, and thus
+{\COtwoE}-reducing changes without mandating exactly which ones, and thus
 allowing for creative and efficient responses. Continuing the junk-food
-analogy, some
-municipalities such as New York city and Philadelphia have imposed a
+analogy,
+some municipalities including New York City and Philadelphia have imposed a
 per-calorie tax on soft drinks, with some studies finding that doing so
 indeed reduced consumption~\cite{sodatax}.
-%MWH: The above is not a parenthetical, so the earlier junkfood analogy should not be, either
 
@@
 sponsorship, or other means.
+In this process, a primary concern
+should be to find ways of {reducing} the financial burden of such a
+surcharge on those who may be disproportionately affected by it---students
+without grant support, participants from developing areas, etc.
+Wealthier participants should subsidize the carbon footprints of less
+wealthy ones.
 
@@
 %
-What should ACM do with the funds collected from this surcharge?  One can
-imagine many good uses, including carbon offsetting, defraying the costs of
+What should ACM do with the funds collected from this surcharge?  One
+obvious possibility is purchasing \emph{carbon offsets}~\cite{CarbonOFfsetReport}.
+%
+A carbon offset is sold by a vendor, who uses the funds to finance an
+activity that permanently removes or avoids emitting some amount of
+greenhouse gases.  The veracity and permanence of this activity is certified
+by a watchdog organization. (E.g., planting trees is typically not a certifiable
+activity, since it's hard to guarantee they won't be cut down, but
+installing methane capture devices on landfill sites or buying
+fuel-efficient stoves to replace open-fire cooking in poor communities can
+be.)
+Many organizations, including companies such as Google, Dell, Microsoft,
+General Motors, Delta Airlines, Lyft, and Expedia, as well as universities,
+academic societies, and even energy companies such as Exxon, now use
+carbon offsets to reduce their
+net footprint. ACM conferences should consider doing the same, and the
+purchases should be
+included in the public accounting we are proposing.  Indeed, ACM recently
+declared that conferences can offer attendees the option to pay for an
+offset during registration.  Other scientific societies and conferences are
+going further, rolling a carbon offset into the registration fee for every
+participant.
+
+Beyond buying offsets, one can
+imagine many good uses for the funds generated from a carbon surcharge: defraying the costs of
 virtualizing conferences (livestreaming, etc.), and supporting ``green''
@@
 best replace or approximate the socializing and networking aspects of
-conferences in a virtual setting.  But the very first concern
-should be to find ways of {\em reducing} the financial burden of such a
-surcharge on those who may be disproportionately affected by it---students
-without grant support, participants from developing areas, etc.  That is,
-wealthier participants should subsidize the carbon footprints of less
-wealthy ones.
+conferences in a virtual setting.
 
-Carbon pricing will stimulate creative rethinking of the conference model
-itself---for example, seriously considering alternatives such as
-rapid-turnaround, journal-only publishing models, once-a-year
-megaconferences, and entirely virtual conferences~\cite{NCN}.
+Ultimately, however, carbon offsets and other ``good works'' cannot
+substitute for real reductions in emissions~\cite{anderson2012inconvenient,
+  carbon-offsets-are-not-our-get-out-jail-free-card}: they are, at best, a
+short-term expedient that buys a little time to develop strategies for more
+difficult cuts.
+%
+Indeed, the main goal of carbon pricing should be to stimulate creative
+rethinking of the conference model itself---for example, seriously
+considering alternatives such as rapid-turnaround journal-only publishing
+models, once-a-year megaconferences, and entirely virtual
+conferences~\cite{NCN}.
 %
@@
 
-A key issue in implementation of this proposal will be how to set a price
-that reflects the true social cost of carbon, without unduly harming the
-community.
+Another key issue in implementation of this second proposal will be how to
+set a price that reflects the true social cost of carbon, without unduly
+harming the scientific community.
 %
@@
 One measurable target, in line with the latest evidence from climate
-science, would be to halve actual emissions every decade from now on,
-following a recently proposed
-``Carbon Law''~\cite{Rockstrom1269}.  \bcp{Read the
-  paper!}
+science, would be to tune the parameters with a goal of halving actual
+emissions every decade from now on, following a recently proposed ``Carbon
+Law''~\cite{Rockstrom1269}.
 
@@
 new ways of doing business in the lower-carbon future that is now upon us.
-ACM should do its part by mandating public accounting of conference carbon
-footprints and by putting a concrete price on the carbon that they use.
+We in ACM should do our part by mandating public accounting of conference
+carbon footprints and by putting a concrete price on the carbon that we use.
 
@@
   \item Find a place for the {\tt acm-climate} mailing list~\cite{acm-climate}
-  \item Write a cover letter
-  \begin{itemize}
-  \item We tried hard to remove incrementalism and make it much more
-  hard hitting
-  \item In particular, we make two concrete proposals
-  \item However, we did keep the suggestion of carbon offsets, hopefully
-  in a more convincing / palatable / better argued form
-  \item what else?
-  \end{itemize}
+  \item Recheck and fix citations
+  \item scan the stuff below and delete it
+  % \item Write a cover letter
+  % \begin{itemize}
+  % \item We tried hard to remove incrementalism and make it much more
+  % hard hitting
+  % \item In particular, we make two concrete proposals
+  % \item However, we did keep the suggestion of carbon offsets, hopefully
+  % in a more convincing / palatable / better argued form
+  % \item what else?
+  % \end{itemize}
   \end{itemize}

### 2019-12-26 main.tex
Nother round of polishing

@@
 \newif\ifdraft\drafttrue    % comments / todos for now
-\newif\iflater\latertrue    % comments / todos for final version
+\newif\iflater\laterfalse    % comments / todos for final version
 
@@
 climate change.  A single round-trip flight from Philadelphia to Paris emits
-the equivalent of about 1.8 tons of carbon dioxide (\COtwoE) per
-passenger~\cite{carboncalculator}, which is a significant fraction of the
+the equivalent of about 1.8 tons of carbon dioxide (\COtwoE, or informally
+``carbon'') per
+passenger~\cite{carboncalculator}. This is a significant fraction of the
 {\em total} yearly emissions for an average resident of the US (16.5 tons)
@@
 sources~\cite{elec-air}.
-% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
-% of conferences will require a reduction in the footprint of the attendees'
-% air travel.}\fi
 
-How should ACM respond to these facts?  In 2016, ACM's Special
+How should ACM respond to these facts?
+
+In 2016, ACM's Special
 Interest Group on Programming Languages (SIGPLAN) convened an ad hoc Climate
@@
 but the data gathered should go beyond this to include ground
-transportation, travel to in-person PC meetings, and estimated emissions
-from hotel and food.
+transportation, travel to in-person program committee meetings, and
+estimated emissions from hotels and food.
 
 ACM should develop tools to gather and publicize this data.  For example,
-SIGPLAN recently built an air-travel-focused {\COtwoE} calculator for
-conferences\iflater~\cite{ClimateCommittee}.  Users can upload conference
+SIGPLAN recently built an air-travel-focused carbon calculator for
+conferences~\cite{ClimateCommittee}.  Users can upload conference
 registration data, and the calculator will estimate the {\COtwoE} cost of
-air travel for those sufficiently far away. This calculator could be
-expanded to account for other forms of conference emissions.
+air travel for those sufficiently far away.
+% This calculator could be
+% expanded to account for other forms of conference emissions.
 
@@
 making this information available is a crucial first step: we cannot manage
-what we have not measured.  And ACM is well positioned to develop tools that
-other academic societies can adopt.
+what we have not measured.
 
@@
 
-Similarly, SIGPLAN has been considering how to reduce emissions over the
-past few years, informed by an accounting of its own {\COtwoE}
-emissions. This discussion has led the organizers of two flagship
+Similarly, SIGPLAN has been considering how to reduce emissions, informed by
+an accounting of its own carbon footprint. This discussion has led the
+organizers of two flagship
 conferences (POPL and ICFP) to switch from in-person to online program
@@
   green for Asia.
-% \iflater\bcp{Richard: I find this a
+% \bcp{Richard: I find this a
 %     little hard to follow. Maybe you could simply label the different groups
@@
 
-However, while public accounting of {\COtwoE} will encourage easy
+However, while public accounting of emissions will encourage easy
 reductions, it is not likely, by itself, to induce major shifts in behavior.
@@
 had trouble getting data), with larger dots representing higher emissions.
-Eyeballing this diagram, it seems that carbon-conscious
+Eyeballing this diagram, it might seem that carbon-conscious
 organizers should hold all of these conferences in either the northeast US or
@@
 a breakdown of the participants in that conference, colored according to the
-continent of their work address.  From a superficial glance
-at the colors, it is clear that, though a minority come from far away, the
+continent of their work address.  A glance
+at the colors makes clear that, though a minority come from far away, the
 majority of participants in each conference are local to the region where the
 conference is being held.
-% \iflater\bcp{That point could be strengthened by adding some
-%   numbers (just in the text, not a figure) about what percentage of
-%   conference attendees attend even one additional instance of the same
-%   conference...}\fi{}
+%
 This suggests that always locating conferences in the
@@
 %
-These disparate perspectives suggest that reducing conference emissions may
+These disparate perspectives suggest that significantly reducing conference
+emissions may
 require genuinely painful compromises.  The impulse to ignore the issue is
@@
 analogy,
-some municipalities including New York City and Philadelphia have imposed a
-per-calorie tax on soft drinks, with some studies finding that doing so
-indeed reduced consumption~\cite{sodatax}.
+some municipalities including Berkeley and Philadelphia have imposed a
+per-calorie tax on soft drinks; studies found that doing so
+significantly reduced consumption~\cite{sodatax}.
 
@@
 by decreasing per-participant emissions, decreasing (physical)
-participation, increasing registration fees, soliciting additional corporate
+participation, increasing registration fees, soliciting corporate
 sponsorship, or other means.
@@
 should be to find ways of {reducing} the financial burden of such a
-surcharge on those who may be disproportionately affected by it---students
+surcharge on those disproportionately affected by it---students
 without grant support, participants from developing areas, etc.
-Wealthier participants should subsidize the carbon footprints of less
+Well-funded participants should subsidize the carbon surcharges of less
 wealthy ones.
@@
 Ideally, at some point, governments will impose carbon pricing uniformly,
-and all {\COtwoE}-intensive activities will have to pay it. But ACM can send a
+and all carbon-intensive activities will have to pay it. But ACM can send a
 strong message about the importance of this issue---and get ahead of the
@@
 greenhouse gases.  The veracity and permanence of this activity is certified
-by a watchdog organization. (E.g., planting trees is typically not a certifiable
-activity, since it's hard to guarantee they won't be cut down, but
+by a watchdog organization. (E.g., planting trees is often considered not to
+be a certifiable
+activity, since it's hard to guarantee they won't be cut down;
+reductions from
 installing methane capture devices on landfill sites or buying
-fuel-efficient stoves to replace open-fire cooking in poor communities can
-be.)
+fuel-efficient stoves to replace open-fire cooking in poor communities are
+easier to predict.)
 Many organizations, including companies such as Google, Dell, Microsoft,
@@
 purchases should be
-included in the public accounting we are proposing.  Indeed, ACM recently
-declared that conferences can offer attendees the option to pay for an
-offset during registration.  Other scientific societies and conferences are
-going further, rolling a carbon offset into the registration fee for every
-participant.
+included in the public accounting we are proposing.
+% For example\bcp{beter
+%   transition?}, ACM recently
+% declared that conferences can offer attendees the option to pay for an
+% offset during registration.  Other scientific societies and conferences are
+% going further, rolling a carbon offset into the registration fee for every
+% participant.
 
@@
 virtualizing conferences (livestreaming, etc.), and supporting ``green''
-computing research. As an example of the last, ACM could help fund a
+computing research~\cite{rolnick2019tackling}.  As an example
+of the last, ACM could help fund a
 cross-cutting research initiative specifically aimed at understanding how to
@@
   carbon-offsets-are-not-our-get-out-jail-free-card}: they are, at best, a
-short-term expedient that buys a little time to develop strategies for more
+short-term expedient that buys time to agree on more
 difficult cuts.
@@
 considering alternatives such as rapid-turnaround journal-only publishing
-models, once-a-year megaconferences, and entirely virtual
+models, yearly megaconferences, and entirely virtual
 conferences~\cite{NCN}.
@@
 
-Another key issue in implementation of this second proposal will be how to
+Another key issue in implementing this second proposal will be how to
 set a price that reflects the true social cost of carbon, without unduly
@@
 Initial data gathering will play a role, and a steady and predictable annual
-ramp on the price is a necessary component, but setting both the initial
+increase is a necessary component, but setting both the initial
 price and the slope of the ramp will likely be difficult political
@@
 their own pace: Organizations at every scale, including ACM, must confront
-their own contributions, raise awareness among their members, and establish
-new ways of doing business in the lower-carbon future that is now upon us.
-We in ACM should do our part by mandating public accounting of conference
-carbon footprints and by putting a concrete price on the carbon that we use.
+their own contributions, raise awareness and foster discussion among their
+members~\cite{acm-climate}, and establish new ways of doing business in the
+lower-carbon future that is now upon us.  We in ACM should do our part by
+mandating public accounting of conference carbon footprints and by putting a
+concrete price on the carbon that we use.
 
@@
 
+\bibliographystyle{unsrt}
+\bibliography{local}
+
+\iflater
+
 \iflater
@@
 
-\bibliographystyle{unsrt}
-\bibliography{local}
-
-\iflater
 \newpage

### 2019-12-26 main.tex
Final changes for CACM version

@@
 ``carbon'') per
-passenger~\cite{carboncalculator}. This is a significant fraction of the
+passenger~\cite{ClimateCommittee}. This is a significant fraction of the
 {\em total} yearly emissions for an average resident of the US (16.5 tons)
@@
 considering alternatives such as rapid-turnaround journal-only publishing
-models, yearly megaconferences, and entirely virtual
+models, yearly mega-conferences, and entirely virtual
 conferences~\cite{NCN}.

### 2019-12-26 main.tex
Use better images from Crista

@@
 \includegraphics[width=6in]{SIGPLAN-confs2.png}
-\includegraphics[width=6in]{ParticipantsOriginAll2.png}
+\includegraphics[width=6in]{ParticipantsOriginAll3.png}
 % \includegraphics[width=6in]{Distances.png}
@@
 their own pace: Organizations at every scale, including ACM, must confront
-their own contributions, raise awareness and foster discussion among their
+their own contributions, raise awareness and build consensus among their
 members~\cite{acm-climate}, and establish new ways of doing business in the

### 2020-02-10 main.tex
Notes

--- 2019-12-26 main.tex
@@
 \newif\ifdraft\drafttrue    % comments / todos for now
-\newif\iflater\laterfalse    % comments / todos for final version
+\newif\iflater\latertrue    % comments / todos for final version
 
@@
 
-\usepackage{xcolor}
-\definecolor{darkred}{rgb}{0.7, 0.0, 0.0}
-\newcommand{\bcp}[1]{\ifdraft{\bf\textcolor{darkred}{[\bf bcp: #1]}}\fi}
-\newcommand{\mwh}[1]{\ifdraft{\bf\textcolor{darkred}{[\bf mwh: #1]}}\fi}
+\newcommand{\bcp}[1]{\ifdraft{\bf [bcp: #1]}\fi}
+\newcommand{\mwh}[1]{\ifdraft{\bf [mwh: #1]}\fi}
 \newcommand{\COtwoE}{CO$_2$e}
@@
 % (Figure \ref{fig:warming}).
-This is a present-day emergency: the UN's Intergovernmental
+This is not a problem that can be left to future
+generations: The UN's Intergovernmental
 Panel on Climate Change (IPCC) says a 40\% decrease in
@@
 % (Figure~\ref{fig:IPCC-cover}).
-Reductions on this scale
+Achieving reductions on this scale will
 require urgent and sustained commitment at all levels of society---not only
@@
 
-Indeed, scientific societies have an especially important role to
-play, since, for many members, travel to conferences represents a
+Indeed, scientific societies arguably have an especially important role to
+play, since, for many of their members, travel to conferences may be a
 substantial or even dominant part of their individual contribution to
 climate change.  A single round-trip flight from Philadelphia to Paris emits
-the equivalent of about 1.8 tons of carbon dioxide (\COtwoE, or informally
-``carbon'') per
-passenger~\cite{carboncalculator}. This is a significant fraction of the
+the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
+passenger\iflater\bcp{Citation?}\fi, a significant fraction of the
 {\em total} yearly emissions for an average resident of the US (16.5 tons)
@@
 Moreover, these emissions have no near-term technological
-fix, as jet fuel is difficult to replace with renewable energy
+fix, since jet fuel is difficult to replace with renewable energy
 sources~\cite{elec-air}.
+% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
+% of conferences will require a reduction in the footprint of the attendees'
+% air travel.}\fi
 
-How should ACM respond to these facts?
-
-In 2016, ACM's Special
+How should ACM respond to these facts?  Two years ago, \mwh{Three, now?} ACM's Special
 Interest Group on Programming Languages (SIGPLAN) convened an ad hoc Climate
-Committee to consider this question~\cite{ClimateCommittee}.  After
-investigating many options~\cite{ClimateCommitteReport}, we are putting
-forward two concrete proposals. First, all ACM
+Committee to research this question~\cite{ClimateCommittee}.  After
+investigating many options~\cite{ClimateCommitteReport} and intensive and
+wide-ranging discussions, we are putting forward two concrete
+proposals. First, all ACM 
 conferences should publicly {\em account} for the {\COtwoE} emitted as a
@@
 the conference. Second, ACM should put a \emph{price} on 
-carbon in conference budgets, creating a steady pressure on organizers to reduce
+carbon in conference budgets, to create pressure on organizers to reduce
 their footprints.
@@
 %
-These reports should be collected in a central place, in a uniform format.
-Most conferences' footprints will be dominated by participants' air travel,
-but the data gathered should go beyond this to include ground
-transportation, travel to in-person program committee meetings, and
-estimated emissions from hotels and food.
+Public accounting will nudge conference organizers and attendees to change
+their behavior. By analogy, chain restaurants in the US are required by law
+to post calorie counts of food items on menu boards, and studies show that
+enlightened customers order, on average, up to 50 fewer calories a day as a
+result~\cite{menu}.  
+%MWH: I don't know why this should be a parenthetical.  Why would the
+%analogy be an aside? It is meant to connect the reader's understanding to
+%what we are proposing, explciitly; it is meant to be read, not skipped
+%over. 
+% BCP: In general, parenthetical remarks are not meant to be skipped over:
+% they are meant to be read as asides.  I think this reads better in parens,
+% but I won't fight for it.
 
-ACM should develop tools to gather and publicize this data.  For example,
-SIGPLAN recently built an air-travel-focused carbon calculator for
-conferences~\cite{ClimateCommittee}.  Users can upload conference
-registration data, and the calculator will estimate the {\COtwoE} cost of
-air travel for those sufficiently far away.
-% This calculator could be
-% expanded to account for other forms of conference emissions.
+As a case in point, SIGPLAN has been considering how to reduce emissions
+over the past couple of years, informed by an accounting of its own
+{\COtwoE} expenditures. This discussion has led the organizers of two of its
+flagship conferences (POPL and ICFP) to switch from in-person to on-line
+program committee meetings (joining a trend among other SIGs). This
+discussion has also prompted several conferences to increase investments in
+livestreaming and video recording to support remote participation.
 
-There is some the reputational risk to ACM in taking the step of publicizing
-its carbon footprint: the numbers are likely to be high, and they may be
-used to criticize both the ACM and the broader academic community.  But
-making this information available is a crucial first step: we cannot manage
-what we have not measured.
+\PARAGRAPH{Carbon offsets}
+%
+SIGPLAN has also explored how to mitigate conference emissions by purchasing
+\emph{carbon offsets}~\cite{CarbonOFfsetReport}.
+%
+A carbon offset is sold by a vendor who uses the funds to finance an
+activity that permanently removes or avoids emitting some amount of
+greenhouse gases.  The veracity and permanence of this activity is certified
+by a watchdog organization. (Typically, planting trees is not a certifiable
+activity, since it's hard to guarantee the trees won't be cut down, but
+installing methane capture devices on landfill sites or buying
+fuel-efficient stoves to replace open-fire cooking in poor communities can
+be.)
 
-\PARAGRAPH{Easy trimming}
-One effect of public accounting will be to nudge conference organizers and
-attendees to change their behavior. By analogy, chain restaurants in the US
-are now required by law to post calorie counts of food items on menu boards;
-studies show that enlightened customers order, on average, up to 50 fewer
-calories a day~\cite{menu}.
+Many organizations, including companies such as Google, Dell, Microsoft,
+General Motors, Delta Airlines, Lyft, and Expedia, as well as universities,
+academic societies, and even energy companies such as Exxon, now use
+carbon offsets in order to reduce their 
+net footprint. ACM conferences can do the same, and the purchases can be
+mentioned in the public accounting we are proposing.  In fact, ACM recently
+decided that conferences could offer attendees the option to pay for an
+offset during registration.  Other scientific societies are going further,
+rolling carbon offsets into the registration fees for all
+participants.\iflater\bcp{In a later draft, we may be able to say that ICML,
+NIPS, etc. are a case in point.  Ask Alexandre Lacoste about this.}\fi
 
-Similarly, SIGPLAN has been considering how to reduce emissions, informed by
-an accounting of its own carbon footprint. This discussion has led the
-organizers of two flagship
-conferences (POPL and ICFP) to switch from in-person to online program
-committee meetings, joining a trend among other SIGs, and has prompted
-several conferences to increase investments in livestreaming and video
-recording to support remote participation.
+We acknowledge that carbon offsets are a 
+  controversial 
+  tool~\cite{anderson2012inconvenient,
+  carbon-offsets-are-not-our-get-out-jail-free-card}.  Offsets are a
+useful short-term expedient, not a replacement for more difficult
+cutting. 
+% MWH: It's a bit odd to have two lengthy parenthetical asides in the same
+% paragraph. I think this one is on point, so need not be an aside.
+% BCP: The whole paragraph was in parens, before.  However, the new paragraph
+% header does a decent job of "closing the aside," so I agree the parens can
+% go.
 
-\begin{figure}[t!]
+% \mwh{Should
+% we mention the FCRC numbers here? 45 out of 2,700 used {\COtwoE} offsets}
+% That's getting too much down into the details, I think -- also, these
+% numbers reflect a hastily organized, preliminary version of what we really
+% want to do.
+
+\PARAGRAPH{Data gathering}
+%
+Beyond encouraging conferences to trim their emissions, public 
+accounting will generate useful data for future organizers.  
+The data should be made
+available in 
+a central place, in a uniform format.  Most
+conferences' footprints will be dominated by participants' air travel, but the
+data gathered should go beyond this to include air travel to in-person PC
+meetings and estimated hotel and food {\COtwoE} footprints.
+
+ACM should develop tools to gather and publicize this data. 
+SIGPLAN recently built an air-travel-focused {\COtwoE} calculator for
+conferences\iflater~\cite{carboncalculator}\bcp{Trim citation?  Is it even
+ the right one? Link seems
+ dead!  But Richard says ``As a reader, my first question is, how do I get
+ access to this calculator? Can you cite this calculator in some way? It
+ would be great to give a citation to where people can learn about it or
+ use it, even if just, "To learn more about the calculator, e-mail xxx"''}\fi. 
+Users can upload conference registration data, and the
+calculator will estimate the {\COtwoE} cost of air travel for those
+sufficiently far away. This calculator could be expanded to account for
+other forms of conference emissions.
+
+We acknowledge that there is some the reputational risk to ACM in taking the
+step of publicizing its carbon footprint: the numbers are likely to be high,
+and may be used to criticize both the ACM and the broader academic
+community.  Nevertheless, making this information available is a crucial
+first step: we cannot manage what we have not measured.
+
+\SECTION{Put a Price on Carbon}
+
+While public accounting of {\COtwoE} expenditures will incentivize some 
+reductions, it is not likely, by itself, to induce major changes.
+%
+Science is a fundamentally social process, and the conference system
+accelerates scientific research through high-bandwidth interaction,
+direct dissemination of results, network building, and serendipitous
+cross-fertilization.
+%
+Organizers and attendees will be reticent to make changes that might
+threaten these benefits.
+
+\begin{figure}[t]
 \centering
@@
   14, in Gothenburg, Sweden) represents .9 tons of {\COtwoE} per participant; the
-  largest (ICFP 16, in Nara, Japan) represents 1.94 tons per
+  largest dot (ICFP 16, in Nara, Japan) represents 1.94 tons per
   participant. Bottom: Breakdown of continent-of-origin for participants in
   each conference.  Colored bars represent percentages of participants whose
-  home city is in each continent: blue for North America, orange for Europe,
-  green for Asia.
-% \bcp{Richard: I find this a
-%     little hard to follow. Maybe you could simply label the different groups
-%     of conferences by their location? I think the "(NA)", "(EU)", and
-%     "(IAS)" are a bit hard to read and decode at first, so that I didn't
-%     understand where each conference was located at first glance. (Adrianne
-%     had the same response when I showed her.) Maybe just put text under
-%     each, maybe with a horizontal bracket, that says "North America,"
-%     "Europe," and "Rest of world," with the bracket and/or words color-coded
-%     as well.}\fi
-%
+  home city is in each continent.  \iflater\bcp{Richard: I find this a
+    little hard to follow. Maybe you could simply label the different groups
+    of conferences by their location? I think the "(NA)", "(EU)", and
+    "(IAS)" are a bit hard to read and decode at first, so that I didn't
+    understand where each conference was located at first glance. (Adrianne
+    had the same response when I showed her.) Maybe just put text under
+    each, maybe with a horizontal bracket, that says "North America,"
+    "Europe," and "Rest of world," with the bracket and/or words color-coded
+    as well.}\fi 
 % Bottom: Distances traveled by participants, one way, in Km; middle dashed
@@
 \PARAGRAPH{Hard choices}
-
-However, while public accounting of emissions will encourage easy
-reductions, it is not likely, by itself, to induce major shifts in behavior.
-%
-Science is a fundamentally social process, and the conference system
-accelerates scientific research through high-bandwidth interaction,
-direct dissemination of results, network building, and serendipitous
-cross-fertilization.
-%
-Organizers and attendees will naturally be reluctant to consider changes
-that might threaten these benefits.
 
@@
 location that minimizes emissions from participants' travel.  Using recent
-registration data from four \mbox{SIGPLAN} conferences, Figure~\ref{fig:confs}
+registration data from four SIGPLAN conferences, Figure~\ref{fig:confs}
 shows two ways of looking at the relation between locations and emissions. 
@@
 conference over the past ten years (excluding a few for which we
-had trouble getting data), with larger dots representing higher emissions.
-Eyeballing this diagram, it might seem that carbon-conscious
+had trouble getting data), with larger dots representing more emissions.
+Eyeballing this diagram, it seems that carbon-conscious 
 organizers should hold all of these conferences in either the northeast US or
@@
 a breakdown of the participants in that conference, colored according to the
-continent of their work address.  A glance
-at the colors makes clear that, though a minority come from far away, the
+continent of their work address.  From a superficial glance
+at the colors, it is clear that, though a minority come from far away, the
 majority of participants in each conference are local to the region where the
-conference is being held.
-%
-This suggests that always locating conferences in the
-same one or two places would significantly impact the diversity of the
+conference is being held.\iflater\bcp{That point could be strengthened by adding some
+  numbers (just in the text, not a figure) about what percentage of
+  conference attendees attend even one additional instance of the same
+  conference...}\fi{} This suggests that locating
+conferences every year in the same one or two places would significantly
+damage the 
 research community by discouraging participation from other parts of the
-world.  Indeed, one might conclude that, from the point of view of
+world.\bcp{Richard: The damage to the research community is a little unclear to me on first read.
+I would reword to something like "that always locating conferences in the same one or two places would significantly damage the heterogeneity of the research community by discouraging participation from other parts of the world".
+Or perhaps some better world, but more specificity of the damage.}  Indeed, one might conclude that, from the point of view of
 strengthening the research community, conferences should move around as 
-much as possible!
+much as possible.
 %
@@
 %
-These disparate perspectives suggest that significantly reducing conference
-emissions may
-require genuinely painful compromises.  The impulse to ignore the issue is
-entirely understandable.
+These disparate perspectives suggest that reducing conference emissions will
+require genuinely painful compromises.  The impulse to avoid thinking about
+it is entirely understandable!
 
@@
 
-\SECTION{Put a Price on Carbon}
+\PARAGRAPH{A cost for carbon}
 %
 This dilemma is a microcosm of one faced by all of society. To address it,
-many policy experts advocate using some form of carbon pricing to impose a
-concrete, immediate cost on
+many policy experts advocate some form of carbon pricing (either taxes or
+cap-and-trade schemes), imposing a concrete, immediate cost on
 emissions~\cite{carbonprice}. Doing so makes manifest the hidden
 environmental cost of emissions, incentivizing
-{\COtwoE}-reducing changes without mandating exactly which ones, and thus
+{\COtwoE}-reducing changes without specifying exactly which, and thus
 allowing for creative and efficient responses. Continuing the junk-food
-analogy,
-some municipalities including Berkeley and Philadelphia have imposed a
-per-calorie tax on soft drinks; studies found that doing so
-significantly reduced consumption~\cite{sodatax}.
+analogy, some 
+municipalities such as New York city and Philadelphia have imposed a
+per-calorie tax on soft drinks, with some studies finding that doing so
+indeed reduced consumption~\cite{sodatax}.
+%MWH: The above is not a parenthetical, so the earlier junkfood analogy should not be, either
 
@@
 by decreasing per-participant emissions, decreasing (physical)
-participation, increasing registration fees, soliciting corporate
+participation, increasing registration fees, soliciting additional corporate
 sponsorship, or other means.
-In this process, a primary concern
-should be to find ways of {reducing} the financial burden of such a
-surcharge on those disproportionately affected by it---students
-without grant support, participants from developing areas, etc.
-Well-funded participants should subsidize the carbon surcharges of less
-wealthy ones.
 
 Ideally, at some point, governments will impose carbon pricing uniformly,
-and all carbon-intensive activities will have to pay it. But ACM can send a
+and all {\COtwoE}-intensive activities will have to pay it. But ACM can send a
 strong message about the importance of this issue---and get ahead of the
@@
  %
-What should ACM do with the funds collected from this surcharge?  One
-obvious possibility is purchasing \emph{carbon offsets}~\cite{CarbonOFfsetReport}.
-%
-A carbon offset is sold by a vendor, who uses the funds to finance an
-activity that permanently removes or avoids emitting some amount of
-greenhouse gases.  The veracity and permanence of this activity is certified
-by a watchdog organization. (E.g., planting trees is often considered not to
-be a certifiable
-activity, since it's hard to guarantee they won't be cut down;
-reductions from
-installing methane capture devices on landfill sites or buying
-fuel-efficient stoves to replace open-fire cooking in poor communities are
-easier to predict.)
-Many organizations, including companies such as Google, Dell, Microsoft,
-General Motors, Delta Airlines, Lyft, and Expedia, as well as universities,
-academic societies, and even energy companies such as Exxon, now use
-carbon offsets to reduce their
-net footprint. ACM conferences should consider doing the same, and the
-purchases should be
-included in the public accounting we are proposing.
-% For example\bcp{beter
-%   transition?}, ACM recently
-% declared that conferences can offer attendees the option to pay for an
-% offset during registration.  Other scientific societies and conferences are
-% going further, rolling a carbon offset into the registration fee for every
-% participant.
-
-Beyond buying offsets, one can
-imagine many good uses for the funds generated from a carbon surcharge: defraying the costs of
+What should ACM do with the funds collected from this surcharge?  One can
+imagine many good uses, including carbon offsetting, defraying the costs of
 virtualizing conferences (livestreaming, etc.), and supporting ``green''
-computing research~\cite{rolnick2019tackling}.  As an example
-of the last, ACM could help fund a
+computing research. As an example of the last, ACM could help fund a 
 cross-cutting research initiative specifically aimed at understanding how to
 best replace or approximate the socializing and networking aspects of
-conferences in a virtual setting.
+conferences in a virtual setting.  But the very first concern
+should be to find ways of {\em reducing} the financial burden of such a
+surcharge on those who may be disproportionately affected by it---students
+without grant support, participants from developing areas, etc.  That is,
+wealthier participants should subsidize the carbon footprints of less
+wealthy ones.
 
-Ultimately, however, carbon offsets and other ``good works'' cannot
-substitute for real reductions in emissions~\cite{anderson2012inconvenient,
-  carbon-offsets-are-not-our-get-out-jail-free-card}: they are, at best, a
-short-term expedient that buys time to agree on more
-difficult cuts.
-%
-Indeed, the main goal of carbon pricing should be to stimulate creative
-rethinking of the conference model itself---for example, seriously
-considering alternatives such as rapid-turnaround journal-only publishing
-models, yearly megaconferences, and entirely virtual
-conferences~\cite{NCN}.
+Carbon pricing will stimulate creative rethinking of the conference model 
+itself---for example, seriously considering alternatives such as
+rapid-turnaround, journal-only publishing models, once-a-year
+megaconferences, and entirely virtual conferences~\cite{NCN}.
 %
@@
 
-Another key issue in implementing this second proposal will be how to
-set a price that reflects the true social cost of carbon, without unduly
-harming the scientific community.
+A key issue in implementation of this proposal will be how to set a price
+that reflects the true social cost of carbon, without unduly harming the
+community.
 %
 Initial data gathering will play a role, and a steady and predictable annual
-increase is a necessary component, but setting both the initial
+ramp on the price is a necessary component, but setting both the initial
 price and the slope of the ramp will likely be difficult political
@@
 One measurable target, in line with the latest evidence from climate
-science, would be to tune the parameters with a goal of halving actual
-emissions every decade from now on, following a recently proposed ``Carbon
-Law''~\cite{Rockstrom1269}.
+science, would be to halve emissions every decade from now on, following the
+``Carbon Law'' proposed in a recent paper~\cite{Rockstrom1269}.  \bcp{Read the
+  paper and cite it!
+  https://doi.org/10.1126/science.aah3443
+  }
 
@@
 
-The climate crisis is too urgent to leave to world leaders to address at
+The climate crisis is too urgent to leave for world leaders to address at
 their own pace: Organizations at every scale, including ACM, must confront
-their own contributions, raise awareness and foster discussion among their
-members~\cite{acm-climate}, and establish new ways of doing business in the
-lower-carbon future that is now upon us.  We in ACM should do our part by
-mandating public accounting of conference carbon footprints and by putting a
-concrete price on the carbon that we use.
+their own contributions, raise awareness among their members, and establish
+new ways of doing business in the lower-carbon future that is coming at us.
 
@@
 
+\iflater
+\bcp{Things still missing:
+  \begin{itemize}
+  \item do we need any more citations for the claims we make?
+  \item Find a place for this: {\tt acm-climate} mailing list~\cite{acm-climate}
+  \end{itemize}
+}
+\fi
+
 \bibliographystyle{unsrt}
@@
 \iflater
-
-\iflater
-\bcp{Things still missing:
-  \begin{itemize}
-  \item do we need any more citations for the claims we make?
-  \item Find a place for the {\tt acm-climate} mailing list~\cite{acm-climate}
-  \item Recheck and fix citations
-  \item scan the stuff below and delete it
-  % \item Write a cover letter
-  % \begin{itemize}
-  % \item We tried hard to remove incrementalism and make it much more
-  % hard hitting
-  % \item In particular, we make two concrete proposals
-  % \item However, we did keep the suggestion of carbon offsets, hopefully
-  % in a more convincing / palatable / better argued form
-  % \item what else?
-  % \end{itemize}
-  \end{itemize}
-}
-\fi
-
 \newpage
